---
title: "Ash free dry weight"
author: "jillashey"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")
if (!require("dplyr")) install.packages("dplyr")
if (!require("plyr")) install.packages("plyr")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("formattable")) install.packages("formattable")
if (!require("data.table")) install.packages("data.table")
if (!require("htmltools")) install.packages("htmltools")
if (!require("webshot")) install.packages("webshot")
if (!require("agricolae")) install.packages("agricolae")
if (!require("sjPlot")) install.packages("sjPlot")
# load packages
library(plyr)
library(dplyr)
library(tidyverse)
library(plotrix)
library(Rmisc)
library(gridExtra)
library(formattable)
library(data.table)
library(htmltools)
library(webshot)
library(agricolae)
library(lme4)
library(car)
library(sjPlot)
```

Dry weight - organic and inorganic compounds 
Ash-free dry weight - organic compounds 

## Read in ash-free dry weight raw data, surface area, homogenate volume and metadata 
```{r}
AFDW <- read.csv("../data/Physiology/AFDW.csv", header = T, na.strings = "") %>% filter(!is.na(PlugID)) %>% dplyr::select(PlugID, Fraction, initial.mass.g, dry.pan.mass.g, burnt.pan.mass.g)

# Load surface area, homogenate volume and master fragment sheet 
surface_area <- read.csv("../data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)
homogenate_vol <- read.csv("../data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 
master <- read.csv("../data/MasterFragmentSheet.csv", header = T, na.strings = "") %>% filter(Metric == "P/R, molec") %>% dplyr::select(PlugID, Site, Treatment, Timepoint)

metadata <- join_all(list(AFDW, surface_area, homogenate_vol, master), by = "PlugID", type = "left")
#metadata$vol_mL <- gsub("NA", "", metadata$vol_mL)

# Remove field samples for now 
metadata_exp <- metadata %>%
  filter(Treatment != "Field")
```

Standardize for volume input
```{r}
sym <- 5
host <- 4

AFDW_df <- metadata_exp %>%
  mutate(dry.pan.mass.g.ml = case_when(Fraction=="Symbiont" ~ dry.pan.mass.g/sym, Fraction=="Host" ~dry.pan.mass.g/host),
         burnt.pan.mass.g.ml = case_when(Fraction=="Symbiont" ~ burnt.pan.mass.g/sym, Fraction=="Host" ~burnt.pan.mass.g/host))
```

Standardize and calculate tissue biomass
```{r}
AFDW_df <- AFDW_df %>%
  mutate(dry.pan.mass.g.vol.corr = dry.pan.mass.g.ml*vol_mL, # Standardize dry pan mass to homogenate volume
         burnt.pan.mass.g.vol.corr = burnt.pan.mass.g.ml*vol_mL, # Standardize burnt pan mass to homogenate volume
        dry.bioimass.g = (dry.pan.mass.g.vol.corr - initial.mass.g), # Calculate dry biomass
         DW.mg.cm2 = ((dry.bioimass.g)*1000) / Area_cm, # Calculate dry biomass standardized by surface area 
        dry.bioimass.g = (dry.pan.mass.g.vol.corr - burnt.pan.mass.g.vol.corr),
        AFDW.mg.cm2 = ((dry.bioimass.g)*1000) / Area_cm)
```

Write to csv 
```{r}
write.csv(AFDW_df, "../output/Physiology/AFDW/AFDW_biomass.csv")
```

# Stats and plotting - host 

Read in data and subset host 
```{r}
biomass <- read.csv("../output/Physiology/AFDW/AFDW_biomass.csv")

biomass_host <- biomass %>%
  #dplyr::mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", as.character(Treatment))) %>%
  dplyr::mutate(Treatment = as.factor(Treatment)) %>%
  dplyr::filter(Fraction == "Host") %>%
  na.omit()
```

Check for outliers 
```{r}
# Look for outliers visually
ggplot(biomass_host, aes(x = Treatment, y = AFDW.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(biomass_host$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_host$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- biomass_host[biomass_host$AFDW.mg.cm2 < lower_bound | biomass_host$AFDW.mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
biomass_host$z_score <- scale(biomass_host$AFDW.mg.cm2)
outliers_z <- biomass_host[abs(biomass_host$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(biomass_host$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_host$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
biomass_host_filtered <- biomass_host[biomass_host$AFDW.mg.cm2 >= lower_bound & biomass_host$AFDW.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(biomass_host)
dim(biomass_host_filtered)
```

6 outliers removed. 

Assess normality of data 
```{r}
# Histogram
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(biomass_host_filtered$AFDW.mg.cm2)
qqline(biomass_host_filtered$AFDW.mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(biomass_host_filtered$AFDW.mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
biomass_host_filtered$AFDW.mg.cm2_sqrt <- sqrt(biomass_host_filtered$AFDW.mg.cm2)
shapiro.test(biomass_host_filtered$AFDW.mg.cm2_sqrt)
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_host_filtered$AFDW.mg.cm2_sqrt)
qqline(biomass_host_filtered$AFDW.mg.cm2_sqrt, col = "red")

# log 
biomass_host_filtered$AFDW.mg.cm2_log <- log(biomass_host_filtered$AFDW.mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(biomass_host_filtered$AFDW.mg.cm2_log)  # Check normality
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_host_filtered$AFDW.mg.cm2_log)
qqline(biomass_host_filtered$AFDW.mg.cm2_log, col = "red")

# Inverse 
biomass_host_filtered$AFDW.mg.cm2_inv <- 1 / (biomass_host_filtered$AFDW.mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(biomass_host_filtered$AFDW.mg.cm2_inv)
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_host_filtered$AFDW.mg.cm2_inv)
qqline(biomass_host_filtered$AFDW.mg.cm2_inv, col = "red")
```

Both log and sqrt are normal. Going to stick with sqrt, as this has been the transformation that I have applied to other phys variables. 

Set TP0 to Ambient for stats purposes
```{r}
biomass_host_filtered <- biomass_host_filtered %>%
  mutate(Treatment = ifelse(Timepoint == "TP0", "Ambient", as.character(Treatment)))
```

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
biomass_host_filtered$Treatment <- as.factor(biomass_host_filtered$Treatment)
biomass_host_filtered$Timepoint <- as.factor(biomass_host_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(AFDW.mg.cm2_sqrt ~ Treatment * Timepoint, data = biomass_host_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(AFDW.mg.cm2_sqrt ~ Treatment * Timepoint, data = biomass_host_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

No significant differences 

Run post-hoc test for Treatment and Timepoint
```{r}
TukeyHSD(anova_result, "Treatment", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)

TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
biomass_host_filtered <- biomass_host_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

# Update to Acclimation for plotting purposes 
biomass_host_filtered <- biomass_host_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Plot untransformed data 
```{r}
# Set Month as factor 
biomass_host_filtered$Month <- factor(biomass_host_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

afdw_host <- ggplot(biomass_host_filtered, aes(x = Month, y = AFDW.mg.cm2, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted")+
  theme_classic() +
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
  theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(face = "bold", size = 30,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); afdw_host
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")

ggsave("../output/Physiology/AFDW/AFDW_Host_Treatment_Timepoint.pdf", afdw_host, width = 20, height = 15)
ggsave("../output/Physiology/AFDW/AFDW_Host_Treatment_Timepoint.png", afdw_host, width = 20, height = 15)
```

Plot sqrt data 
```{r}
# Set Month as factor 
biomass_host_filtered$Month <- factor(biomass_host_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

afdw_host_sqrt <- ggplot(biomass_host_filtered, aes(x = Month, y = AFDW.mg.cm2_sqrt, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted")+
  theme_classic() +
  xlab("") +
  ylab("Host Tissue Biomass (sqrt-transformed mg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
  theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(face = "bold", size = 30,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); afdw_host_sqrt
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")

ggsave("../output/Physiology/AFDW/AFDW_Host_Sqrt_Treatment_Timepoint.pdf", afdw_host_sqrt, width = 20, height = 15)
ggsave("../output/Physiology/AFDW/AFDW_Host_Sqrt_Treatment_Timepoint.png", afdw_host_sqrt, width = 20, height = 15)
```

Calculate mean, sd and sem of each group for plotting 
```{r}
biomass_host_summary <- biomass_host_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_host = mean(AFDW.mg.cm2, na.rm = TRUE),
    sd_host = sd(AFDW.mg.cm2, na.rm = TRUE),
    sem_host = sd_host / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Add letters for plotting 
```{r}
month_letters_manual <- tibble::tibble(
  Month = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"),
  Letters = c("a", "a", "a", "a", "a", "a", "a", "a"),
  y_position = c(9.24, 11.23, 10.81, 6.17, 11.19, 12.15, 9.88, 8.82) # 0.2 offset from highest sd 
)
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
biomass_host_summary$Month <- factor(biomass_host_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
host_dot_sem <- ggplot(biomass_host_summary, aes(x = Month, y = mean_host, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_host - sem_host, 
                    ymax = mean_host + sem_host), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  geom_text(data = month_letters_manual,
            aes(x = Month, y = y_position, label = Letters),
            inherit.aes = FALSE,
            position = position_dodge(width = 0.5),
            size = 10) +
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top"); host_dot_sem
ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sem.pdf", host_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sem.png", host_dot_sem, width = 20, height = 15)

# sd
host_dot_sd <- ggplot(biomass_host_summary, aes(x = Month, y = mean_host, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_host - sd_host, 
                    ymax = mean_host + sd_host),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  geom_text(data = month_letters_manual,
            aes(x = Month, y = y_position, label = Letters),
            inherit.aes = FALSE,
            position = position_dodge(width = 0.5),
            size = 10) +
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.position = "top",
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );host_dot_sd

ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sd.pdf", host_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sd.png", host_dot_sd, width = 22, height = 12)
```

# Stats and plotting - sym 

Read in data and subset sym 
```{r}
biomass <- read.csv("../output/Physiology/AFDW/AFDW_biomass.csv")

biomass_sym <- biomass %>%
  dplyr::mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", as.character(Treatment))) %>%
  dplyr::mutate(Treatment = as.factor(Treatment)) %>%
  dplyr::filter(Fraction == "Symbiont") %>%
  na.omit()
```

Check for outliers 
```{r}
# Look for outliers visually
ggplot(biomass_sym, aes(x = Treatment, y = AFDW.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(biomass_sym$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_sym$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- biomass_sym[biomass_sym$AFDW.mg.cm2 < lower_bound | biomass_sym$AFDW.mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
biomass_sym$z_score <- scale(biomass_sym$AFDW.mg.cm2)
outliers_z <- biomass_sym[abs(biomass_sym$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(biomass_sym$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_sym$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
biomass_sym_filtered <- biomass_sym[biomass_sym$AFDW.mg.cm2 >= lower_bound & biomass_sym$AFDW.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(biomass_sym)
dim(biomass_sym_filtered)
```

8 outliers removed. 

Assess normality of data 
```{r}
# Histogram
ggplot(biomass_sym_filtered, aes(x = AFDW.mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(biomass_sym_filtered$AFDW.mg.cm2)
qqline(biomass_sym_filtered$AFDW.mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(biomass_sym_filtered$AFDW.mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
biomass_sym_filtered$AFDW.mg.cm2_sqrt <- sqrt(biomass_sym_filtered$AFDW.mg.cm2)
shapiro.test(biomass_sym_filtered$AFDW.mg.cm2_sqrt)
ggplot(biomass_sym_filtered, aes(x = AFDW.mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_sym_filtered$AFDW.mg.cm2_sqrt)
qqline(biomass_sym_filtered$AFDW.mg.cm2_sqrt, col = "red")

# log 
biomass_sym_filtered$AFDW.mg.cm2_log <- log(biomass_sym_filtered$AFDW.mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(biomass_sym_filtered$AFDW.mg.cm2_log)  # Check normality
ggplot(biomass_sym_filtered, aes(x = AFDW.mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_sym_filtered$AFDW.mg.cm2_log)
qqline(biomass_sym_filtered$AFDW.mg.cm2_log, col = "red")

# Inverse 
biomass_sym_filtered$AFDW.mg.cm2_inv <- 1 / (biomass_sym_filtered$AFDW.mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(biomass_sym_filtered$AFDW.mg.cm2_inv)
ggplot(biomass_sym_filtered, aes(x = AFDW.mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_sym_filtered$AFDW.mg.cm2_inv)
qqline(biomass_sym_filtered$AFDW.mg.cm2_inv, col = "red")
```

Sqrt is normal.

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
biomass_sym_filtered$Treatment <- as.factor(biomass_sym_filtered$Treatment)
biomass_sym_filtered$Timepoint <- as.factor(biomass_sym_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(AFDW.mg.cm2_sqrt ~ Treatment * Timepoint, data = biomass_sym_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(AFDW.mg.cm2_sqrt ~ Treatment * Timepoint, data = biomass_sym_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

No significant differences 

Run post-hoc test for Treatment and Timepoint
```{r}
TukeyHSD(anova_result, "Treatment", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)

TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
biomass_sym_filtered <- biomass_sym_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Plot untransformed data 
```{r}
# Set Month as factor 
biomass_sym_filtered$Month <- factor(biomass_sym_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

afdw_sym <- ggplot(biomass_sym_filtered, aes(x = Month, y = AFDW.mg.cm2, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted")+
  theme_classic() +
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
  theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(face = "bold", size = 30,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); afdw_sym
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")

ggsave("../output/Physiology/AFDW/AFDW_Sym_Treatment_Timepoint.pdf", afdw_sym, width = 20, height = 15)
ggsave("../output/Physiology/AFDW/AFDW_Sym_Treatment_Timepoint.png", afdw_sym, width = 20, height = 15)
```

Calculate mean, sd and sem of each group for plotting 
```{r}
biomass_sym_summary <- biomass_sym_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_sym = mean(AFDW.mg.cm2, na.rm = TRUE),
    sd_sym = sd(AFDW.mg.cm2, na.rm = TRUE),
    sem_sym = sd_sym / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
biomass_sym_summary$Month <- factor(biomass_sym_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
sym_dot_sem <- ggplot(biomass_sym_summary, aes(x = Month, y = mean_sym, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_sym - sem_sym, 
                    ymax = mean_sym + sem_sym), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  xlab("") +
  ylab("Symbiont Tissue Biomass (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); sym_dot_sem
ggsave("../output/Physiology/AFDW/AFDW_Sym_Treatment_Timepoint_dotwhisker_sem.pdf", sym_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/AFDW/AFDW_Sym_Treatment_Timepoint_dotwhisker_sem.png", sym_dot_sem, width = 20, height = 15)

# sd
sym_dot_sd <- ggplot(biomass_sym_summary, aes(x = Month, y = mean_sym, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_sym - sd_sym, 
                    ymax = mean_sym + sd_sym),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  xlab("") +
  ylab("Symbiont Tissue Biomass (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );sym_dot_sd

ggsave("../output/Physiology/AFDW/AFDW_Sym_Treatment_Timepoint_dotwhisker_sd.pdf", sym_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/AFDW/AFDW_Sym_Treatment_Timepoint_dotwhisker_sd.png", sym_dot_sd, width = 22, height = 12)
```

# Remove February and rerun stats - host

Read in data and subset host 
```{r}
biomass <- read.csv("../output/Physiology/AFDW/AFDW_biomass.csv")

biomass_host <- biomass %>%
  dplyr::filter(Fraction == "Host") %>%
  na.omit()
```

Remove Feb samples 
```{r}
biomass_host <- biomass_host %>%
  filter(!Timepoint == "TP0")
```

Check for outliers 
```{r}
# Look for outliers visually
ggplot(biomass_host, aes(x = Treatment, y = AFDW.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(biomass_host$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_host$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- biomass_host[biomass_host$AFDW.mg.cm2 < lower_bound | biomass_host$AFDW.mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
biomass_host$z_score <- scale(biomass_host$AFDW.mg.cm2)
outliers_z <- biomass_host[abs(biomass_host$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(biomass_host$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_host$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
biomass_host_filtered <- biomass_host[biomass_host$AFDW.mg.cm2 >= lower_bound & biomass_host$AFDW.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(biomass_host)
dim(biomass_host_filtered)
```

3 outliers removed 

Assess normality of data 
```{r}
# Histogram
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(biomass_host_filtered$AFDW.mg.cm2)
qqline(biomass_host_filtered$AFDW.mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(biomass_host_filtered$AFDW.mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
biomass_host_filtered$AFDW.mg.cm2_sqrt <- sqrt(biomass_host_filtered$AFDW.mg.cm2)
shapiro.test(biomass_host_filtered$AFDW.mg.cm2_sqrt)
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_host_filtered$AFDW.mg.cm2_sqrt)
qqline(biomass_host_filtered$AFDW.mg.cm2_sqrt, col = "red")

# log 
biomass_host_filtered$AFDW.mg.cm2_log <- log(biomass_host_filtered$AFDW.mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(biomass_host_filtered$AFDW.mg.cm2_log)  # Check normality
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_host_filtered$AFDW.mg.cm2_log)
qqline(biomass_host_filtered$AFDW.mg.cm2_log, col = "red")

# Inverse 
biomass_host_filtered$AFDW.mg.cm2_inv <- 1 / (biomass_host_filtered$AFDW.mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(biomass_host_filtered$AFDW.mg.cm2_inv)
ggplot(biomass_host_filtered, aes(x = AFDW.mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(biomass_host_filtered$AFDW.mg.cm2_inv)
qqline(biomass_host_filtered$AFDW.mg.cm2_inv, col = "red")
```

Both log and sqrt are normal. Going to stick with sqrt, as this has been the transformation that I have applied above. 

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
biomass_host_filtered$Treatment <- as.factor(biomass_host_filtered$Treatment)
biomass_host_filtered$Timepoint <- as.factor(biomass_host_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(AFDW.mg.cm2_sqrt ~ Treatment * Timepoint, data = biomass_host_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(AFDW.mg.cm2_sqrt ~ Treatment * Timepoint, data = biomass_host_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

No variables are significant.

Run post-hoc test for Treatment and Timepoint
```{r}
TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
TukeyHSD(anova_result, "Treatment", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
biomass_host_filtered <- biomass_host_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Calculate mean, sd and sem of each group for plotting 
```{r}
biomass_host_summary <- biomass_host_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_host = mean(AFDW.mg.cm2, na.rm = TRUE),
    sd_host = sd(AFDW.mg.cm2, na.rm = TRUE),
    sem_host = sd_host / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
biomass_host_summary$Month <- factor(biomass_host_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
host_dot_sem <- ggplot(biomass_host_summary, aes(x = Month, y = mean_host, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_host - sem_host, 
                    ymax = mean_host + sem_host), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); host_dot_sem
ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sem_noFeb.pdf", host_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sem_noFeb.png", host_dot_sem, width = 20, height = 15)

# sd
host_dot_sd <- ggplot(biomass_host_summary, aes(x = Month, y = mean_host, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_host - sd_host, 
                    ymax = mean_host + sd_host),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );host_dot_sd

ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sd_noFeb.pdf", host_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/AFDW/AFDW_Treatment_Timepoint_dotwhisker_sd_noFeb.png", host_dot_sd, width = 22, height = 12)
```