---
title: "Carbohydrates"
author: "jillashey"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is modified from A. Huffmyer's original [script](https://github.com/AHuffmyer/SymbioticIntegration/blob/main/Mcap2021/Scripts/Carbohydrates.Rmd). AH and I ran multiple plates but only plate 3 had Astrangia samples. Using only plate 3, this script analyzes and plots carbohydrate data for the Astrangia 2021 experiment. The carb assays were done on 4/6/22 with AH 

## Load libraries
```{r}
if (!require("dplyr")) install.packages("dplyr")
if (!require("readr")) install.packages("readr")
if (!require("stringr")) install.packages("stringr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("grid")) install.packages("grid")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lattice")) install.packages("lattice")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lsmeans")) install.packages("lsmeans")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("car")) install.packages("car")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("corrr")) install.packages("corrr")
if (!require("GGally")) install.packages("GGally")
if (!require("cowplot")) install.packages("cowplot")
if (!require("multcomp")) install.packages("multcomp")

library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
library(car)
library(kableExtra)
library(ggcorrplot)
library(corrr)
library(GGally)
library(cowplot)
library(multcomp)
library(ggstatsplot)
```

## Load data 
```{r}
carbs_data <- read.csv("../data/Physiology/Carbohydrates/Carb_Data.csv")
carbs_meta <- read.csv("../data/Physiology/Carbohydrates/Carb_Meta.csv")
```

## Separate by run? 
```{r}
# merge data and metadata
carbs_data <- merge(carbs_meta, carbs_data, by = c("Well", "Run"))

run1 <- carbs_data %>% 
  filter(Run == "1")
run2 <- carbs_data %>% 
  filter(Run == "2")
run3 <- carbs_data %>% 
  filter(Run == "3")
run4 <- carbs_data %>% 
  filter(Run == "4")
run5 <- carbs_data %>% 
  filter(Run == "5")

# Average blanks for each run and calculate blank correction
blank1 <- run1 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run1$abs.corr <- run1$X485 - blank1$blk.avg

blank2 <- run2 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run2$abs.corr <- run2$X485 - blank2$blk.avg

blank3 <- run3 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run3$abs.corr <- run3$X485 - blank3$blk.avg

blank4 <- run4 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run4$abs.corr <- run4$X485 - blank4$blk.avg

blank5 <- run5 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run5$abs.corr <- run5$X485 - blank5$blk.avg
```

## Plot standard curve for each run 
```{r}
# run1
standard1 <- run1 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard1, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard1 <- lm(Concentration ~ abs.corr, data = standard1)
summary(lmstandard1)

# run2
standard2 <- run2 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard2, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard2 <- lm(Concentration ~ abs.corr, data = standard2)
summary(lmstandard2)

# run3
standard3 <- run3 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard3, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard3 <- lm(Concentration ~ abs.corr, data = standard3)
summary(lmstandard3)

# run4
standard4 <- run4 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard4, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard4 <- lm(Concentration ~ abs.corr, data = standard4)
summary(lmstandard4)

# run5
standard5 <- run5 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard5, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard5 <- lm(Concentration ~ abs.corr, data = standard5)
summary(lmstandard5)
```

## Calculate concentration values from standard curve 
```{r}
sample1 <- run1 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample1$Concentration <- predict(lmstandard1, newdata = sample1) # using model to get concentration

sample2 <- run2 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample2$Concentration <- predict(lmstandard2, newdata = sample2) # using model to get concentration

sample3 <- run3 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample3$Concentration <- predict(lmstandard3, newdata = sample3) # using model to get concentration

sample4 <- run4 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample4$Concentration <- predict(lmstandard4, newdata = sample4) # using model to get concentration

sample5 <- run5 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample5$Concentration <- predict(lmstandard5, newdata = sample5) # using model to get concentration
```

Bind dfs together and average samples
```{r}
carb_all <- rbind(sample1, sample2, sample3, sample4, sample5)

carb_all <- carb_all %>%
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_concentration = mean(Concentration)) %>%
  filter(mean_concentration > 0)
```

## Read in metadata, surface area, and homogenate volume
```{r}
surface_area <- read.csv("../data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("../data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("../data/MasterFragmentSheet.csv", header = T, na.strings = "") %>% 
  dplyr::filter(Metric == "P/R, molec") %>% 
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge carb data with metadata 
```{r}
all <- carb_all %>% 
  dplyr::rename(PlugID = ID) %>%
  full_join(metadata, by = "PlugID")
```

## Normalize to homogenate vol and surface area 
```{r}
# AH calculation (for larvae): Samples$Carb.mg <- (Samples$Concentration * (Samples$Resuspension_volume/1000) * (1000/Samples$Homo_vol))
# KW calculation (for adults?): samples.carb.meta$Carb.mgcm2 <- (samples.carb.meta$Concentration * samples.carb.meta$Homogenate_Vol.mL * 10)/samples.carb.meta$Surface_Area_cm2 

# Accounting for dilution factor (1000/10) and Normalizing to homogenate volume and surface area
all$Carb.mg.cm2 <- (all$mean_concentration * all$vol_mL * 10)/all$Area_cm ## do I multiple by 100 (since I took 100 ul for the assay) or 10 (because I took 1/10th of the volume from the aliquotted homogenate)

all <- all %>%
  mutate(carb_ug = mean_concentration/vol_mL,
         carb_ug.cm2 = carb_ug / Area_cm,
         carb_mg.cm2 = carb_ug.cm2/1000)
```

Save carb data
```{r}
write.csv(all, "../output/Physiology/Carbohydrates/Carbohydrates.csv")
```

# Stats and plotting 

Read in data
```{r}
carb <- read.csv("../output/Physiology/Carbohydrates/Carbohydrates.csv")

carb <- carb %>%
  mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Check for outliers 
```{r}
# Look for outliers visually
ggplot(carb, aes(x = Treatment, y = Carb.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(carb$Carb.mg.cm2, 0.25)
Q3 <- quantile(carb$Carb.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- carb[carb$Carb.mg.cm2 < lower_bound | carb$Carb.mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
carb$z_score <- scale(carb$Carb.mg.cm2)
outliers_z <- carb[abs(carb$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(carb$Carb.mg.cm2, 0.25)
Q3 <- quantile(carb$Carb.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
carb_exp_filtered <- carb[carb$Carb.mg.cm2 >= lower_bound & carb$Carb.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(carb)
dim(carb_exp_filtered)
```

6 outliers removed. 

Assess normality of data 
```{r}
# Histogram
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(carb_exp_filtered$Carb.mg.cm2)
qqline(carb_exp_filtered$Carb.mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(carb_exp_filtered$Carb.mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
carb_exp_filtered$Carb.mg.cm2_sqrt <- sqrt(carb_exp_filtered$Carb.mg.cm2)
shapiro.test(carb_exp_filtered$Carb.mg.cm2_sqrt)
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(carb_exp_filtered$Carb.mg.cm2_sqrt)
qqline(carb_exp_filtered$Carb.mg.cm2_sqrt, col = "red")

# log 
carb_exp_filtered$Carb.mg.cm2_log <- log(carb_exp_filtered$Carb.mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(carb_exp_filtered$Carb.mg.cm2_log)  # Check normality
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(carb_exp_filtered$Carb.mg.cm2_log)
qqline(carb_exp_filtered$Carb.mg.cm2_log, col = "red")

# Inverse 
carb_exp_filtered$Carb.mg.cm2_inv <- 1 / (carb_exp_filtered$Carb.mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(carb_exp_filtered$Carb.mg.cm2_inv)
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(carb_exp_filtered$Carb.mg.cm2_inv)
qqline(carb_exp_filtered$Carb.mg.cm2_inv, col = "red")
```

Sqrt transformation produces normality. Remove NAs produced from negative values 
```{r}
carb_exp_filtered <- carb_exp_filtered %>%
  filter(Carb.mg.cm2_sqrt != "NaN")
```

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
carb_exp_filtered$Treatment <- as.factor(carb_exp_filtered$Treatment)
carb_exp_filtered$Timepoint <- as.factor(carb_exp_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(Carb.mg.cm2_sqrt ~ Treatment * Timepoint, data = carb_exp_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(Carb.mg.cm2_sqrt ~ Treatment * Timepoint, data = carb_exp_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

Treatment is significant but Timepoint and interaction is not. This contrasts with our other phys data, which is primarily different by time point. 

Run post-hoc test for Treatment and Timepoint
```{r}
TukeyHSD(anova_result, "Treatment", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)

TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
carb_exp_filtered <- carb_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Plot untransformed data 
```{r}
# Set Month as factor 
carb_exp_filtered$Month <- factor(carb_exp_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

carb <- ggplot(carb_exp_filtered, aes(x = Month, y = Carb.mg.cm2, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted")+
  theme_classic() +
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
  theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(face = "bold", size = 30,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); carb
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")

ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint.pdf", carb, width = 20, height = 15)
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint.png", carb, width = 20, height = 15)
```

Calculate mean, sd and sem of each group for plotting 
```{r}
carb_summary <- carb_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_carb = mean(Carb.mg.cm2, na.rm = TRUE),
    sd_carb = sd(Carb.mg.cm2, na.rm = TRUE),
    sem_carb = sd_carb / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
carb_summary$Month <- factor(carb_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
carb_dot_sem <- ggplot(carb_summary, aes(x = Month, y = mean_carb, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_carb - sem_carb, 
                    ymax = mean_carb + sem_carb), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); carb_dot_sem
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sem.pdf", carb_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sem.png", carb_dot_sem, width = 20, height = 15)

# sd
carb_dot_sd <- ggplot(carb_summary, aes(x = Month, y = mean_carb, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_carb - sd_carb, 
                    ymax = mean_carb + sd_carb),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );carb_dot_sd

ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sd.pdf", carb_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sd.png", carb_dot_sd, width = 22, height = 12)
```

# Remove February and rerun stats

Read in data
```{r}
carb <- read.csv("../output/Physiology/Carbohydrates/Carbohydrates.csv")
```

Remove field and Feb samples 
```{r}
carb_exp <- carb %>%
  filter(!Treatment == "Field") %>%
  filter(!Timepoint == "TP0")
```

Check for outliers 
```{r}
# Look for outliers visually
ggplot(carb_exp, aes(x = Treatment, y = Carb.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(carb_exp$Carb.mg.cm2, 0.25)
Q3 <- quantile(carb_exp$Carb.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- carb_exp[carb_exp$Carb.mg.cm2 < lower_bound | carb_exp$Carb.mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
carb_exp$z_score <- scale(carb_exp$Carb.mg.cm2)
outliers_z <- carb_exp[abs(carb_exp$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(carb_exp$Carb.mg.cm2, 0.25)
Q3 <- quantile(carb_exp$Carb.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
carb_exp_filtered <- carb_exp[carb_exp$Carb.mg.cm2 >= lower_bound & carb_exp$Carb.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(carb_exp)
dim(carb_exp_filtered)
```

7 outliers removed 

Assess normality of data 
```{r}
# Histogram
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(carb_exp_filtered$Carb.mg.cm2)
qqline(carb_exp_filtered$Carb.mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(carb_exp_filtered$Carb.mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
carb_exp_filtered$Carb.mg.cm2_sqrt <- sqrt(carb_exp_filtered$Carb.mg.cm2)
shapiro.test(carb_exp_filtered$Carb.mg.cm2_sqrt)
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(carb_exp_filtered$Carb.mg.cm2_sqrt)
qqline(carb_exp_filtered$Carb.mg.cm2_sqrt, col = "red")

# log 
carb_exp_filtered$Carb.mg.cm2_log <- log(carb_exp_filtered$Carb.mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(carb_exp_filtered$Carb.mg.cm2_log)  # Check normality
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(carb_exp_filtered$Carb.mg.cm2_log)
qqline(carb_exp_filtered$Carb.mg.cm2_log, col = "red")

# Inverse 
carb_exp_filtered$Carb.mg.cm2_inv <- 1 / (carb_exp_filtered$Carb.mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(carb_exp_filtered$Carb.mg.cm2_inv)
ggplot(carb_exp_filtered, aes(x = Carb.mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(carb_exp_filtered$Carb.mg.cm2_inv)
qqline(carb_exp_filtered$Carb.mg.cm2_inv, col = "red")
```

Sqrt transformation produces normality. Remove NAs produced from negative values 
```{r}
carb_exp_filtered <- carb_exp_filtered %>%
  filter(Carb.mg.cm2_sqrt != "NaN")
```

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
carb_exp_filtered$Treatment <- as.factor(carb_exp_filtered$Treatment)
carb_exp_filtered$Timepoint <- as.factor(carb_exp_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(Carb.mg.cm2_sqrt ~ Treatment * Timepoint, data = carb_exp_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(Carb.mg.cm2_sqrt ~ Treatment * Timepoint, data = carb_exp_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

No variables are significant.

Run post-hoc test for Treatment and Timepoint
```{r}
TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
TukeyHSD(anova_result, "Treatment", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
carb_exp_filtered <- carb_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Calculate mean, sd and sem of each group for plotting 
```{r}
carb_summary <- carb_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_carb = mean(Carb.mg.cm2, na.rm = TRUE),
    sd_carb = sd(Carb.mg.cm2, na.rm = TRUE),
    sem_carb = sd_carb / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
carb_summary$Month <- factor(carb_summary$Month, levels = c("March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
carb_dot_sem <- ggplot(carb_summary, aes(x = Month, y = mean_carb, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_carb - sem_carb, 
                    ymax = mean_carb + sem_carb), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); carb_dot_sem
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sem_noFeb.pdf", carb_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sem_noFeb.png", carb_dot_sem, width = 20, height = 15)

# sd
carb_dot_sd <- ggplot(carb_summary, aes(x = Month, y = mean_carb, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_carb - sd_carb, 
                    ymax = mean_carb + sd_carb),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );carb_dot_sd

ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sd_noFeb.pdf", carb_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_Treatment_Timepoint_dotwhisker_sd_noFeb.png", carb_dot_sd, width = 22, height = 12)
```

