---
title: "Carbohydrates"
author: "jillashey"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is modified from A. Huffmyer's original [script](https://github.com/AHuffmyer/SymbioticIntegration/blob/main/Mcap2021/Scripts/Carbohydrates.Rmd). AH and I ran multiple plates but only plate 3 had Astrangia samples. Using only plate 3, this script analyzes and plots carbohydrate data for the Astrangia 2021 experiment. The carb assays were done on 4/6/22 with AH 

## Load libraries
```{r}
if (!require("dplyr")) install.packages("dplyr")
if (!require("readr")) install.packages("readr")
if (!require("stringr")) install.packages("stringr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("grid")) install.packages("grid")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lattice")) install.packages("lattice")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lsmeans")) install.packages("lsmeans")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("car")) install.packages("car")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("corrr")) install.packages("corrr")
if (!require("GGally")) install.packages("GGally")
if (!require("cowplot")) install.packages("cowplot")
if (!require("multcomp")) install.packages("multcomp")

library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
library(car)
library(kableExtra)
library(ggcorrplot)
library(corrr)
library(GGally)
library(cowplot)
library(multcomp)
```

## Load data 
```{r}
data <- read.csv("data/Physiology/Carbohydrates/20220406_Carbohydrates_Data.csv")
Meta <- read.csv("data/Physiology/Carbohydrates/20220406_Carbohydrates_Meta.csv")
```

## Calculate standard curve
Merge files and rename cols
```{r}
data.1 <- merge(Meta, data, by = c("Well", "Run"))

# Blank correction for each run separately
Blank <- data.1 %>% 
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
data.1$abs.corr <- data.1$X485 - Blank$blk.avg
```

Plot standard curve 
```{r}
Standard <- data.1 %>% 
  filter(Sample.Type == "Standard")
Standard.plot <- ggplot(data = Standard, aes(x=Concentration, y=abs.corr))+
  #facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot
```

Extract standard curve. 
```{r}
lmstandard <- lm (Concentration ~ abs.corr, data = Standard)
lmsummary <- summary(lmstandard) 
```

## Calculations (initial curve)

Obtain concentration values for samples from standard curve.  
```{r}
Samples <- data.1 %>% #subsetting Samples
  filter(Sample.Type == "Sample")

Samples$Concentration <- predict(lmstandard, newdata = Samples) #using model to get concentration
```

Accounting for dilution factor (1000/10) and normalizing to homogenate volume 
```{r}
Samples$Carb.uL <- (Samples$Concentration * Samples$Homo_vol * 10)
```

Remove AH samples 
```{r}
Samples <- Samples %>% 
  filter(!Sample.ID == "X")

Samples <- Samples %>% 
  filter(!Sample.ID == "W")
```


Plot Astrangia
```{r}
Carb.Plot.Ast <- Samples%>%
  ggplot(aes(x=Treatment, y=Carb.uL, fill = Treatment)) +
  #facet_grid(~Sample.ID)+
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
#  stat_summary(fun=mean, aes(group=Fraction, color = Fraction), position = position_dodge(width = 0.5))  + 
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  #  ylim(0,0.5) +
  xlab("Temperature") + ylab(expression("Total Carbohydrate " (mg~uL))) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, size=0.75), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  #  theme(axis.text = element_text(size = 30, color = "black"),
  #        axis.title = element_text(size = 36, color = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Carb.Plot.Ast
ggsave("output/Physiology/Carbohydrates/Carbohydrates.pdf", Carb.Plot.Ast)
```







20230420
I now have processed most of my Astrangia 2021 samples for the host fraction

## Load data 
```{r}
carbs_data <- read.csv("../data/Physiology/Carbohydrates/Carb_Data.csv")
carbs_meta <- read.csv("../data/Physiology/Carbohydrates/Carb_Meta.csv")
```

## Separate by run? 
```{r}
# merge data and metadata
carbs_data <- merge(carbs_meta, carbs_data, by = c("Well", "Run"))

run1 <- carbs_data %>% 
  filter(Run == "1")
run2 <- carbs_data %>% 
  filter(Run == "2")
run3 <- carbs_data %>% 
  filter(Run == "3")
run4 <- carbs_data %>% 
  filter(Run == "4")
run5 <- carbs_data %>% 
  filter(Run == "5")

# Average blanks for each run and calculate blank correction
blank1 <- run1 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run1$abs.corr <- run1$X485 - blank1$blk.avg

blank2 <- run2 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run2$abs.corr <- run2$X485 - blank2$blk.avg

blank3 <- run3 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run3$abs.corr <- run3$X485 - blank3$blk.avg

blank4 <- run4 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run4$abs.corr <- run4$X485 - blank4$blk.avg

blank5 <- run5 %>%
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X485))
run5$abs.corr <- run5$X485 - blank5$blk.avg
```

## Plot standard curve for each run 
```{r}
# run1
standard1 <- run1 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard1, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard1 <- lm(Concentration ~ abs.corr, data = standard1)
summary(lmstandard1)

# run2
standard2 <- run2 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard2, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard2 <- lm(Concentration ~ abs.corr, data = standard2)
summary(lmstandard2)

# run3
standard3 <- run3 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard3, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard3 <- lm(Concentration ~ abs.corr, data = standard3)
summary(lmstandard3)

# run4
standard4 <- run4 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard4, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard4 <- lm(Concentration ~ abs.corr, data = standard4)
summary(lmstandard4)

# run5
standard5 <- run5 %>%
  filter(Sample.Type =="Standard")
  
ggplot(data = standard5, aes(x=Concentration, y=abs.corr))+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lmstandard5 <- lm(Concentration ~ abs.corr, data = standard5)
summary(lmstandard5)
```

## Calculate concentration values from standard curve 
```{r}
sample1 <- run1 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample1$Concentration <- predict(lmstandard1, newdata = sample1) # using model to get concentration

sample2 <- run2 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample2$Concentration <- predict(lmstandard2, newdata = sample2) # using model to get concentration

sample3 <- run3 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample3$Concentration <- predict(lmstandard3, newdata = sample3) # using model to get concentration

sample4 <- run4 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample4$Concentration <- predict(lmstandard4, newdata = sample4) # using model to get concentration

sample5 <- run5 %>%
  filter(Sample.Type == "Sample") # subsetting samples 
sample5$Concentration <- predict(lmstandard5, newdata = sample5) # using model to get concentration
```

Bind dfs together and average samples
```{r}
carb_all <- rbind(sample1, sample2, sample3, sample4, sample5)

carb_all <- carb_all %>%
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_concentration = mean(Concentration))
```

## Read in metadata, surface area, and homogenate volume
```{r}
surface_area <- read.csv("../data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("../data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("../data/MasterFragmentSheet.csv", header = T, na.strings = "") %>% 
  dplyr::filter(Metric == "P/R, molec") %>% 
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge carb data with metadata 
```{r}
carb_all <- carb_all %>% 
  dplyr::rename(PlugID = ID)

all <- merge(metadata, carb_all, by = "PlugID")
```

## Normalize to homogenate vol and surface area 
```{r}
# AH calculation (for larvae): Samples$Carb.mg <- (Samples$Concentration * (Samples$Resuspension_volume/1000) * (1000/Samples$Homo_vol))
# KW calculation (for adults?): samples.carb.meta$Carb.mgcm2 <- (samples.carb.meta$Concentration * samples.carb.meta$Homogenate_Vol.mL * 10)/samples.carb.meta$Surface_Area_cm2 

# Accounting for dilution factor (1000/10) and Normalizing to homogenate volume and surface area
all$Carb.mg.cm2 <- (all$mean_concentration * all$vol_mL * 10)/all$Area_cm ## do I multiple by 100 (since I took 100 ul for the assay) or 10 (because I took 1/10th of the volume from the aliquotted homogenate)
```

Save carb data
```{r}
write.csv(all, "../output/Physiology/Carbohydrates/Carbohydrates.csv")
```



## Plotting and stats
Read in data
```{r}
carb <- read.csv("../output/Physiology/Carbohydrates/Carbohydrates.csv")
```

Assign Month to TPs
```{r}
carb <- carb %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "Febuary", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Remove field data and change Treatment for TP0 to be "Acclimation"
```{r}
carb <- carb %>%
#   filter(!Timepoint == "TP0") %>%
   filter(!Treatment == "Field") %>%
   filter(Carb.mg.cm2 <= 15)
   mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", Treatment))
```

Set levels for plotting by month 
```{r}
carb$Month <- factor(carb$Month, levels = c("Febuary", "March", "Early April", "Late April", "May", "June", "July", "August"))
```

Plot 
```{r}
plot <- ggplot(carb, aes(x = Month, y = Carb.mg.cm2, colour = Treatment)) +
  geom_boxplot(aes(colour=Treatment), lwd=2) +
  #geom_point(aes(fill=Treatment, group=Treatment), colour="black", pch = 21, size=4, position = position_jitterdodge(0.3), alpha=0.9) + 
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) +
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  #hjust(x = 0) +
  #facet_wrap(~Assay, ncol = 2) +
  theme_classic() +
    theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); plot 

ggsave("../output/Physiology/Carbohydrates/Carbohydrates_20240521.pdf", plot, width = 15, height = 10)
ggsave("../output/Physiology/Carbohydrates/Carbohydrates_20240521.png", plot, width = 15, height = 10)
```

Summarize data
```{r}
## Summarize data w/o field 
mean.carb_exp <- summarySE(carb, measurevar = "Carb.mg.cm2", groupvars = c("Treatment", "Timepoint"))
mean.carb_site_exp <- summarySE(carb, measurevar = "Carb.mg.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
```

Looking at this plot, I see that not all of the treatments/timepoints have n=8. I need to look through my data and notebooks so that I can make sure I have all the data that I need and that I have processed everything. 

Going to do stats just for Temperate coral conference 
```{r}
model <- aov(Carb.mg.cm2 ~ Treatment*Timepoint, data = carb)
summary(model)

#                     Df Sum Sq Mean Sq F value  Pr(>F)   
# Treatment             2   30.7  15.368   3.328 0.03962 * 
# Timepoint             6   84.0  13.994   3.030 0.00892 **
# Treatment:Timepoint   6   42.3   7.051   1.527 0.17632   
# Residuals           107  494.1   4.618                   
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

## Check ANOVA assumptions 
# ANOVA assumes data is normally distributed and variance across groups is homogenous 
# 1. Check homogeneity of variance assumption
# a. Residuals vs fits plot
plot(model, 1)

# b. Levene's test 
leveneTest(residuals(model) ~ Timepoint*Treatment, data = carb)
## p-value is greater than 0.05, so variance is homogeneous 

# 2. Check normality assumption 
# a. Normality plot of residuals 
plot(model, 2) 
hist(residuals(model))

#b. Shapiro-Wilk test 
aov_residuals <- residuals(object = model)
shapiro.test(x = aov_residuals)
# pvalue is less than 0.05, normality is violated

## Compute summary stats 
model.tables(model, type = "means", se = T)

## Multiple pairwise-comparisons between the means of groups 
# Tukey HSD 
TukeyHSD(model, which = "Timepoint")
## TP2 differs from TP7

TukeyHSD(model, which = "Treatment")
## Heat differs from amb (but just barely w/ pvalue of 0.0504046)
```

