---
title: "Chlorophyll-a"
author: "jillashey"
date: "2023-02-23"
output: html_document
---

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")
if (!require("dplyr")) install.packages("dplyr")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("formattable")) install.packages("formattable")
if (!require("data.table")) install.packages("data.table")
if (!require("htmltools")) install.packages("htmltools")
if (!require("webshot")) install.packages("webshot")

# load packages
library(plyr)
library(dplyr)
library(tidyverse)
library(plotrix)
library(Rmisc)
library(ggplot2)
library(gridExtra)
library(formattable)
library(data.table)
library(htmltools)
library(webshot)
library(lme4)
library(MuMIn)
library(cowplot)
library(corrplot)
library(lmerTest)
```

## Import data 
```{r}
## There is definitely a better way to do this

# Plate 1
p1_data <- read.csv("data/Physiology/Chlorophyll/20230118_plate1_data.csv")
p1_platemap <- read.csv("data/Physiology/Chlorophyll/20230118_plate1_platemap.csv")
p1 <- left_join(p1_data, p1_platemap)

# Plate 2
p2_data <- read.csv("data/Physiology/Chlorophyll/20230126_plate2_data.csv")
p2_platemap <- read.csv("data/Physiology/Chlorophyll/20230126_plate2_platemap.csv")
p2 <- left_join(p2_data, p2_platemap, by = "Well")

# Plate 3
p3_data <- read.csv("data/Physiology/Chlorophyll/20230210_plate3_data.csv")
p3_platemap <- read.csv("data/Physiology/Chlorophyll/20230210_plate3_platemap.csv")
p3 <- left_join(p3_data, p3_platemap)

# Plate 4
p4_data <- read.csv("data/Physiology/Chlorophyll/20230214_plate4_data.csv")
p4_platemap <- read.csv("data/Physiology/Chlorophyll/20230214_plate4_platemap.csv")
p4 <- left_join(p4_data, p4_platemap)

# Plate 5
p5_data <- read.csv("data/Physiology/Chlorophyll/20230216_plate5_data.csv")
p5_platemap <- read.csv("data/Physiology/Chlorophyll/20230216_plate5_platemap.csv")
p5 <- left_join(p5_data, p5_platemap)
```

## Calculate chlorophyll concentrations for each plate 
```{r}
# Group by run and sample ID
p1 <- p1 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)
  
p2 <- p2 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)

p3 <- p3 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)

p4 <- p4 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)

p5 <- p5 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)
```

## Bind all dfs together 
```{r}
all <- rbind(p1, p2, p3, p4, p5)

# Remove blanks
all <- all %>% 
  filter(!ID == "Blank") %>%
  dplyr::rename(PlugID = ID)
```

## Read in surface area, homogenate volume and metadata 
```{r}
surface_area <- read.csv("data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("data/MasterFragmentSheet.csv", header = T, na.strings = "") %>%
  dplyr::filter(Metric == "P/R, molec") %>%
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

# Merge dataframes
metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
#metadata$vol_mL <- gsub("NA", "", metadata$vol_mL)
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge metadata and chl data and normalize data 
```{r}
chl <- left_join(metadata, all) %>%
       mutate(chla.ug.cm2 = (chla.ug.ml * vol_mL) / Area_cm, # I need to calculate surface area squared?
             chlc2.ug.cm2 = (chlc2.ug.ml * vol_mL) / Area_cm) %>%
       na.omit() ## omitting NAs for now 
```

## Summarize 
```{r}
mean.chla <- summarySE(chl, measurevar = "chla.ug.cm2", groupvars = c("Treatment", "Timepoint"))
mean.chlc2 <- summarySE(chl, measurevar = "chlc2.ug.cm2", groupvars = c("Treatment", "Timepoint"))

mean.chla_site <- summarySE(chl, measurevar = "chla.ug.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
mean.chlc2_site <- summarySE(chl, measurevar = "chlc2.ug.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
# I did not equally sample across site, which was a mistake on my part. I need to figure out how to present the data with site 
```

## Plot 
```{r}
# Remove field data for now 
chl_exp <- chl %>%
  filter(!Treatment == "Field")

# plot by exp only -- chl a
plot_chla_trt <- ggplot(chl_exp, aes(x = Timepoint, y = chla.ug.cm2, fill = Treatment)) +
  geom_boxplot(lwd = 1, color = "black") +
  #geom_point(aes(color = Treatment), size = 2) +
  #geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_fill_manual(values = c("blue", "red")) +
  ylab("Chlorophyll-a (ug/cm²)") +
  xlab("") +
  theme_classic() +
  theme(legend.title = element_text(size = 28),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))) 
plot_chla_trt

ggsave("output/Physiology/Chlorophyll/Chla_Treatment.pdf", plot_chla_trt, width = 50, height = 25, units = "cm")
ggsave("output/Physiology/Chlorophyll/Chla_Treatment.png", plot_chla_trt, width = 50, height = 25, units = "cm")

# plot by exp only -- chl c2
plot_chlc2_trt <- ggplot(chl_exp, aes(x = Timepoint, y = chlc2.ug.cm2, fill = Treatment)) +
  geom_boxplot(lwd = 1, color = "black") +
  #geom_point(aes(color = Treatment), size = 2) +
  #geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_fill_manual(values = c("blue", "red")) +
  ylab("Chlorophyll-a (ug/cm²)") +
  xlab("") +
  theme_classic() +
  theme(legend.title = element_text(size = 28),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))) 
plot_chlc2_trt

ggsave("output/Physiology/Chlorophyll/Chlc2_Treatment.pdf", plot_chlc2_trt, width = 50, height = 25, units = "cm")
ggsave("output/Physiology/Chlorophyll/Chlc2_Treatment.png", plot_chlc2_trt, width = 50, height = 25, units = "cm")
```

## Stats 
## Test for differences 
For now (as of 20230420), I'm using code based on this [tutorial](http://www.sthda.com/english/wiki/two-way-anova-test-in-r). I need to do some more investigations on stats stuff 
```{r}
## Compute two-way ANOVA test 
model <- aov(log(chla.ug.cm2) ~ Treatment*Timepoint, data = chl_exp) 
summary(model)

## Compute summary stats 
model.tables(model, type = "means", se = T)
```































## Read in dfs and merge into one large df
```{r}
## platemaps 
df.platemaps <- 
  list.files(path = "data/Physiology/Chlorophyll-a", pattern = "platemap.csv", full.names = TRUE) %>% 
  #setnames(.) %>% 
  map_dfr(read.csv, .id = "file.ID") %>% # join all files together in one df by file ID (ie day/date of run; plate 1, 2, etc)
  group_by(file.ID)

df.platemaps$file.ID <- gsub("data/Physiology/Chlorophyll-a", "", df.platemaps$file.ID)
df.platemaps$file.ID <- gsub("_platemap.csv", "", df.platemaps$file.ID)

## data
df.data <- 
  list.files(path = "data/Physiology/Chlorophyll-a", pattern = "data.csv", full.names = TRUE) %>% 
  map_dfr(read.csv, .id = "file.ID") %>% # join all files together in one df by file ID (ie day/date of run; plate 1, 2, etc)
  group_by(file.ID)

df.data$file.ID <-gsub("data/Physiology/Chlorophyll-a", "", df.data$file.ID)
df.data$file.ID <- gsub("_data.csv", "", df.data$file.ID)

# cut off last empty rows in df.data 
df.data <- df.data[1:480,]

## Merge platemap and data for each plate 
df <- left_join(df.data, df.platemaps)
#str(df)

# Remove unneeded columns 
df <- subset(df, select = -c(X, X.1))
```

## Calculate chlorophyll concentrations 
```{r}
# Avg all technical replicates for each sample duplicates per plate 
df <- df %>%
  filter(!is.na(ID))

# Group by run and sample ID
df.1 <- df %>% 
  dplyr::group_by(file.ID, ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750))

# Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
df.1 <- df.1 %>% 
  dplyr::mutate(adj630 = mean_A630 - mean_A750,
                adj663 = mean_A663 - mean_A750)

# Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
## units are ug/mL
df.1 <- df.1 %>% 
  dplyr::mutate(chla.ug.ml = (11.43 * adj663) - (0.64 * adj630),
                chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663))

# path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
df.1 <- df.1 %>%
    dplyr::mutate(chla.ug.ml = chla.ug.ml*0.584,
        chlc2.ug.ml = chlc2.ug.ml*0.584)
```

## Standardize to surface area and homogenate volume 
```{r}
# Load surface area 
sa <- read.csv("data/Physiology/PR/PR_SurfaceArea.csv")%>% 
  dplyr::select(ID, Site, Treatment, Timepoint, Area_cm)
str(sa)

## join chl and surface area dfs 
df.chl <- left_join(sa, df.1)

# Load homogenate volume 
homog.vol <- read.csv("data/Physiology/homogenate_volume.csv") %>% 
  select(PlugID, vol_mL) %>% 
  rename_with(.cols = 1, ~ "ID")
str(homog.vol)

## join chl and homogenate dfs
df.chl <- left_join(df.chl, homog.vol)


# Account for homogenate volume and normalize to surface area 
df.chl <- df.chl %>%
    mutate(chla.ug.cm2 = (chla.ug.ml * vol_mL) / Area_cm, # I need to calculate surface area squared?
         chlc2.ug.cm2 = (chlc2.ug.ml * vol_mL) / Area_cm) # I need to calculate surface area squared?
         #chla.ug.cells = (chla.ug.ml * vol_mL) / cells, # this is for cell density 
         #chlc2.ug.cells = (chlc2.ug.ml * vol_mL) / cells) # this is for cell density 

# Remove blanks and NAs
df.chl <- filter(df.chl, !ID %in% c("NA", "BK"))

# Save chlorophyll a 
df.chl %>% 
  select(ID, Site, Treatment, Timepoint, chla.ug.ml, chlc2.ug.ml, chla.ug.cm2, chlc2.ug.cm2) %>%
  write.csv(file = "Output/Physiology/chlorophyll.csv")
```

The df df.chl has timepoint and treatment in it. I'm going to drop the NAs in the 'chlc2.ug.cm2' column then plot by treatment and timepoint 
```{r}
# Remove NAs in chlc2.ug.cm2 column 
df.chl <- df.chl %>% 
  drop_na(chlc2.ug.cm2)
```

## Quality control!
```{r}
## Check range
# chl a
range(df.chl$chla.ug.ml)
range(df.chl$chla.ug.cm2)

# chl c
range(df.chl$chlc2.ug.ml)
range(df.chl$chlc2.ug.cm2)

# filter 
dplyr::filter(df.chl, chlc2.ug.cm2 > 1.5) # none higher than 1.5
dplyr::filter(df.chl, chla.ug.cm2 < 0) # output should be zero since negative values are not possible. 
dplyr::filter(df.chl, chlc2.ug.cm2 < 0) # output should be zero since negative values are not possible. 
# remove negative values from chlc2.ug.cm2
df.chl <- df.chl[df.chl$chlc2.ug.cm2 >= 0,]
```

## Calculate mean values for each TP and treatment group 
```{r}
df.chl <- df.chl %>% distinct()

# select certain cols and list chla and chlc2 values by sample 
df.chl_select <- df.chl %>%
  select(ID, Site, Treatment, Timepoint, Area_cm, vol_mL, chla.ug.cm2, chlc2.ug.cm2) %>% 
  gather("Measurement", "Value", 7:8) %>%
  distinct()

# Calculate means by Site, Treatment, and Timepoint 
chl_means <- summarySE(df.chl_select, measurevar = "Value", groupvars = c("Site", "Treatment", "Timepoint", "Measurement")) %>% 
  filter(N > 2)

# Calculate means by Treatment and Timepoint 
chl_means_NoSite <- summarySE(df.chl_select, measurevar = "Value", groupvars = c("Treatment", "Timepoint", "Measurement")) %>% 
  filter(N > 2)
```

## Separate df into chla and chlc2
```{r}
chla_means <- chl_means %>%
  filter(Measurement=="chla.ug.cm2")

chlc2_means <- chl_means %>%
  filter(Measurement=="chlc2.ug.cm2")

## no site 
chla_means_NoSite <- chl_means_NoSite %>% 
    filter(Measurement=="chla.ug.cm2")

chlc2_means_NoSite <- chl_means_NoSite %>% 
  filter(Measurement=="chlc2.ug.cm2")
```


## Plotting timeseries with mean table 
```{r}
cols <- c("blue", "green", "red")
 
# chla cm2 plot 
chla_means_plot <- ggplot(chla_means, aes(x=Timepoint, y=Value, color=Treatment, shape=Site)) +
  geom_point() +
  geom_errorbar(aes(ymin = Value - se, ymax = Value + se, width = 0.2)) +
  scale_color_manual(values = cols) + 
  xlab("") +
  ylab(expression("Chlorophyll a" ~(ug ~cm^-2))) +
  #theme(legend.position = "none")
  theme_classic() +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 20, color = "black", face = "bold.italic"), # Change axis line
  axis.line = element_line(colour = "black")) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 13)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.title = element_text(size = 15))
chla_means_plot

# chlc2 cm2 plot 
chlc2_means_plot <- ggplot(chlc2_means, aes(x=Timepoint, y=Value, color=Treatment, shape=Site)) +
  geom_point() +
  geom_errorbar(aes(ymin = Value - se, ymax = Value + se, width = 0.2)) +
  scale_color_manual(values = cols) + 
  xlab("") +
  ylab(expression("Chlorophyll c2" ~(ug ~cm^-2))) +
  #theme(legend.position = "none")
  theme_classic() +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 20, color = "black", face = "bold.italic"), # Change axis line
  axis.line = element_line(colour = "black")) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 13)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.title = element_text(size = 15))
chlc2_means_plot





### PLOTS WITH NO SITE VARIABLE 
# chla cm2 plot 
chla_means_NoSite_plot <- ggplot(chla_means_NoSite, aes(x=Timepoint, y=Value, color=Treatment)) +
  geom_point() +
  geom_errorbar(aes(ymin = Value - se, ymax = Value + se, width = 0.2)) +
  scale_color_manual(values = cols) + 
  xlab("") +
  ylab(expression("Chlorophyll a" ~(ug ~cm^-2))) +
  #theme(legend.position = "none")
  theme_classic() +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 20, color = "black", face = "bold.italic"), # Change axis line
  axis.line = element_line(colour = "black")) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 13)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.title = element_text(size = 15))
chla_means_NoSite_plot

# chlc2 cm2 plot 
chlc2_means_NoSite_plot <- ggplot(chlc2_means_NoSite, aes(x=Timepoint, y=Value, color=Treatment)) +
  geom_point() +
  geom_errorbar(aes(ymin = Value - se, ymax = Value + se, width = 0.2)) +
  scale_color_manual(values = cols) + 
  xlab("") +
  ylab(expression("Chlorophyll c2" ~(ug ~cm^-2))) +
  #theme(legend.position = "none")
  theme_classic() +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 20, color = "black", face = "bold.italic"), # Change axis line
  axis.line = element_line(colour = "black")) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 13)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.title = element_text(size = 15))
chlc2_means_NoSite_plot
```

## Stats
Raw data distribution view 
```{r}
hist(chla_means$Value)
hist(chlc2_means$Value)

hist(df.chl$chla.ug.cm2)
hist(df.chl$chlc2.ug.cm2)
```

### Linear mixed models 
```{r}
# separate into chla and chlc2 dfs
chla <- df.chl_select %>%
  filter(Measurement=="chla.ug.cm2")

chlc2 <- df.chl_select %>%
  filter(Measurement=="chlc2.ug.cm2")

## Chla per cm2
chl_LMER_a <- lmer(log(Value) ~ Timepoint*Treatment + (1|Timepoint:Treatment), na.action = na.omit, data = chla)
```








