---
title: "Chlorophyll-a"
author: "jillashey"
date: "2023-02-23"
output: html_document
---

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")
if (!require("dplyr")) install.packages("dplyr")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("formattable")) install.packages("formattable")
if (!require("data.table")) install.packages("data.table")
if (!require("htmltools")) install.packages("htmltools")
if (!require("webshot")) install.packages("webshot")

# load packages
library(plyr)
library(dplyr)
library(tidyverse)
library(plotrix)
library(Rmisc)
library(ggplot2)
library(gridExtra)
library(formattable)
library(data.table)
library(htmltools)
library(webshot)
library(lme4)
library(MuMIn)
library(cowplot)
library(corrplot)
library(lmerTest)
library(car)
library(emmeans)
library(ggtext)
```

## Import data 
```{r}
## There is definitely a better way to do this

# Plate 1
p1_data <- read.csv("../data/Physiology/Chlorophyll/20230118_plate1_data.csv")
p1_platemap <- read.csv("../data/Physiology/Chlorophyll/20230118_plate1_platemap.csv")
p1 <- left_join(p1_data, p1_platemap)

# Plate 2
p2_data <- read.csv("../data/Physiology/Chlorophyll/20230126_plate2_data.csv")
p2_platemap <- read.csv("../data/Physiology/Chlorophyll/20230126_plate2_platemap.csv")
p2 <- left_join(p2_data, p2_platemap, by = "Well")

# Plate 3
p3_data <- read.csv("../data/Physiology/Chlorophyll/20230210_plate3_data.csv")
p3_platemap <- read.csv("../data/Physiology/Chlorophyll/20230210_plate3_platemap.csv")
p3 <- left_join(p3_data, p3_platemap)

# Plate 4
p4_data <- read.csv("../data/Physiology/Chlorophyll/20230214_plate4_data.csv")
p4_platemap <- read.csv("../data/Physiology/Chlorophyll/20230214_plate4_platemap.csv")
p4 <- left_join(p4_data, p4_platemap)

# Plate 5
p5_data <- read.csv("../data/Physiology/Chlorophyll/20230216_plate5_data.csv")
p5_platemap <- read.csv("../data/Physiology/Chlorophyll/20230216_plate5_platemap.csv")
p5 <- left_join(p5_data, p5_platemap)
```

## Calculate chlorophyll concentrations for each plate 
```{r}
# Group by run and sample ID
p1 <- p1 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)
  
p2 <- p2 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)

p3 <- p3 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)

p4 <- p4 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)

p5 <- p5 %>% 
  dplyr::group_by(ID) %>%
  dplyr::summarise(mean_A630 = mean(X630), # calculate means 
            mean_A663 = mean(X663), 
            mean_A750 = mean(X750)) %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750, # # Get the acetone blank 750 absorbance for each file (ie plate) and subtract from 630 and 663 values for each sample 
         adj663 = mean_A663 - mean_A750,
         chla.ug.ml = (11.43 * adj663) - (0.64 * adj630), # Calculate chla and chlc values based on equations from Jeffery and Humphrey 1975
         chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663),
         chla.ug.ml = chla.ug.ml*0.584, # path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
         chlc2.ug.ml = chlc2.ug.ml*0.584)
```

## Bind all dfs together 
```{r}
all <- rbind(p1, p2, p3, p4, p5)

# Remove blanks
all <- all %>% 
  filter(!ID == "Blank") %>%
  dplyr::rename(PlugID = ID)
```

## Read in surface area, homogenate volume and metadata 
```{r}
surface_area <- read.csv("../data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("../data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("../data/MasterFragmentSheet.csv", header = T, na.strings = "") %>%
  dplyr::filter(Metric == "P/R, molec") %>%
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

# Merge dataframes
metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
#metadata$vol_mL <- gsub("NA", "", metadata$vol_mL)
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge metadata and chl data and normalize data 
```{r}
chl <- left_join(metadata, all) %>%
       mutate(chla.ug.cm2 = (chla.ug.ml * vol_mL) / Area_cm, # I need to calculate surface area squared?
             chlc2.ug.cm2 = (chlc2.ug.ml * vol_mL) / Area_cm) %>%
       na.omit() ## omitting NAs for now 
```

## Summarize 
```{r}
mean.chla <- summarySE(chl, measurevar = "chla.ug.cm2", groupvars = c("Treatment", "Timepoint"))
mean.chlc2 <- summarySE(chl, measurevar = "chlc2.ug.cm2", groupvars = c("Treatment", "Timepoint"))

mean.chla_site <- summarySE(chl, measurevar = "chla.ug.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
mean.chlc2_site <- summarySE(chl, measurevar = "chlc2.ug.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
# I did not equally sample across site, which was a mistake on my part. I need to figure out how to present the data with site 
```

## Save data 
```{r}
write.csv(chl, "../output/Physiology/Chlorophyll/Chlorophyll.csv")
```

# Chlorophyll a

## Read in chl data
```{r}
chl <- read.csv("../output/Physiology/Chlorophyll/Chlorophyll.csv")
```

## Check for outliers 
```{r}
# Remove field data for now 
chl_exp <- chl %>%
  filter(!Treatment == "Field")

# Look for outliers visually
ggplot(chl_exp, aes(x = Treatment, y = chla.ug.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(chl_exp$chla.ug.cm2, 0.25)
Q3 <- quantile(chl_exp$chla.ug.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- chl_exp[chl_exp$chla.ug.cm2 < lower_bound | chl_exp$chla.ug.cm2 > upper_bound, ]
print(outliers)

# Z-score method
chl_exp$z_score <- scale(chl_exp$chla.ug.cm2)
outliers_z <- chl_exp[abs(chl_exp$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(chl_exp$chla.ug.cm2, 0.25)
Q3 <- quantile(chl_exp$chla.ug.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
chl_exp_filtered <- chl_exp[chl_exp$chla.ug.cm2 >= lower_bound & chl_exp$chla.ug.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(chl_exp)
dim(chl_exp_filtered)
```

12 outliers removed. 

Data is not normal, perform data transformation
```{r}
ggplot(chl_exp_filtered, aes(x = chla.ug.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

chl_exp_filtered$chla_sqrt <- sqrt(chl_exp_filtered$chla.ug.cm2)
shapiro.test(chl_exp_filtered$chla_sqrt)
qqnorm(chl_exp_filtered$chla_sqrt)
qqline(chl_exp_filtered$chla_sqrt, col = "red")
```

Assess normality of square root transformed data 
```{r}
# Histogram
ggplot(chl_exp_filtered, aes(x = chla_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(chl_exp_filtered$chla_sqrt)
qqline(chl_exp_filtered$chla_sqrt, col = "red")

# Shapiro-Wilk Test
shapiro.test(chl_exp_filtered$chla_sqrt)
```

Set TP0 to Ambient for stats purposes
```{r}
chl_exp_filtered <- chl_exp_filtered %>%
  mutate(Treatment = ifelse(Timepoint == "TP0", "Ambient", as.character(Treatment)))
```

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
chl_exp_filtered$Treatment <- as.factor(chl_exp_filtered$Treatment)
chl_exp_filtered$Timepoint <- as.factor(chl_exp_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(chla_sqrt ~ Treatment * Timepoint, data = chl_exp_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(chla_sqrt ~ Treatment * Timepoint, data = chl_exp_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

Timepoint is slightly significant but Treatment and interaction is not. 

Run post-hoc test for Timepoint
```{r}
TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)

plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
chl_exp_filtered <- chl_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

# Update to Acclimation for plotting purposes 
chl_exp_filtered <- chl_exp_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Plot sqrt
```{r}
# Set Month as factor 
chl_exp_filtered$Month <- factor(chl_exp_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# Create the plot
chla <- ggplot(chl_exp_filtered, aes(x = Month, y = chla_sqrt, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  theme_classic() +
  labs(x = "",
       y = "Chlorophyll-a (sqrt-transformed µg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
      theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(size = 30,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); chla
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")
ggsave("../output/Physiology/Chlorophyll/Chla_Sqrt_Treatment_Timepoint.pdf", chla, width = 20, height = 15)
ggsave("../output/Physiology/Chlorophyll/Chla_Sqrt_Treatment_Timepoint.png", chla, width = 20, height = 15)
```

Plot untransformed data 
```{r}
anova_summary <- summary(anova_result)

# Extract p-values
p_treatment <- anova_summary[[1]]$`Pr(>F)`[1]  # Treatment p-value
p_timepoint <- anova_summary[[1]]$`Pr(>F)`[2]  # Timepoint p-value
p_interaction <- anova_summary[[1]]$`Pr(>F)`[3]  # Interaction p-value

# Format text for annotation
anova_text <- paste0("Two-way ANOVA Results:",
                     "Treatment p = ", signif(p_treatment, 3),
                     "Timepoint p = ", signif(p_timepoint, 3),
                     "Interaction p = ", signif(p_interaction, 3))

# Set Month as factor 
chl_exp_filtered$Month <- factor(chl_exp_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# Create the plot
chla <- ggplot(chl_exp_filtered, aes(x = Month, y = chla.ug.cm2, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  theme_classic() +
  labs(x = "",
       y = "Chlorophyll-a (µg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
      theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(size = 30,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); chla
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint.pdf", chla, width = 20, height = 15)
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint.png", chla, width = 20, height = 15)
```

Calculate mean, sd and sem of each group for plotting 
```{r}
chla_summary <- chl_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_chla = mean(chla.ug.cm2, na.rm = TRUE),
    sd_chla = sd(chla.ug.cm2, na.rm = TRUE),
    sem_chla = sd_chla / sqrt(n),  # SEM formula
    mean_chlasqrt = mean(chla_sqrt, na.rm = TRUE),
    sd_chlasqrt = sd(chla_sqrt, na.rm = TRUE),
    sem_chlasqrt = sd_chlasqrt / sqrt(n),
    .groups = "drop"
  )
```

Add letters for plotting 
```{r}
month_letters_manual <- tibble::tibble(
  Month = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"),
  Letters = c("a", "a", "a", "a", "a", "a", "a", "a"),
  y_position = c(1.77, 1.65, 2.12, 2.29, 2.39, 2.01, 1.78, 2.76) # 0.2 offset from highest sd 
)
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
chla_summary$Month <- factor(chla_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
chla_dot_sem <- ggplot(chla_summary, aes(x = Month, y = mean_chla, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_chla - sem_chla, 
                    ymax = mean_chla + sem_chla), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  geom_text(data = month_letters_manual,
            aes(x = Month, y = y_position, label = Letters),
            inherit.aes = FALSE,
            position = position_dodge(width = 0.5),
            size = 10) +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  labs(y = expression(Chlorophyll-a~(µg~cm^{-2})), x = "") +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); chla_dot_sem
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sem.pdf", chla_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sem.png", chla_dot_sem, width = 20, height = 15)

# sd
chla_dot_sd <- ggplot(chla_summary, aes(x = Month, y = mean_chla, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_chla - sd_chla, 
                    ymax = mean_chla + sd_chla),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  geom_text(data = month_letters_manual,
            aes(x = Month, y = y_position, label = Letters),
            inherit.aes = FALSE,
            position = position_dodge(width = 0.5),
            size = 10) +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  labs(y = expression(bold(Chlorophyll-a~(µg~cm^{-2}))), x = "") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.position = "top",
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );chla_dot_sd

ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sd.pdf", chla_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sd.png", chla_dot_sd, width = 22, height = 12)
```

# Remove February and rerun stats

Read in chl data
```{r}
chl <- read.csv("../output/Physiology/Chlorophyll/Chlorophyll.csv")
```

Remove field and Feb samples 
```{r}
chl_exp <- chl %>%
  filter(!Treatment == "Field") %>%
  filter(!Timepoint == "TP0")
```

Check for outliers 
```{r}
# Look for outliers visually
ggplot(chl_exp, aes(x = Treatment, y = chla.ug.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(chl_exp$chla.ug.cm2, 0.25)
Q3 <- quantile(chl_exp$chla.ug.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- chl_exp[chl_exp$chla.ug.cm2 < lower_bound | chl_exp$chla.ug.cm2 > upper_bound, ]
print(outliers)

# Z-score method
chl_exp$z_score <- scale(chl_exp$chla.ug.cm2)
outliers_z <- chl_exp[abs(chl_exp$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(chl_exp$chla.ug.cm2, 0.25)
Q3 <- quantile(chl_exp$chla.ug.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
chl_exp_filtered <- chl_exp[chl_exp$chla.ug.cm2 >= lower_bound & chl_exp$chla.ug.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(chl_exp)
dim(chl_exp_filtered)
```

Assess normality of data 
```{r}
# Histogram
ggplot(chl_exp_filtered, aes(x = chla.ug.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(chl_exp_filtered$chla.ug.cm2)
qqline(chl_exp_filtered$chla.ug.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(chl_exp_filtered$chla.ug.cm2)
```

Data is not normal, perform sqrt transformation
```{r}
ggplot(chl_exp_filtered, aes(x = chla.ug.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

chl_exp_filtered$chla_sqrt <- sqrt(chl_exp_filtered$chla.ug.cm2)
shapiro.test(chl_exp_filtered$chla_sqrt)
qqnorm(chl_exp_filtered$chla_sqrt)
qqline(chl_exp_filtered$chla_sqrt, col = "red")
```

Assess normality of sqrt transformed data 
```{r}
# Histogram
ggplot(chl_exp_filtered, aes(x = chla_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(chl_exp_filtered$chla_sqrt)
qqline(chl_exp_filtered$chla_sqrt, col = "red")

# Shapiro-Wilk Test
shapiro.test(chl_exp_filtered$chla_sqrt)
```

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
chl_exp_filtered$Treatment <- as.factor(chl_exp_filtered$Treatment)
chl_exp_filtered$Timepoint <- as.factor(chl_exp_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(chla_sqrt ~ Treatment * Timepoint, data = chl_exp_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(chla_sqrt ~ Treatment * Timepoint, data = chl_exp_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```
Timepoint is marginally significant (p=0.043) but Treatment and interaction is not. 

Run post-hoc test for Timepoint
```{r}
TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Assign month to TPs
```{r}
chl_exp_filtered <- chl_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Calculate mean, sd and sem of each group for plotting 
```{r}
chla_summary <- chl_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_chla = mean(chla.ug.cm2, na.rm = TRUE),
    sd_chla = sd(chla.ug.cm2, na.rm = TRUE),
    sem_chla = sd_chla / sqrt(n),  # SEM formula
    mean_chlasqrt = mean(chla_sqrt, na.rm = TRUE),
    sd_chlasqrt = sd(chla_sqrt, na.rm = TRUE),
    sem_chlasqrt = sd_chlasqrt / sqrt(n),
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
chla_summary$Month <- factor(chla_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
chla_dot_sem <- ggplot(chla_summary, aes(x = Month, y = mean_chla, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_chla - sem_chla, 
                    ymax = mean_chla + sem_chla), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  labs(y = expression(Chlorophyll-a~(µg~cm^{-2})), x = "") +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); chla_dot_sem
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sem_noFeb.pdf", chla_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sem_noFeb.png", chla_dot_sem, width = 20, height = 15)

# sd
chla_dot_sd <- ggplot(chla_summary, aes(x = Month, y = mean_chla, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_chla - sd_chla, 
                    ymax = mean_chla + sd_chla),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  labs(y = expression(bold(Chlorophyll-a~(µg~cm^{-2}))), x = "") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.position = "top",
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );chla_dot_sd

ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sd_noFeb.pdf", chla_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/Chlorophyll/Chla_Treatment_Timepoint_dotwhisker_sd_noFeb.png", chla_dot_sd, width = 22, height = 12)
```


