---
title: "Color_Score"
author: "jillashey"
date: "3/12/2021"
output: html_document
---

Modified from [ES code](https://github.com/hputnam/Acclim_Dynamics/blob/master/Scripts/Photographic_Bleaching.Rmd)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading in required libraries. 
```{r}
if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 
if ("ggpubr" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggpubr') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcompView" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcompView') 
if ("rstatix" %in% rownames(installed.packages()) == 'FALSE') install.packages('rstatix') 
#Read in required libraries
##### Include Versions of libraries
library("vegan")
library("ggpubr")
library("gridExtra")
library("plyr") 
library("emmeans")
library("multcompView")
library("rstatix")
library("Rmisc")
```

Reading in quality controlled version of ImageJ data.
Protocol for ImageJ Color Score Analysis found [here](https://emmastrand.github.io/EmmaStrand_Notebook/ImageJ-Analysis-Protocols/).  
```{r}
Data <- read.csv("data/ColorScore/Color_Score.csv", header=T, sep=",", na.string="NA") #read in data file
unique(Data$Treatment)
unique(Data$Timepoint)
```

Normalizing to color standards. 
```{r}
Data$Red.Norm.Coral <- Data$Red.Coral/Data$Red.Standard
Data$Green.Norm.Coral <- Data$Green.Coral/Data$Green.Standard
Data$Blue.Norm.Coral <- Data$Blue.Coral/Data$Blue.Standard

par(mfrow=c(1,3))
boxplot(Data$Red.Standard)
boxplot(Data$Green.Standard)
boxplot(Data$Blue.Standard)

which((Data$Red.Standard)==min((Data$Red.Standard)))
which((Data$Green.Standard)==min((Data$Green.Standard)))
which((Data$Blue.Standard)==min((Data$Blue.Standard)))

par(mfrow=c(1,3))
boxplot(Data$Red.Norm.Coral)
boxplot(Data$Green.Norm.Coral)
boxplot(Data$Blue.Norm.Coral)



```

```{r}
color.score <- as.matrix(cbind(Data$Red.Norm.Coral,Data$Green.Norm.Coral,Data$Blue.Norm.Coral)) #create matrix
rownames(color.score) <- Data$Fragment.ID #name columns in dataframe
dist <- vegdist(color.score, method="euclidean") #calculate distance matrix of color scores
PCA.color <- princomp(dist) #run principal components Analysis
summary(PCA.color) # view variance explained by PCs
Color <- as.data.frame(PCA.color$scores[,1]) #extract PC1
Color$Fragment.ID <- rownames(color.score)
Color  <- cbind(Color, Data$Timepoint, Data$Treatment)
colnames(Color) <- c("Color.Score", "Fragment.ID", "Timepoint", "Treatment")        
Color$Group <- paste(Color$Timepoint, Color$Treatment)
```

Output: Color Score dataframe and checking for outliers in the dataset. 
```{r}
write.table(Color, file = "output/Color_Score_Output.csv", append = FALSE, quote = TRUE, sep = ",",
            eol = "\n", na = "NA", dec = ".", row.names = TRUE,
            col.names = TRUE, qmethod = c("escape", "double"),
            fileEncoding = "")
x<- Color #assign values to test data frame to look for outliers
par(mar=c(10,4,2,2)) #bottom, left, top and right margins respectively
boxplot(Color.Score ~ Group, data = x, lwd = 1, ylab = 'PC1Color', las=2, cex=0.8) #plot boxplot of PC1 color score by treatment and timepoint
```

** Statistical Analysis and Plotting ** 
```{r}
# for use in plotting to avoid having to run PCA again
Color.Score <- read.csv("output/Color_Score_Output.csv", header=T, sep=",", na.string="NA") #read in data file
mod1 <- aov(sqrt(Color.Score+200) ~ Treatment*Timepoint, data=Color.Score) #run an ANOVA ## add Timepoint as factor later 
hist(residuals(mod1)) #look at normality of data
boxplot(residuals(mod1)) #look at normality of data
summary(mod1)
#posthoc analyses
marginal <- emmeans(mod1, ~ Treatment)
#letter display of posthoc results
phlett <- multcomp::cld(marginal,
          alpha=0.05,
          Letters=letters,
          adjust="tukey")
phlett <- phlett[order(phlett$Treatment),]
phlett
```

Calculating means, sample size, and standard error for timepoint and treatment groups. 
```{r}


All.Means <- summarySE(Color.Score, measurevar="Color.Score", groupvars=c("Treatment", "Timepoint")) %>% 
  filter(N>2) %>% na.omit()
#All.Means <- ddply(Color.Score, c('Timepoint','Species','Treatment'), summarize,
                  #  mean= mean(Bleaching.Score, na.rm=T), #mean pnet
                 #   N = sum(!is.na(Bleaching.Score)), # sample size
                  #  se = sd(Bleaching.Score, na.rm=T)/sqrt(N)) #SE
#All.Means$se[is.na(All.Means$se)] <- 0
#All.Means$Group <- paste(All.Means$Timepoint, All.Means$Treatment, All.Means$Species)
#All.Means$SpGroup <- paste(All.Means$Treatment, All.Means$Species)
#All.Means <- All.Means[-c(87:94),]
#All.Means$Timepoint <- factor(All.Means$Timepoint, levels = c("Week1", "Week2", "Week3", "Week4", "Week5", "Week6", "Week7", "Week8", "Week9", "Week10", "Week11", "Week12", "Week13", "Week14", "Week15", "Week16")) 
#All.Means <- All.Means %>% subset(Timepoint!="Week5")
```

Plot data 
```{r}
Color.Score_plot <- ggplot(All.Means, aes(x=Timepoint, y = Color.Score, group=Treatment, shape=Treatment)) +
  geom_line() + geom_point() +
  #facet_grid(~Species, scales = "free", labeller = species_labeller) +
  geom_errorbar(aes(ymin=Color.Score-se, ymax=Color.Score+se), width=.2) +
  xlab("Treatment") + 
  ylab("Color Score") + # y-axis label and y axis range
  theme_classic() + 
  theme(legend.position = "left") +
  geom_vline(xintercept = c(7.8), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  #scale_x_discrete(breaks = c("Week1", "Week2", "Week4", "Week6", "Week8", "Week12", "Week16"), labels = c("1 week", "2 week", "4 week", "6 week", "8 week", "12 week", "16 week")) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change axis line
  axis.line = element_line(colour = "black"))
Color.Score_plot
ggsave(file="output/Color_Score.pdf", Color.Score_plot, width = 11, height = 6, units = c("in"))
```



















