---
title: "DailyMeasures"
author: "JAshey"
date: "2/2/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
```

# Check working directory
```{r}
getwd() 
```

# Load data
```{r}
# Load data
daily <- read.csv("data/DailyMeasurements.csv", header = T)
daily <- daily[-13,] # remove tank 14 because it cracked 
daily$Timepoint <- as.Date(daily$Timepoint, "%m/%d/%Y")
daily$Tank <- as.factor(daily$Tank)
daily <- subset(daily, Treatment=="Heat" | Treatment=="Ambient")
```

# Discrete pH calculations from Tris calibrations.
```{r}
path <-("data/Environmental/pH_Tris") #set path to calibration files
file.names<-list.files(path = path, pattern = "csv$") #list all the file names in the folder to get only get the csv files
pH.cals <- data.frame(matrix(NA, nrow=length(file.names), ncol=3, dimnames=list(file.names,c("Date", "Intercept", "Slope")))) #generate a 3 column dataframe with specific column names

for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Calib.Data <-read.table(file.path(path,file.names[i]), header=TRUE, sep=",", na.string="NA", as.is=TRUE) #reads in the data files
  file.names[i]
  model <-lm(mVTris ~ TTris, data=Calib.Data) #runs a linear regression of mV as a function of temperature
  coe <- coef(model) #extracts the coeffecients
  summary(model)$r.squared #extracts the r squared
  plot(Calib.Data$mVTris, Calib.Data$TTris, main = summary(model)$r.squared) #plots the regression data
  pH.cals[i,2:3] <- coe #inserts coefficients in the dataframe
  pH.cals[i,1] <- substr(file.names[i],1,8) #stores the file name in the Date column
}
colnames(pH.cals) <- c("Calib.Date",  "Intercept",  "Slope") #rename columns
pH.cals #view data

#constants for use in pH calculation 
R <- 8.31447215 #gas constant in J mol-1 K-1 
F <- 96485.339924 #Faraday constant in coulombs mol-1

#merge with Seawater chemistry file
SW.chem <- merge(daily, pH.cals, by="Calib.Date")
```

# Calculate total pH. 
```{r}
mvTris <- SW.chem$Temperature*SW.chem$Slope+SW.chem$Intercept #calculate the mV of the tris standard using the temperature mv relationships in the measured standard curves 
STris<-35 #salinity of the Tris
phTris<- (11911.08-18.2499*STris-0.039336*STris^2)*(1/(SW.chem$Temperature+273.15))-366.27059+ 0.53993607*STris+0.00016329*STris^2+(64.52243-0.084041*STris)*log(SW.chem$Temperature+273.15)-0.11149858*(SW.chem$Temperature+273.15) #calculate the pH of the tris (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
SW.chem$pH.Total<-phTris+(mvTris/1000-SW.chem$pH.MV/1000)/(R*(SW.chem$Temperature+273.15)*log(10)/F) #calculate the pH on the total scale (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
```

# Plot daily measurements 
```{r}
# By Treatment
temp.trt = ggplot(SW.chem, aes(x=Treatment, y=Temperature.C)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Temperature°C") +
  theme(axis.text.x = element_text(angle = 90))
temp.trt
sal.trt = ggplot(SW.chem, aes(x=Treatment, y=Salinity.psu)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Salinity.psu") +
    theme(axis.text.x = element_text(angle = 90))
sal.trt
pH.trt = ggplot(SW.chem, aes(x=Treatment, y=pH.Total)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("pH Total Scale") +
  theme(axis.text.x = element_text(angle = 90))
pH.trt
light.trt = ggplot(SW.chem, aes(x=Treatment, y=Light)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Light µmol photons m^2 s^-1") +
    theme(axis.text.x = element_text(angle = 90))
light.trt
flow.trt = ggplot(SW.chem, aes(x=Treatment, y=Flow.mL.sec)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Flow mL/sec") +
    theme(axis.text.x = element_text(angle = 90))
flow.trt
plot.trt <- grid.arrange(temp.trt, sal.trt, pH.trt, light.trt, flow.trt, ncol=3, nrow = 2, clip="off")
ggsave("output/Daily/Daily_Measurements.By.Treatment.pdf", plot.trt, width = 21, height = 21, units = c("in"))

# By Tank
temp.tank = ggplot(SW.chem, aes(x=Tank, y=Temperature.C)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Temperature°C") +
    theme(axis.text.x = element_text(angle = 90))
temp.tank
sal.tank = ggplot(SW.chem, aes(x=Tank, y=Salinity.psu)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Salinity.psu") +
    theme(axis.text.x = element_text(angle = 90))
sal.tank
pH.tank = ggplot(SW.chem, aes(x=Tank, y=pH.Total)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("pH Total Scale") +
    theme(axis.text.x = element_text(angle = 90))
pH.tank
light.tank = ggplot(SW.chem, aes(x=Tank, y=Light)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Light µmol photons m^2 s^-1") +
    theme(axis.text.x = element_text(angle = 90))
light.tank
flow.tank = ggplot(SW.chem, aes(x=Tank, y=Flow.mL.sec)) +
  geom_boxplot(aes(color = Treatment), size = 1) +
  ylab("Flow mL/sec") +
    theme(axis.text.x = element_text(angle = 90))
flow.tank
plot.tank <- grid.arrange(temp.tank, sal.tank, pH.tank, light.tank, flow.tank, ncol=3, nrow = 2, clip="off")
#ggsave("output/Daily/Daily_Measurements.By.Tank.pdf", plot.tank, width = 21, height = 21, units = c("in"))
```

# test for differences between tanks and treatments
```{r}
# By Treatment
mod.temp <- aov(Temperature.C ~ Treatment, data = SW.chem)
mod.temp
summary(mod.temp)
hist(mod.temp$residuals)

mod.sal <- aov(Salinity.psu ~ Treatment, data = SW.chem)
mod.sal
summary(mod.sal)
hist(mod.sal$residuals)

mod.pH <- aov(pH.Total ~ Treatment, data = SW.chem)
mod.pH
summary(mod.pH)
hist(mod.pH$residuals)

mod.light <- aov(Light ~ Treatment, data = SW.chem)
mod.light
summary(mod.light)
hist(mod.light$residuals)

mod.flow <- aov(Flow.mL.sec ~ Treatment, data = SW.chem)
mod.flow
summary(mod.flow)
hist(mod.flow$residuals)

# By Tank
mod.temp <- aov(Temperature.C ~ Tank, data = SW.chem)
mod.temp
summary(mod.temp)
hist(mod.temp$residuals)

mod.sal <- aov(Salinity.psu ~ Tank, data = SW.chem)
mod.sal
summary(mod.sal)
hist(mod.sal$residuals)

mod.pH <- aov(pH.Total ~ Tank, data = SW.chem)
mod.pH
summary(mod.pH)
hist(mod.pH$residuals)

mod.light <- aov(Light ~ Tank, data = SW.chem)
mod.light
summary(mod.light)
hist(mod.light$residuals)

mod.flow <- aov(Flow.mL.sec ~ Tank, data = SW.chem)
mod.flow
summary(mod.flow)
hist(mod.flow$residuals)
```

# Plot by date 
```{r}
# Reshape the data from wide to long format
SW.chem_long <- SW.chem %>%
  pivot_longer(cols = c(Temperature.C, Salinity.psu, pH.Total),
               names_to = "Variable",
               values_to = "Value")

# Create the plot
date_daily<- ggplot(SW.chem_long, aes(x = Timepoint, y = Value, color = Treatment)) +
  geom_point() +
  #geom_line() +
  facet_wrap(~ Variable, scales = "free_y", ncol = 1) +
  labs(title = "Temperature, Salinity, and pH Total by Timepoint",
       x = "Timepoint",
       y = "Value") +
  theme_bw() +
  theme(legend.position = "bottom"); date_daily


#ggsave("output/Environmental/Daily/Daily_Measurements.By.Date_facet.pdf", plot.all, width = 30, height = 25, units = c("in"))
#ggsave("output/Daily/Daily_Measurements.By.Date_facet.png", plot.all, width = 30, height = 25, units = c("in"))
```

Make date plot pretty!
```{r}
SW.chem_clean <- SW.chem %>%
  mutate(
    # Convert Timepoint with proper year format
    Timepoint = as.Date(gsub("^0021", "2021", Timepoint)),
    # Convert numeric Date to proper date format
    Date = as.Date(as.character(Date), "%Y%m%d")
  ) %>%
  pivot_longer(
    cols = c(Temperature.C, Salinity.psu, pH.Total),
    names_to = "Parameter",
    values_to = "Value"
  ) 

# Create final plot with vertical line
daily_plot <- ggplot(SW.chem_clean, aes(x = Date, y = Value, color = Treatment)) +
  geom_point(size = 2.5) +
  geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = as.Date("2021-02-06"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-03-04"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-04-02"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-04-29"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-05-28"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-06-25"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-07-23"), linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = as.Date("2021-08-25"), linetype = "dashed", color = "black", size = 1) +
  facet_wrap(~ Parameter, ncol = 1, scales = "free_y", strip.position = "right") +
  scale_color_manual(values = c(Ambient = "#377EB8", Heat = "#E41A1C")) +
  scale_x_date(
    breaks = seq(min(SW.chem_clean$Date), max(SW.chem_clean$Date), by = "1 month"),
    labels = date_format("%b"),
    expand = expansion(mult = 0.05)
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(axis.ticks = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 15),
    #strip.placement = "outside",
    strip.text = element_text(size = 20, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black", size = 30),
    panel.spacing = unit(1, "lines")
  ); daily_plot
ggsave("output/Environmental/Daily/Daily_Measurements.By.Date_facet.pdf", daily_plot, width = 15, height = 20)
ggsave("output/Environmental/Daily/Daily_Measurements.By.Date_facet.png", daily_plot, width = 15, height = 20)
```


# Plot DMs for acclimation 

# Subset Acclimation 
```{r}
daily <- read.csv("data/DailyMeasurements.csv", header = T)
daily <- daily[-13,] # remove tank 14 because it cracked 
daily$Timepoint <- as.Date(daily$Timepoint, "%m/%d/%Y")
daily$Tank <- as.factor(daily$Tank)
daily_acc <- daily %>%
  filter(Treatment == "Acclimation")
```

# Discrete pH calculations from Tris calibrations.
```{r}
path <-("data/Environmental/pH_Tris") #set path to calibration files
file.names<-list.files(path = path, pattern = "csv$") #list all the file names in the folder to get only get the csv files
pH.cals <- data.frame(matrix(NA, nrow=length(file.names), ncol=3, dimnames=list(file.names,c("Date", "Intercept", "Slope")))) #generate a 3 column dataframe with specific column names

for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Calib.Data <-read.table(file.path(path,file.names[i]), header=TRUE, sep=",", na.string="NA", as.is=TRUE) #reads in the data files
  file.names[i]
  model <-lm(mVTris ~ TTris, data=Calib.Data) #runs a linear regression of mV as a function of temperature
  coe <- coef(model) #extracts the coeffecients
  summary(model)$r.squared #extracts the r squared
  plot(Calib.Data$mVTris, Calib.Data$TTris, main = summary(model)$r.squared) #plots the regression data
  pH.cals[i,2:3] <- coe #inserts coefficients in the dataframe
  pH.cals[i,1] <- substr(file.names[i],1,8) #stores the file name in the Date column
}
colnames(pH.cals) <- c("Calib.Date",  "Intercept",  "Slope") #rename columns
pH.cals #view data

#constants for use in pH calculation 
R <- 8.31447215 #gas constant in J mol-1 K-1 
F <- 96485.339924 #Faraday constant in coulombs mol-1

#merge with Seawater chemistry file
SW.chem_acc <- merge(daily_acc, pH.cals, by="Calib.Date")
```

# Calculate total pH. 
```{r}
mvTris <- SW.chem_acc$Temperature*SW.chem_acc$Slope+SW.chem_acc$Intercept #calculate the mV of the tris standard using the temperature mv relationships in the measured standard curves 
STris<-35 #salinity of the Tris
phTris<- (11911.08-18.2499*STris-0.039336*STris^2)*(1/(SW.chem_acc$Temperature+273.15))-366.27059+ 0.53993607*STris+0.00016329*STris^2+(64.52243-0.084041*STris)*log(SW.chem_acc$Temperature+273.15)-0.11149858*(SW.chem_acc$Temperature+273.15) #calculate the pH of the tris (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
SW.chem_acc$pH.Total<-phTris+(mvTris/1000-SW.chem_acc$pH.MV/1000)/(R*(SW.chem_acc$Temperature+273.15)*log(10)/F) #calculate the pH on the total scale (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
```

# Plot 
```{r}
SW.chem_acc <- SW.chem_acc %>%
  filter(pH.Total > 7) %>%
  filter(Salinity.psu > 27) %>%
  mutate(Temperature.C_adj = Temperature.C - 3)

SW.chem_acc_clean <- SW.chem_acc %>%
  mutate(
    # Convert Timepoint with proper year format
    Timepoint = as.Date(gsub("^0021", "2021", Timepoint)),
    # Convert numeric Date to proper date format
    Date = as.Date(as.character(Date), "%Y%m%d")
  ) %>%
  pivot_longer(
    cols = c(Temperature.C_adj, Salinity.psu, pH.Total),
    names_to = "Parameter",
    values_to = "Value"
  ) 

# Create final plot with vertical line
daily_plot <- ggplot(SW.chem_acc_clean, aes(x = Date, y = Value)) +
  geom_point(size = 2.5) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ Parameter, ncol = 1, scales = "free_y", strip.position = "right") +
  scale_x_date(
    breaks = seq(min(SW.chem_acc_clean$Date), 
                max(SW.chem_acc_clean$Date), 
                by = "1 week"),
    labels = date_format("%b %d"),
    expand = expansion(mult = 0.05)
  ) +
  labs(y = NULL, x = NULL) +
  theme_bw() +
  theme(axis.ticks = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 15),
    #strip.placement = "outside",
    strip.text = element_text(size = 20, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black", size = 30),
    panel.spacing = unit(1, "lines")
  ); daily_plot
ggsave("output/Environmental/Daily/Acclimation_Daily_Measurements.By.Date_facet.pdf", daily_plot, width = 15, height = 20)
ggsave("output/Environmental/Daily/Acclimation_Daily_Measurements.By.Date_facet.png", daily_plot, width = 15, height = 20)
```

Plot as boxplot 
```{r}
daily_plot_box <- ggplot(SW.chem_acc_clean, aes(x = Treatment, y = Value)) +
  geom_boxplot()+
  facet_wrap(~ Parameter, ncol = 1, scales = "free_y", strip.position = "right")+
    labs(y = NULL, x = NULL) +
  theme_bw() +
  theme(axis.ticks = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 15),
    #strip.placement = "outside",
    strip.text = element_text(size = 20, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black", size = 30),
    panel.spacing = unit(1, "lines")
  ); daily_plot_box
ggsave("output/Environmental/Daily/Acclimation_Daily_Measurements_facet.pdf", daily_plot_box, width = 15, height = 20)
ggsave("output/Environmental/Daily/Acclimation_Daily_Measurements_facet.png", daily_plot_box, width = 15, height = 20)
```

# Plot PAR 
```{r}
# Load data
daily <- read.csv("data/DailyMeasurements.csv", header = T)
daily <- daily[-13,] # remove tank 14 because it cracked 
daily$Timepoint <- as.Date(daily$Timepoint, "%m/%d/%Y")
daily$Tank <- as.factor(daily$Tank)

# Plot by trt
daily_light <- ggplot(daily, aes(x = Treatment, y = Light, color = Treatment, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA, size = 1, alpha = 0.4) +  
  labs(y = "PAR", x = NULL) +
  scale_color_manual(values = c(Acclimation = "grey60", Ambient = "#377EB8", Heat = "#E41A1C")) +
  scale_fill_manual(values = c(Acclimation = "grey60", Ambient = "#377EB8", Heat = "#E41A1C")) +
  theme_bw() +
  theme(axis.ticks = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 15),
    #strip.placement = "outside",
    strip.text = element_text(size = 20, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black", size = 30),
    axis.title = element_text(size = 32, color = "black"),
    panel.spacing = unit(1, "lines")
  ); daily_light

ggsave("output/Environmental/Daily/Light_Measurements_Treatment.pdf", daily_light, width = 15, height = 10)
ggsave("output/Environmental/Daily/Light_Measurements_Treatment.png", daily_light, width = 15, height = 10)
```




