---
title: "glmmSeq - Astrangia mRNA"
output: html_document
date: "2023-10-02"
---

This script analyzses the raw mRNA count data using the glmmSeq package as described in the [vignette](https://cran.r-project.org/web/packages/glmmSeq/vignettes/glmmSeq.html) and [manual](https://cran.r-project.org/web/packages/glmmSeq/glmmSeq.pdf). This script is based on ZD's [script](https://github.com/imkristenbrown/Heron-Pdam-gene-expression/blob/master/BioInf/scripts/glmmSeq/Slope_Base/glmmSeq.Rmd).

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

sessionInfo() 
# R version 4.3.1 (2023-06-16)
# Platform: x86_64-apple-darwin20 (64-bit)
# Running under: macOS Ventura 13.5.1
```

Install and load packages 
```{r install packages, eval=FALSE}
BiocManager::install("qvalue", force = TRUE)
install.packages("rlang")
install.packages("glmmSeq") #from CRAN
devtools::install_github("myles-lewis/glmmSeq") # from github
install.packages("plotly")
BiocManager::install("genefilter")
BiocManager::install("DESeq2")

library(lme4)
library(glmmSeq)
library(dplyr) #load dplyr, which is required to run this tutorial but not included in the vignette
library(kableExtra) #load kableExtra, which is also required to run this tutorial but not included in the vignette
library(plotly) #library plotly to make interactive plots
library(genefilter)
library(DESeq2)
library(pheatmap)

sessionInfo()
```


## Load and clean data 

Read in raw count data for mRNA 
```{r}
raw_counts <- read.csv("../data/Molecular/mRNA/Apoc_gene_count_matrix_bowtie.csv")
rownames(raw_counts) <- raw_counts[,1] #set first column that contains gene names as rownames
raw_counts <- raw_counts[,-1] #remove the column with gene names
```

Remove extraneous info from sample name (eg trimmed.AST.1065_R1_001.fastq.gz.bam.gtf should be AST-1065)
```{r}
colnames(raw_counts)<-gsub("_[^_]+$", "",colnames(raw_counts)) #get rid of "_001.fastq.gz.bam.gtf"
colnames(raw_counts)<-gsub("_[^_]+$", "",colnames(raw_counts)) #get rid of "_R1"
colnames(raw_counts)<-gsub("trimmed.", "",colnames(raw_counts)) #get rid of "trimmed."
colnames(raw_counts)<-gsub("AST.", "AST-",colnames(raw_counts)) # replace "AST." with "AST-"
colnames(raw_counts)<-gsub("align.", "", colnames(raw_counts)) #get rid of "align."
```

Read in metadata 
```{r}
meta <- read.csv("../data/Molecular/RNA_metadata.csv")
meta <- dplyr::arrange(meta, ID) # rearrange metadata so IDs are sorted in descending order 

# Set variables as factors 
meta$Timepoint <- factor(meta$Timepoint, levels = c("TP0", "TP5", "TP7"))
meta$Treatment <- factor(meta$Treatment, levels = c("Acclimation","Ambient", "Heat"))
```

Data sanity checks!
```{r}
all(rownames(meta$ID)) %in% colnames(raw_counts) #are all of the sample names (rows) in the metadata df in the gene count matrix? Should be TRUE. Says FALSE right now, but a visual inspection shows me that all of them are there...
all(rownames(meta$ID) == colnames(raw_counts)) #are they the same in the same order? Should be TRUE
```

Remove AST-1105 from the data, as it did not have good mapping and has basically no counts. 
```{r}
# Remove column from gene count matrix 
raw_counts <- raw_counts %>%
  dplyr::select(-"AST-1105")
  
# Remove row from metadata
meta <- meta %>% 
  filter(ID != "AST-1105")

all(rownames(meta$ID) %in% colnames(raw_counts)) # must come out TRUE
```

## Filter dataset 

Use filtered data instead of raw counts as input
```{r}
filt_outrm <- raw_counts[rowSums(!as.matrix(raw_counts)) < ncol(raw_counts), ] #here we remove all genes that were not expressed in any of our samples
```

Use pOverA filtering to reduce the dataset 
```{r}
ffun<-filterfun(pOverA(0.5,10))  #set up filtering parameters--LOOK INTO FILTERING PARAMETERS
filt_outrm_poa <- genefilter((filt_outrm), ffun) #apply filter
sum(filt_outrm_poa) #count number of genes left

filt_outrm_poa <- filt_outrm[filt_outrm_poa,] #keep only rows that passed filter
```

Data sanity checks!
```{r}
all(rownames(meta$ID)) %in% colnames(filt_outrm_poa) #are all of the sample names (rows) in the metadata df in the gene count matrix? Should be TRUE. Says FALSE right now, but a visual inspection shows me that all of them are there...
all(rownames(meta$ID) == colnames(filt_outrm_poa)) #are they the same in the same order? Should be TRUE
```

## Calculate dispersion 

"Using negative binomial models requires gene dispersion estimates to be made. This can be achieved in a number of ways. A common way to calculate this for gene i is to use the equation:

Dispersioni = (variancei - meani)/meani2

Starting with raw counts, so using [DESeq2](http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html), but could also import a normalized gene count matrix (i.e. transcripts per million) and calculate disperson manually as in the [vignette](https://cran.r-project.org/web/packages/glmmSeq/vignettes/glmmSeq.html).

```{r}
dds_filt_outrm <- DESeqDataSetFromMatrix(countData = filt_outrm_poa,
                              colData = meta,
                              design = ~1) # CHECK IF DESIGN SHOULD BE ~1

dds_filt_outrm <- DESeq(dds_filt_outrm) #apply DEseq dispersion calculation

dispersions_filt_outrm <- setNames(dispersions(dds_filt_outrm), rownames(filt_outrm_poa)) #save dispersions as vector for each gene

dds_norm_counts <- as.data.frame(counts(dds_filt_outrm, normalized=TRUE)) #save normalized counts from DEseq object for downstream use

vst <- vst(dds_filt_outrm)
```

Using vst object, make a heatmap and PCA plot of samples
```{r}
# Plot heat-map of sample-to-sample distances
sampleDists <- dist(t(assay(vst))) # calculate distance matrix
sampleDistMatrix <- as.matrix(sampleDists) # create distance matrix
rownames(sampleDistMatrix) <- colnames(vst) # assign row names
colnames(sampleDistMatrix) <- NULL # assign col names 
heatmap <- pheatmap(sampleDistMatrix, # plot matrix
         clustering_distance_rows = sampleDists, # cluster rows
         clustering_distance_cols = sampleDists) # cluster cols

# Plot PCAs of all samples 
vst_PCAdata <- plotPCA(vst, intgroup = c("Treatment", "Timepoint"), returnData=TRUE) # create PCA loadings 
percentVar <- round(100*attr(vst_PCAdata, "percentVar")) #plot PCA of samples with all data
PCAplot <- ggplot(vst_PCAdata, aes(PC1, PC2, color=Treatment, shape = Timepoint)) + 
   geom_point(size=3) +
   geom_text(aes(label=name),hjust=0, vjust=0) +
   xlab(paste0("PC1: ",percentVar[1],"% variance")) +
   ylab(paste0("PC2: ",percentVar[2],"% variance")) +
   #scale_color_manual(values = c(control="cadetblue3", mid="palevioletred", high="darkgreen", unknown="black")) +
   coord_fixed() +
   #ggtitle("P. lobata - all genes") +
   theme_bw() + #Set background color
   theme(panel.border = element_blank(), # Set border
         #panel.grid.major = element_blank(), #Set major gridlines
         #panel.grid.minor = element_blank(), #Set minor gridlines
         axis.line = element_line(colour = "black"), #Set axes color
         plot.background=element_blank()) #Set the plot background
PCAplot 

# Looks like we have 2 outliers, but most of the samples are clustering around one another
```

Save vst object for downstream analysis 
```{r}
vst <- assay(vst) # call only the transformed counts
vst <- as.data.frame(vst)
vst <- tibble::rownames_to_column(vst, var = "Gene")

#write.csv(vst, "../output/vst_mRNA.csv") #save this as csv for downstream analysis
```

There is also an option to include size factors for each gene.
```{r, eval=FALSE}
sizeFactors_filt <- estimateSizeFactorsForMatrix(filt_outrm_poa)

## look into why there is such a range of size factors for each sample 
```

## Fit models 

To fit a model for one gene over time we use a formula such as: gene expression ~ fixed effects + random effects

We want to use the treatment as fixed effect and timepoint as the random effect: gene expression ~ Treatment + (1|Timepoint).

To fit this model for all genes we can use the glmmSeq function.
```{r, eval=FALSE}
results <- glmmSeq(~ Treatment + (1|Timepoint),
                   countdata = filt_outrm_poa,
                   metadata = meta,
                   dispersion = dispersions_filt_outrm,
                   progress = TRUE)

#this fitting takes a while, so I save it as an RDS and load this in when needed instead of re-running everytime (unless something upstream of here changes)

saveRDS(results, file = "../output/Molecular/mRNA/glmmSeq.rds")  #save RDS
```

Look at output 
```{r}
names(attributes(results))

results@modelData
results@errors
summary(results@stats)

summary(results, gene = "evm.TU.chromosome_10.776 ")
```

Obtain Q values
```{r}
results <- glmmQvals(results)

# Treatment
# ---------
# Not Significant     Significant 
#           19116            2664 
```

Only has information on 

Model fit stats
```{r}
stats <- summary(results)
```







