---
title: "Physiology"
author: "jillashey"
date: "2023-02-16"
output: html_document
---

## Overview 


## Load libraries 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# load packages
library(tidyverse)
library(broom)
```

Did soluble protein test run w/ Holo, Host, and Sym fractions on 2/16/23. I'm now going to compare the soluble protein results across holo, host, and sym. 

## Import data
```{r}
# raw data
raw.data <- read.csv("data/Physiology/Protein/20230216_testplate_data.csv")
raw.data$X562 <- as.numeric(raw.data$X562)

# plate map
platemap <- read.csv("data/Physiology/Protein/20230216_testplate_platemap.csv")

data <- full_join(raw.data, platemap, by = "Well") %>% na.omit() # join dfs by Well 

# Add extra column for holo, host, sym
data$partner <- gsub(".*_", "", data$ID)
```

## Separate out standards and plot standards 
```{r}
# separate out standards from samples 
standards <- data %>% filter(grepl("Standard", ID)) %>% dplyr::rename(Standard = ID)

# group by std name and concentration 
st.name <- c("StandardA", "StandardB", "StandardC", "StandardD", "StandardE", "StandardF", "StandardG", "StandardH", "StandardI")
concentration <- c("2000", "1500", "1000", "750", "500", "250", "125", "25", "0")

standards <- standards %>%
  dplyr::group_by(Standard) %>%       
  dplyr::summarise(X562 = mean(X562)) %>%       # avg standards    
  mutate(X562.adj = X562 - X562[Standard == "StandardI"]) # subtract avg standard from standard I (blank)

standards$BSA_ug.mL <- concentration[match(standards$Standard, st.name)] # match concentration to std name 
standards$BSA_ug.mL <- as.numeric(standards$BSA_ug.mL) # make 

nlm <- lm(BSA_ug.mL ~ X562.adj, data = standards) # fit linear model 
summary(nlm)

# plot standard data 
st.plot <- standards %>%
  ggplot(aes(x = X562.adj, y = BSA_ug.mL)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  geom_smooth(method = "lm") + 
  labs(title = paste("Adj R2 = ", signif(summary(nlm)$adj.r.squared, 5),
                     "; Soluble Protein Standard Curve Pierca BCA Purple Kit")); st.plot
```

## Calculate protein based on standard curve
```{r}
data <- data %>% 
  dplyr::group_by(ID) %>%
  summarise(X562 = mean(X562)) %>% 
  mutate(X562.adj = X562 - standards$X562[standards$Standard == "StandardI"],  
         prot_ug.mL.562 = map_dbl(X562.adj, ~ predict(nlm, newdata = data.frame(X562.adj = .))))

st.plot + 
  geom_point(data = data, aes(x = X562.adj, y = prot_ug.mL.562), pch = "X", cex = 5, alpha = 0.3)
  #labs(title = "March 3 2021 - Purple Kit Values")
```

## Plot protein concentrations by partner (holo, host, sym). This data is not normalized to SA or homogenate vol. 
```{r}
# Isolate sample data only 
sample.data <- data[1:18,]

# Add extra col for partner and ID
sample.data$partner <- gsub(".*_", "", sample.data$ID)
sample.data$Plug_ID <- gsub("_.*", "", sample.data$ID)


# Plot by partner 
sample.data %>% 
  ggplot(aes(x = partner, y = prot_ug.mL.562, fill = partner)) +
  #geom_boxplot() +
  geom_point() +
  ggtitle("Soluble protein concentration (ug/mL)")

# Plot by ID
sample.data %>% 
  ggplot(aes(x = Plug_ID, y = prot_ug.mL.562, color = partner)) +
  geom_point(size = 5) +
  ggtitle("Soluble protein concentration (ug/mL)")

```












Compare with TOTAL protein (soluble and insoluble )
## Load data 
## Import data
```{r}
# raw data
raw.data2 <- read.csv("data/Physiology/Protein/20230217_testplate2_data.csv")
raw.data2$X562 <- as.numeric(raw.data2$X562)

# plate map
platemap <- read.csv("data/Physiology/Protein/20230216_testplate_platemap.csv") # same plate map for both 20230216 (soluble only) and 20230217 (total protein)

data2 <- full_join(raw.data2, platemap, by = "Well") %>% na.omit() # join dfs by Well 

# Add extra column for holo, host, sym
#data$partner <- gsub(".*_", "", data$ID)
```

## Separate out standards and plot standards 
```{r}
# separate out standards from samples 
standards2 <- data2 %>% filter(grepl("Standard", ID)) %>% dplyr::rename(Standard = ID)

# group by std name and concentration 
st.name <- c("StandardA", "StandardB", "StandardC", "StandardD", "StandardE", "StandardF", "StandardG", "StandardH", "StandardI")
concentration <- c("2000", "1500", "1000", "750", "500", "250", "125", "25", "0")

standards2 <- standards2 %>%
  dplyr::group_by(Standard) %>%       
  dplyr::summarise(X562 = mean(X562)) %>%       # avg standards    
  mutate(X562.adj = X562 - X562[Standard == "StandardI"]) # subtract avg standard from standard I (blank)

standards2$BSA_ug.mL <- concentration[match(standards2$Standard, st.name)] # match concentration to std name 
standards2$BSA_ug.mL <- as.numeric(standards2$BSA_ug.mL) # make 

nlm <- lm(BSA_ug.mL ~ X562.adj, data = standards2) # fit linear model 
summary(nlm)

# plot standard data 
st.plot2 <- standards2 %>%
  ggplot(aes(x = X562.adj, y = BSA_ug.mL)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  geom_smooth(method = "lm") + 
  labs(title = paste("Adj R2 = ", signif(summary(nlm)$adj.r.squared, 5),
                     "; Soluble Protein Standard Curve Pierca BCA Purple Kit")); st.plot2
```

## Calculate protein based on standard curve
```{r}
data2 <- data2 %>% 
  dplyr::group_by(ID) %>%
  summarise(X562 = mean(X562)) %>% 
  mutate(X562.adj = X562 - standards$X562[standards$Standard == "StandardI"],  
         prot_ug.mL.562 = map_dbl(X562.adj, ~ predict(nlm, newdata = data.frame(X562.adj = .))))

st.plot2 + 
  geom_point(data = data2, aes(x = X562.adj, y = prot_ug.mL.562), pch = "X", cex = 5, alpha = 0.3)
  #labs(title = "March 3 2021 - Purple Kit Values")
```

## Plot protein concentrations by partner (holo, host, sym). This data is not normalized to SA or homogenate vol. 
```{r}
# Isolate sample data only 
sample.data2 <- data2[1:18,]

# Add extra col for partner and ID
sample.data2$partner <- gsub(".*_", "", sample.data2$ID)
sample.data2$Plug_ID <- gsub("_.*", "", sample.data2$ID)


# Plot by partner 
sample.data2 %>% 
  ggplot(aes(x = partner, y = prot_ug.mL.562, fill = partner)) +
  #geom_boxplot() +
  geom_point() +
  ggtitle("Soluble protein concentration (ug/mL)")

# Plot by ID
sample.data2 %>% 
  ggplot(aes(x = Plug_ID, y = prot_ug.mL.562, color = partner)) +
  geom_point(size = 5) +
  ggtitle("Total protein concentration (ug/mL)")

```

## Plot by soluble and insoluble 
```{r}
# add extra column in each df
sample.data$type <- "Soluble"
sample.data2$type <- "Total"

# check to make sure cols are in same order 
colnames(sample.data)
colnames(sample.data2)

# join dfs
test <- rbind(sample.data, sample.data2)

# plot by ID
test %>% 
  ggplot(aes(x = Plug_ID, y = prot_ug.mL.562, color = partner, shape = type)) +
  geom_point(size = 5) +
  ggtitle("Total protein concentration (ug/mL)")












# join SP and IP dfs
join <- full_join(sample.data, sample.data2, by = "ID")

# plot soluble v insoluble 
plot(join$prot_ug.mL.562.x ~ join$prot_ug.mL.562.y)

# plot by ID
sample.data2 %>% 
  ggplot(aes(x = Plug_ID, y = prot_ug.mL.562, color = partner)) +
  geom_point(size = 5) +
  ggtitle("Total protein concentration (ug/mL)")
```



## Normalize by surface area 













