---
title: LarvalResp_Plotting
author: "AS Huffmyer"
date: '2020'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
Plotting larval resp data, Astrangia Aug 2020

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```


# **Data loading**  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
Rdata<-read.csv("output/LarvalResp/oxygen_R_calc.csv") #load data
```

# **Analysis and plotting**  

Format columns.   

```{r}
Rdata$Org.Number<-as.factor(Rdata$Org.Number)
```

## Respiration  

Plot data as a scatterplot   
```{r}
r_plot<-Rdata %>%
    ggplot(., aes(x = Org.Number, y = R.nmol.org.min)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(group=Org.Number), color="black", outlier.size = 0, position = position_dodge(1), lwd=1) +
    geom_point(aes(group=Org.Number), fill="darkgray", pch = 21, size=4, position = position_dodge(0.2)) + 
    xlab("Number of Larvae") + 
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.015, 0.005), labels = scales::number_format(accuracy = 0.001, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot


ggsave("output/LarvalResp/Respiration.png", r_plot, dpi=300, w=8.5, h=5, units="in")
```

These data indicate that the most consistent measurements of respiration are between 15 and 40 larvae per well. Variability decreases as the number of larvae increase, so using 20-40 larvae per well seems to be the most appropriate. Respiration per larvae decreases at 50 larvae, which may indicate overloading of the well. The consistency in respiration per larvae across larval density indicates a consistent measurement of respiration! 

Target density = 15-40 larvae per well