---
title: "miRanda output parsing"
author: "Jill Ashey"
date: "2024-05-06"
output: html_document
---

This script will read in the miranda output and assign the correct gene id to each comparison.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Read in miranda results. 
```{r}
miranda <- read.delim("~/Desktop/PutnamLab/Astrangia/2020-2021/miranda/miranda_strict_all_parsed.txt", header = F) %>% 
  rename(mirna = V1,
        Target = V2, 
        Score = V3, 
        Energy_kcal_mol = V4, 
        Query_start_end = V5, 
        Subject_start_end = V6,
        Length = V7, 
        Subject_Identity = V8, 
        Query_Identity = V9)

head(miranda)
```

This will likely take a while to load because the file is so large. 

The miRNA column corresponds to the differentially expressed miRNAs. The target column corresponds to the 3'UTR sequence of a differentially expressed gene. To get the corresponding gene name for the 3'UTR, read in gene 3'UTR text file.
```{r}
gene_3utr <- read.delim("output/Molecular/interactions/uniq_modified_closest_genes_3UTRid.txt", header = F) %>%
  dplyr::select(V1, V12) %>%
  rename(UTR = V1, 
         gene_id = V12)

# Remove everything after the space in the UTR column
gene_3utr$UTR <- sub("\\s.*", "", gene_3utr$UTR)
```

Merge the miranda results with the 3UTR info.
```{r}
miranda_gene <- miranda %>%
  inner_join(gene_3utr, by = c("Target" = "UTR"))
```

Remove `>` in miRNA id name
```{r}
miranda_gene$mirna <- gsub(">", "", miranda_gene$mirna)
```

Save file
```{r}
write.csv(miranda_gene, file = "~/Desktop/PutnamLab/Astrangia/2020-2021/miranda/miranda_gene_ids_all.csv")
```







