---
title: "miRanda output parsing"
author: "Jill Ashey"
date: "2024-05-06"
output: html_document
---

This script will read in the miranda output and assign the correct gene id to each comparison.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Read in miranda results. 
```{r}
miranda <- read.delim("~/Desktop/PutnamLab/Astrangia/2020-2021/miranda/miranda_strict_all_parsed.txt", header = F) %>% 
  rename(mirna = V1,
        Target = V2, 
        Score = V3, 
        Energy_kcal_mol = V4, 
        Query_start_end = V5, 
        Subject_start_end = V6,
        Length = V7, 
        Subject_Identity = V8, 
        Query_Identity = V9)

head(miranda)
```

This will likely take a while to load because the file is so large. 

The miRNA column corresponds to the differentially expressed miRNAs. The target column corresponds to the 3'UTR sequence of a differentially expressed gene. To get the corresponding gene name for the 3'UTR, read in gene 3'UTR text file.
```{r}
gene_3utr <- read.delim("output/Molecular/interactions/uniq_modified_closest_genes_3UTRid.txt", header = F) %>%
  dplyr::select(V1, V12) %>%
  rename(UTR = V1, 
         gene_id = V12)

# Remove everything after the space in the UTR column
gene_3utr$UTR <- sub("\\s.*", "", gene_3utr$UTR)
```

Merge the miranda results with the 3UTR info.
```{r}
miranda_gene <- miranda %>%
  inner_join(gene_3utr, by = c("Target" = "UTR"))
```

Remove `>` in miRNA id name
```{r}
miranda_gene$mirna <- gsub(">", "", miranda_gene$mirna)
```

Save file
```{r}
write.csv(miranda_gene, file = "~/Desktop/PutnamLab/Astrangia/2020-2021/miranda/miranda_gene_ids_all.csv")
```







## 20260108 

Using tab file to determine interactions with a 13-mer seed match with <= 2 mismatches and exact pairing between 10-11bp. 
```{r}
#!/usr/bin/env Rscript

# Diagnostic script to explore miranda results and test filtering criteria

# Function to parse miranda output
parse_miranda_output <- function(input_file) {
  lines <- readLines(input_file)
  
  results <- list()
  
  i <- 1
  while (i <= length(lines)) {
    # Look for alignment blocks
    if (grepl("^   Forward:", lines[i])) {
      forward_line <- lines[i]
      query_line <- lines[i + 1]
      alignment_line <- lines[i + 2]
      ref_line <- lines[i + 3]
      energy_line <- lines[i + 4]
      
      # Find the score line (starts with >)
      score_line_idx <- NA
      for (j in (i+5):min(i+15, length(lines))) {
        if (grepl("^>", lines[j])) {
          score_line_idx <- j
          break
        }
      }
      
      if (!is.na(score_line_idx)) {
        score_line <- lines[score_line_idx]
        
        # Extract sequences
        query_seq <- gsub(".*[35]'\\s+([ATGCUatgcu]+)\\s+[35]'.*", "\\1", query_line)
        ref_seq <- gsub(".*[35]'\\s+([ATGCUatgcu]+)\\s+[35]'.*", "\\1", ref_line)
        alignment <- gsub(".*\\s{3,}(.+)\\s*$", "\\1", alignment_line)
        
        # Extract energy
        energy <- as.numeric(gsub(".*Energy:\\s+([0-9.-]+).*", "\\1", energy_line))
        
        # Parse score line
        score_parts <- strsplit(gsub("^>", "", score_line), "\t")[[1]]
        
        if (length(score_parts) >= 9) {
          results[[length(results) + 1]] <- list(
            miRNA = score_parts[1],
            target = score_parts[2],
            score = as.numeric(score_parts[3]),
            energy = as.numeric(score_parts[4]),
            miRNA_pos = score_parts[5],
            target_pos = score_parts[6],
            align_len = as.numeric(score_parts[7]),
            identity = score_parts[8],
            similarity = score_parts[9],
            query_seq = query_seq,
            ref_seq = ref_seq,
            alignment = alignment
          )
        }
      }
    }
    i <- i + 1
  }
  
  return(results)
}

# Function to analyze complementarity
analyze_pairing <- function(query_seq, ref_seq) {
  query_clean <- toupper(gsub("[^ATGCUatgcu]", "", query_seq))
  ref_clean <- toupper(gsub("[^ATGCUatgcu]", "", ref_seq))
  
  len <- min(nchar(query_clean), nchar(ref_clean))
  
  matches <- 0
  wobbles <- 0  # G:U wobble pairs
  
  pairing_pattern <- character(len)
  
  for (i in 1:len) {
    q <- substr(query_clean, i, i)
    r <- substr(ref_clean, i, i)
    
    # Watson-Crick pairs (miRNA is 3'->5', target is 5'->3')
    if ((q == "A" && (r == "T" || r == "U")) ||
        ((q == "T" || q == "U") && r == "A") ||
        (q == "G" && r == "C") ||
        (q == "C" && r == "G")) {
      matches <- matches + 1
      pairing_pattern[i] <- "|"
    } else if ((q == "G" && (r == "T" || r == "U")) ||
               ((q == "T" || q == "U") && r == "G")) {
      wobbles <- wobbles + 1
      pairing_pattern[i] <- ":"
    } else {
      pairing_pattern[i] <- " "
    }
  }
  
  return(list(
    matches = matches,
    wobbles = wobbles,
    mismatches = len - matches - wobbles,
    length = len,
    pairing_pattern = paste(pairing_pattern, collapse = "")
  ))
}

# Function to analyze seed regions with different criteria
analyze_seed_regions <- function(results) {
  seed_analysis <- data.frame()
  
  for (i in 1:length(results)) {
    r <- results[[i]]
    
    # Analyze different seed region lengths
    for (seed_len in c(6, 7, 8, 9, 10, 11, 12, 13, 14)) {
      if (nchar(r$query_seq) >= seed_len) {
        # Extract seed region (first N bases of alignment)
        seed_query <- substr(r$query_seq, 1, seed_len)
        seed_ref <- substr(r$ref_seq, 1, seed_len)
        
        pairing <- analyze_pairing(seed_query, seed_ref)
        
        seed_analysis <- rbind(seed_analysis, data.frame(
          target_num = i,
          miRNA = r$miRNA,
          target = r$target,
          energy = r$energy,
          score = r$score,
          seed_length = seed_len,
          matches = pairing$matches,
          wobbles = pairing$wobbles,
          mismatches = pairing$mismatches,
          match_percent = round(100 * pairing$matches / seed_len, 1),
          stringsAsFactors = FALSE
        ))
      }
    }
  }
  
  return(seed_analysis)
}

# Main execution
cat("=== Miranda Output Diagnostic Tool ===\n\n")

input_file <- "~/Desktop/PutnamLab/Astrangia/2020-2021/miranda/miranda_strict_all_1kb_apoc_shortstack.tab"  # Change to your input file name

# Parse all results
cat("Parsing miranda output...\n")
all_results <- parse_miranda_output(input_file)
cat("Found", length(all_results), "total predictions\n\n")

if (length(all_results) == 0) {
  cat("No results found. Check your input file format.\n")
  quit()
}

# Display first few examples
cat("=== First 3 Examples ===\n")
for (i in 1:min(3, length(all_results))) {
  r <- all_results[[i]]
  cat("\nTarget", i, ":\n")
  cat("miRNA:  ", r$miRNA, "\n")
  cat("Target: ", r$target, "\n")
  cat("Score:  ", r$score, "  Energy: ", r$energy, "\n")
  cat("Query (3'->5'): ", r$query_seq, "\n")
  cat("Alignment:      ", r$alignment, "\n")
  cat("Ref (5'->3'):   ", r$ref_seq, "\n")
  
  pairing <- analyze_pairing(r$query_seq, r$ref_seq)
  cat("Pairing:        ", pairing$pairing_pattern, "\n")
  cat("Matches:", pairing$matches, "Wobbles:", pairing$wobbles, "Mismatches:", pairing$mismatches, "\n")
}

# Analyze seed regions
cat("\n=== Seed Region Analysis ===\n")
seed_data <- analyze_seed_regions(all_results)

# Summary statistics by seed length
cat("\nTargets passing different seed matching criteria:\n")
cat("(Seed length | Max mismatches allowed | Targets passing)\n")
for (seed_len in c(7, 8, 9, 10, 11, 12, 13)) {
  for (max_mm in 0:3) {
    passing <- sum(seed_data$seed_length == seed_len & seed_data$mismatches <= max_mm)
    cat(sprintf("  %2d bases  |  ≤%d mismatches         | %4d targets\n", 
                seed_len, max_mm, passing))
  }
}

# Distribution of energy scores
cat("\n=== Energy Distribution ===\n")
energies <- sapply(all_results, function(x) x$energy)
cat("Min energy:", min(energies), "\n")
cat("Mean energy:", round(mean(energies), 2), "\n")
cat("Median energy:", round(median(energies), 2), "\n")
cat("Energies < -20:", sum(energies < -20), "targets\n")
cat("Energies < -25:", sum(energies < -25), "targets\n")

# Save detailed results for exploration
write.csv(seed_data, "seed_analysis.csv", row.names = FALSE)
cat("\nDetailed seed analysis saved to: seed_analysis.csv\n")

# Save all alignments for manual inspection
alignment_details <- data.frame()
for (i in 1:length(all_results)) {
  r <- all_results[[i]]
  pairing <- analyze_pairing(r$query_seq, r$ref_seq)
  
  alignment_details <- rbind(alignment_details, data.frame(
    miRNA = r$miRNA,
    target = r$target,
    score = r$score,
    energy = r$energy,
    align_length = r$align_len,
    query_seq = r$query_seq,
    alignment = r$alignment,
    ref_seq = r$ref_seq,
    pairing_pattern = pairing$pairing_pattern,
    matches = pairing$matches,
    mismatches = pairing$mismatches,
    stringsAsFactors = FALSE
  ))
}

write.csv(alignment_details, "all_alignments.csv", row.names = FALSE)
cat("All alignment details saved to: all_alignments.csv\n")

cat("\n=== Recommendations ===\n")
cat("1. Open 'all_alignments.csv' to manually inspect alignments\n")
cat("2. Open 'seed_analysis.csv' to explore seed matching patterns\n")
cat("3. Look for conserved pairing patterns in positions 2-8 or 2-7\n")
cat("4. Consider combining seed match + energy threshold (e.g., <-20 kcal/mol)\n")
cat("5. Check literature for cnidarian/coral-specific miRNA:target rules\n")
```


input_file <- "~/Desktop/PutnamLab/Astrangia/2020-2021/miranda/miranda_strict_all_1kb_apoc_shortstack.tab"  # Change to your input file name







