---
title: "Comparing miRDeep2 output"
author: "Jill Ashey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = F)

library(tidyverse)
```

I am running mirdeep2 on each of my samples individually. Now I am going to look at the output (known and novel miRNAs) from those scripts. 

AST-1065
```{r}
AST_1065_novel <- read.csv("../data/Molecular/smRNA/mirdeep2/AST-1065-30bp_novel.csv")
length(unique(AST_1065_novel$provisional.id)) # 316 unique genomic locations
length(unique(AST_1065_novel$consensus.mature.sequence)) # 280 unique mature novel miRNA sequences 
```

AST-1147
```{r}
AST_1147_novel <- read.csv("../data/Molecular/smRNA/mirdeep2/AST-1147-30bp_novel.csv")
length(unique(AST_1147_novel$provisional.id)) # 1316 unique genomic locations
length(unique(AST_1147_novel$consensus.mature.sequence)) # 970 unique mature novel miRNA sequences 
```

Merge to see how many seqs they share 
```{r}
test <- full_join(AST_1065_novel, AST_1147_novel, by = "consensus.mature.sequence") %>%
  filter(miRDeep2.score.x > 10) %>%
  filter(miRDeep2.score.y > 10) %>%
  filter(significant.rannovel_filtold.p.value.x == "yes") %>%
  filter(significant.rannovel_filtold.p.value.y == "yes") %>%
  na.omit()
length(unique(test$consensus.mature.sequence)) # 24 seqs shared 
```

1/30/24
Now I have run mirdeep2 on all samples and the known and novel info. I'm going to try reading in all files and comparing if the samples have any sequences in common
```{r}
# Check the current working directory
getwd()

# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list
csv_data <- lapply(file_names, read.csv)

# Extract mature sequences from each CSV file
mature_sequences_list <- lapply(csv_data, function(novel_filt) novel_filt$consensus.mature.sequence)

# Find the intersection of mature sequences across all files
shared_mature_sequences <- Reduce(intersect, mature_sequences_list)
## These are all of the sequences shared across all samples 






# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list
csv_data <- lapply(file_names, read.csv)

# Extract mature sequences from each CSV file
mature_sequences_list <- lapply(csv_data, function(novel_filt) novel_filt$consensus.mature.sequence)

# Initialize a list to store groups of shared sequences
shared_sequence_groups <- list()

# Loop through each pair of files and find shared sequences
for (i in 1:(length(file_names) - 1)) {
  for (j in (i + 1):length(file_names)) {
    # Check if i and j are within bounds
    if (i <= length(mature_sequences_list) && j <= length(mature_sequences_list)) {
      # Create a key for the shared sequence group
      key <- paste(file_names[i], file_names[j], sep = " & ")
      
      # Find the intersection of mature sequences between files i and j
      intersection <- intersect(mature_sequences_list[[i]], mature_sequences_list[[j]])
      
      # Store the shared sequences in the shared_sequence_groups list
      shared_sequence_groups[[key]] <- intersection
    } else {
      print("Index i or j out of bounds.")
    }
  }
}
## These are the shared sequences between two files 







# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list
csv_data <- lapply(file_names, read.csv)

# Create a data structure to track which sequences came from which files
sequence_file_mapping <- list()

# Loop through each CSV file and extract mature sequences
for (i in seq_along(file_names)) {
  # Extract mature sequences
  mature_sequences <- csv_data[[i]]$consensus.mature.sequence
  
  # Update sequence_file_mapping
  for (sequence in mature_sequences) {
    if (is.null(sequence_file_mapping[[sequence]])) {
      sequence_file_mapping[[sequence]] <- list()
    }
    sequence_file_mapping[[sequence]][[basename(file_names[i])]] <- TRUE
  }
}

# Find groups of shared sequences across all files
shared_sequences <- list()

for (sequence in names(sequence_file_mapping)) {
  # Check if the sequence is shared across files
  if (length(sequence_file_mapping[[sequence]]) > 1) {
    shared_sequences[[sequence]] <- names(sequence_file_mapping[[sequence]])
  }
}

# Print or further process shared_sequences
print(shared_sequences)
```

Yay!!! I got a large list of the shared sequences across multiple files!!!!! Good job me. Now how to utilize this info usefully...

Save shared seqs as csv -- important to note that these are not filtered by score or anything like that.
```{r}
# Create a data frame to store shared sequences and associated files
shared_sequences_novel_filt <- data.frame(
  sequence = character(),
  files = character(),
  stringsAsFactors = FALSE
)

# Convert shared_sequences list to data frame format
for (sequence in names(shared_sequences)) {
  # Extract sequence and files
  files <- paste(shared_sequences[[sequence]], collapse = ", ")
  
  # Append to the dataframe
  shared_sequences_novel_filt <- rbind(shared_sequences_novel_filt, data.frame(sequence = sequence, files = files))
}

# Save shared sequences as a CSV file
write.csv(shared_sequences_novel_filt, file = "../output/Molecular/smRNA/shared_miRNA_all.csv", row.names = FALSE)
```

Will I need to process the samples individually? Hmmmmmmmm 

Read in data so that I make one massive novel_filt of all files that includes the file name as a column
```{r}
# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list of data frames
csv_data <- lapply(file_names, function(file) {
  # Read CSV file
  novel_filt <- read.csv(file, stringsAsFactors = FALSE)
  
  # Convert miRDeep2.score column to numeric
  novel_filt$miRDeep2.score <- as.numeric(novel_filt$miRDeep2.score)
  
  # Convert read count columns to numeric
  novel_filt$total.read.count <- as.numeric(novel_filt$total.read.count)
  novel_filt$mature.read.count <- as.numeric(novel_filt$mature.read.count)
  novel_filt$loop.read.count <- as.numeric(novel_filt$loop.read.count)
  novel_filt$star.read.count <- as.numeric(novel_filt$star.read.count)
  
  # Add a new column indicating the file name
  novel_filt$file_name <- basename(file)
  
  return(novel_filt)
})

# Combine all data frames into a single dataframe
novel_filt_all <- bind_rows(csv_data)
```

Filter so that mirdeep2 score >= 10, no rfam alerts and significant rannovel_filtold pvalues only. 
```{r}
novel_filt_filt <- novel_filt_all %>%
  filter(!is.na(miRDeep2.score)) %>%
  filter(miRDeep2.score > 10,    # Rows where values in numeric_column are above 10
         significant.rannovel_filtold.p.value == "yes",    # Rows where 'yes' appears in yes_column
        !grepl(c("rRNA/tRNA", "rfam alert"), rfam.alert, ignore.case = TRUE))  # Rows where 'rfam' does not appear in rfam_column
```

We are left with 2154 rows. Lets look at how many sequences we have and how many samples were retained
```{r}
length(unique(novel_filt_filt$consensus.mature.sequence)) # 657 unique seqs
length(unique(novel_filt_filt$file_name)) # 19 files - all samples retained after filtering 
```

Look at the shared miRNAs across samples after filtering 
```{r}
# Group by consensus.mature.sequence and aggregate file names
shared_sequences <- novel_filt_filt %>%
  group_by(consensus.mature.sequence) %>%
  summarize(files_shared = paste(unique(file_name), collapse = ", ")) %>%
  filter(lengths(strsplit(files_shared, ", ")) > 1) # Remove sequences that only have 1 sample associated with it 
```

20240206

Finally finished running mirdeep2 on all of the R1 sequences. Let's look at the novel data
```{r}
novel <- read.csv("../data/Molecular/smRNA/all_novel.csv")
length(unique(novel$provisional.id))
length(unique(novel$consensus.mature.sequence))
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant rannovel_filtold p-value
```{r}
novel_filt <- novel %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

# Save csv 
write.csv(novel_filt, "../data/Molecular/smRNA/novel_filt.csv")
```

Now look at the known data 
```{r}
known <- read.csv("../data/Molecular/smRNA/all_known.csv")
length(unique(known$tag.id))
length(unique(known$consensus.mature.sequence))
```

Because we can see some mapping to known miRNAs in cnidarians, I am only going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant rannovel_filtold p-value
```{r}
known_filt <- known %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

# Save csv 
write.csv(known_filt, "../data/Molecular/smRNA/known_filt.csv")
```

20240226

Ran the quantifier module of mirdeep2 on all samples using the output from mirdeep2 above as input for the mature and precursor sequences. All files have the same number of rows and appear to be ordered in the same way. I'm going to read the files in (there are 19) and merge them together based on the miRNA id. 

```{r}
# Load the first file to initialize the data frame
df <- read.csv("../data/Molecular/smRNA/quantifier/AST-1065_miRNAs_expressed_all_samples_1709048527.csv", sep = "\t", header = T)[,c(1)]
df <- as.data.frame(df)
colnames(df) <- "miRNA"

df2 <- read.csv("../data/Molecular/smRNA/quantifier/AST-1567_miRNAs_expressed_all_samples_1709053231.csv", sep = "\t", header = T)[,c(1)]
df2 <- as.data.frame(df2)
colnames(df2) <- "miRNA"

blah <- cbind(df, df2)
blah <- unique(blah)

are_identical <- identical(df$miRNA, df2$miRNA)
## Can use this to check if rows are identical 

getwd()





test <- merge(df, df2, by = "miRNA")
test <- unique(test)


# Read and left join the rest of the files
file_names <- c(
    "../data/Molecular/smRNA/quantifier/AST-2360_miRNAs_expressed_all_samples_1709055924.csv",
    "../data/Molecular/smRNA/quantifier/AST-1147_miRNAs_expressed_all_samples_1709052026.csv",
    "../data/Molecular/smRNA/quantifier/AST-2398_miRNAs_expressed_all_samples_1709055986.csv",
    "../data/Molecular/smRNA/quantifier/AST-1412_miRNAs_expressed_all_samples_1709053021.csv",
    "../data/Molecular/smRNA/quantifier/AST-2404_miRNAs_expressed_all_samples_1709057037.csv",
    "../data/Molecular/smRNA/quantifier/AST-1560_miRNAs_expressed_all_samples_1709054526.csv",
    "../data/Molecular/smRNA/quantifier/AST-1567_miRNAs_expressed_all_samples_1709053231.csv",
    "../data/Molecular/smRNA/quantifier/AST-2512_miRNAs_expressed_all_samples_1709057192.csv",
    "../data/Molecular/smRNA/quantifier/AST-1617_miRNAs_expressed_all_samples_1709055134.csv",
    "../data/Molecular/smRNA/quantifier/AST-2523_miRNAs_expressed_all_samples_1709057251.csv",
    "../data/Molecular/smRNA/quantifier/AST-1722_miRNAs_expressed_all_samples_1709055319.csv",
    "../data/Molecular/smRNA/quantifier/AST-2563_miRNAs_expressed_all_samples_1709057330.csv",
    "../data/Molecular/smRNA/quantifier/AST-2000_miRNAs_expressed_all_samples_1709055397.csv",
    "../data/Molecular/smRNA/quantifier/AST-2729_miRNAs_expressed_all_samples_1709057412.csv",
    "../data/Molecular/smRNA/quantifier/AST-2007_miRNAs_expressed_all_samples_1709055480.csv",
    "../data/Molecular/smRNA/quantifier/AST-2755_miRNAs_expressed_all_samples_1709057479.csv",
    "../data/Molecular/smRNA/quantifier/AST-2302_miRNAs_expressed_all_samples_1709055687.csv"
)

for (file_name in file_names) {
    # Extract the unique number from the file name
    unique_number <- gsub("AST-", "", substr(file_name, 1, regexpr("_miRNAs", file_name) - 1))
    
    # Read the next file
    next_df <- read.csv(file_name, sep = "\t", header = TRUE)
    
    # Add prefix to selected column names (2 to 6)
    colnames_to_modify <- colnames(next_df)[2:6]
    modified_colnames <- paste0("AST-", unique_number, "_", colnames_to_modify)
    colnames(next_df)[2:6] <- modified_colnames
  
    # Perform left join
    df <- merge(df, next_df, by = "X.miRNA", all.x = TRUE)
}

# Save the final joined dataframe
write.csv(df, "final_joined_file.csv", row.names = FALSE)

```

```{r}
# Define the file names
file_names <- c(
    "../data/Molecular/smRNA/quantifier/AST-2360_miRNAs_expressed_all_samples_1709055924.csv",
    "../data/Molecular/smRNA/quantifier/AST-1147_miRNAs_expressed_all_samples_1709052026.csv",
    "../data/Molecular/smRNA/quantifier/AST-2398_miRNAs_expressed_all_samples_1709055986.csv",
    "../data/Molecular/smRNA/quantifier/AST-1412_miRNAs_expressed_all_samples_1709053021.csv",
    "../data/Molecular/smRNA/quantifier/AST-2404_miRNAs_expressed_all_samples_1709057037.csv",
    "../data/Molecular/smRNA/quantifier/AST-1560_miRNAs_expressed_all_samples_1709054526.csv",
    "../data/Molecular/smRNA/quantifier/AST-1567_miRNAs_expressed_all_samples_1709053231.csv",
    "../data/Molecular/smRNA/quantifier/AST-2512_miRNAs_expressed_all_samples_1709057192.csv",
    "../data/Molecular/smRNA/quantifier/AST-1617_miRNAs_expressed_all_samples_1709055134.csv",
    "../data/Molecular/smRNA/quantifier/AST-2523_miRNAs_expressed_all_samples_1709057251.csv",
    "../data/Molecular/smRNA/quantifier/AST-1722_miRNAs_expressed_all_samples_1709055319.csv",
    "../data/Molecular/smRNA/quantifier/AST-2563_miRNAs_expressed_all_samples_1709057330.csv",
    "../data/Molecular/smRNA/quantifier/AST-2000_miRNAs_expressed_all_samples_1709055397.csv",
    "../data/Molecular/smRNA/quantifier/AST-2729_miRNAs_expressed_all_samples_1709057412.csv",
    "../data/Molecular/smRNA/quantifier/AST-2007_miRNAs_expressed_all_samples_1709055480.csv",
    "../data/Molecular/smRNA/quantifier/AST-2755_miRNAs_expressed_all_samples_1709057479.csv",
    "../data/Molecular/smRNA/quantifier/AST-2302_miRNAs_expressed_all_samples_1709055687.csv"
)

# Create an empty list to store data frames
df_list <- list()

# Read each file into a separate data frame and store it in the list
for (file_name in file_names) {
    # Extract the unique number from the file name
    unique_number <- gsub("AST-", "", substr(file_name, 1, regexpr("_miRNAs", file_name) - 1))
    
    # Read the file and store it in the list
    df <- read.csv(file_name, sep = "\t", header = TRUE)
    
    # Rename columns
    colnames(df)[-1] <- paste0("AST_", unique_number, "_", colnames(df)[-1])
    
    # Store the data frame in the list
    df_list[[unique_number]] <- df
}

# Now df_list contains all your data frames, indexed by the unique numbers
# You can access each data frame using df_list[[unique_number]]

# For example, to access the data frame for AST-1065, you can do:
# ast_1065_df <- df_list[["1065"]]

# Loop through each dataframe in df_list
for (i in seq_along(df_list)) {
    # Remove the last 4 columns
    df_list[[i]] <- df_list[[i]][, -(ncol(df_list[[i]]) - 3):ncol(df_list[[i]])]
}


```






