---
title: "miRNA discovery"
author: "Jill Ashey"
date: "2024-03-06"
output: html_document
---

I ran mirdeep2 on R1 reads from the Astrangia 2021 experiment. The output was files with novel and known miRNAs. In this code, I filter the results to identify putative miRNAs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Load the novel miRNAs
```{r}
novel <- read.csv("../data/Molecular/smRNA/all_novel_20241121.csv")
length(unique(novel$provisional.id))
length(unique(novel$consensus.mature.sequence)) 
```

Filter so that mirdeep2 score >100, no rfam info, at least 10 reads in mature and star read count, and significant randfold p-value. 
```{r}
novel_filt <- novel %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

length(unique(novel_filt$provisional.id))
length(unique(novel_filt$consensus.mature.sequence))

# Save csv 
#write.csv(novel_filt, "../data/Molecular/smRNA/novel_filt.csv")
```

Will need to go through each pdf to determine putative miRNAs. 

Load the known miRNAs
```{r}
known <- read.csv("../data/Molecular/smRNA/all_known_20241121.csv")
length(unique(known$tag.id)) 
length(unique(known$consensus.mature.sequence)) 
```

Because we can see some mapping to known miRNAs in cnidarians, I'm going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant ranfold p-value
```{r}
known_filt <- known %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

length(unique(known_filt$tag.id)) 
length(unique(known_filt$consensus.mature.sequence)) 

# Save csv 
#write.csv(known_filt, "../data/Molecular/smRNA/known_filt.csv")
```

Bind novel and known miRNAs. 

