---
title: "miRNA discovery"
author: "Jill Ashey"
date: "2024-03-06"
output: html_document
---

I ran mirdeep2 on R1 reads from the Astrangia 2021 experiment. The output was files with novel and known miRNAs. In this code, I filter the results to identify putative miRNAs and combine those with miRNA counts for each sample. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Load the novel miRNAs
```{r}
novel <- read.csv("../data/Molecular/smRNA/all_novel.csv")
length(unique(novel$provisional.id)) # 1841 unique genomic ids
length(unique(novel$consensus.mature.sequence)) # 1389 unique miRNA sequences
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant randfold p-value. These filtering parameters are based on XXXXXXX
```{r}
novel_filt <- novel %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

# Save csv 
write.csv(novel_filt, "../data/Molecular/smRNA/novel_filt.csv")
```

Select specific columns for downstream analysis
```{r}
novel_filt <- novel_filt %>%
    rename("ID" = "provisional.id",                                                  
            "estimated_true_positive" = "estimated.probability.that.the.miRNA.candidate.is.a.true.positive", 
            "miRBase" = "miRBase.miRNA") %>%
    mutate(Type = "Novel")
```

Load the known miRNAs
```{r}
known <- read.csv("../data/Molecular/smRNA/all_known.csv")
length(unique(known$tag.id)) # 16 unique genomic ids
length(unique(known$consensus.mature.sequence)) # 14 unique miRNA sequences
```

Because we can see some mapping to known miRNAs in cnidarians, I'm going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant ranfold p-value
```{r}
known_filt <- known %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

# Save csv 
write.csv(known_filt, "../data/Molecular/smRNA/known_filt.csv")
```

Select specific columns for downstream analysis
```{r}
known_filt <- known_filt %>%
    rename("ID" = "tag.id",                                                  
            "estimated_true_positive" = "estimated.probability.that.the.miRNA.is.a.true.positive", 
            "miRBase" = "mature.miRBase.miRNA") %>%
    mutate(Type = "Known")
```

Bind 
```{r}
all_miRNA_filt <- rbind(novel_filt, known_filt) %>%
  select(c("ID", "miRDeep2.score", "estimated_true_positive", "rfam.alert", "total.read.count", "mature.read.count", "loop.read.count", "star.read.count", "significant.randfold.p.value", "consensus.mature.sequence", "consensus.star.sequence", "consensus.precursor.sequence", "precursor.coordinate", "Type"))
```

Read in miRNA counts from each sample
```{r}
# Define a function to read and process CSV files
process_csv <- function(file_path, suffix) {
  df <- read.csv(file_path, sep = "\t", header = TRUE) %>%
    select(X.miRNA, read_count) %>%
    rename(!!paste0("read_count_AST-", suffix) := read_count)
  
  return(df)
}

# List of file paths and corresponding suffixes
file_paths <- c(
  "../data/Molecular/smRNA/quantifier/AST-1065_miRNAs_expressed_all_samples_1709048527.csv",
  "../data/Molecular/smRNA/quantifier/AST-1147_miRNAs_expressed_all_samples_1709052026.csv"
  # Add more file paths here
)

suffixes <- c(
  "1065",
  "1147"
  # Add more suffixes here
)

# Initialize an empty list to store data frames
dfs <- list()

# Iterate through the file paths and process each CSV file
for (i in seq_along(file_paths)) {
  df <- process_csv(file_paths[i], suffixes[i])
  dfs[[i]] <- df
}

# Check if columns are the same across data frames
identical_columns <- map_lgl(dfs[-1], ~identical(dfs[[1]]$X.miRNA, .x$X.miRNA))

# If all columns are the same, bind the data frames together
if (all(identical_columns)) {
  all <- Reduce(function(x, y) cbind(x, y), dfs)
  # Remove the extra miRNA columns
  #all <- select(all, -contains("X.miRNA"))
} else {
  print("Columns are not identical across data frames.")
}
```

miRNA cunts for all samples are in one df but there are still a bunch of extra miRNA id columns. Put the miRNAs in a separate df, remove the miRNA id columns, and then bind the miRNA id df back so there is only one miRNA id column 
```{r}
miRNA <- as.data.frame(all$X.miRNA)

all <- all %>%
  select(-contains("miRNA"))

all <- cbind(miRNA, all) %>%
    rename("id" = "all$X.miRNA")

write.csv(all, "../data/Molecular/smRNA/miRNA_count_matrix.csv")
```






