---
title: "Comparing miRDeep2 output"
author: "Jill Ashey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = F)

library(tidyverse)
```

I am running mirdeep2 on each of my samples individually. Now I am going to look at the output (known and novel miRNAs) from those scripts. 

AST-1065
```{r}
AST_1065_novel <- read.csv("../data/Molecular/smRNA/mirdeep2/AST-1065-30bp_novel.csv")
length(unique(AST_1065_novel$provisional.id)) # 316 unique genomic locations
length(unique(AST_1065_novel$consensus.mature.sequence)) # 280 unique mature novel miRNA sequences 
```

AST-1147
```{r}
AST_1147_novel <- read.csv("../data/Molecular/smRNA/mirdeep2/AST-1147-30bp_novel.csv")
length(unique(AST_1147_novel$provisional.id)) # 1316 unique genomic locations
length(unique(AST_1147_novel$consensus.mature.sequence)) # 970 unique mature novel miRNA sequences 
```

Merge to see how many seqs they share 
```{r}
test <- full_join(AST_1065_novel, AST_1147_novel, by = "consensus.mature.sequence") %>%
  filter(miRDeep2.score.x > 10) %>%
  filter(miRDeep2.score.y > 10) %>%
  filter(significant.rannovel_filtold.p.value.x == "yes") %>%
  filter(significant.rannovel_filtold.p.value.y == "yes") %>%
  na.omit()
length(unique(test$consensus.mature.sequence)) # 24 seqs shared 
```

1/30/24
Now I have run mirdeep2 on all samples and the known and novel info. I'm going to try reading in all files and comparing if the samples have any sequences in common
```{r}
# Check the current working directory
getwd()

# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list
csv_data <- lapply(file_names, read.csv)

# Extract mature sequences from each CSV file
mature_sequences_list <- lapply(csv_data, function(novel_filt) novel_filt$consensus.mature.sequence)

# Find the intersection of mature sequences across all files
shared_mature_sequences <- Reduce(intersect, mature_sequences_list)
## These are all of the sequences shared across all samples 






# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list
csv_data <- lapply(file_names, read.csv)

# Extract mature sequences from each CSV file
mature_sequences_list <- lapply(csv_data, function(novel_filt) novel_filt$consensus.mature.sequence)

# Initialize a list to store groups of shared sequences
shared_sequence_groups <- list()

# Loop through each pair of files and find shared sequences
for (i in 1:(length(file_names) - 1)) {
  for (j in (i + 1):length(file_names)) {
    # Check if i and j are within bounds
    if (i <= length(mature_sequences_list) && j <= length(mature_sequences_list)) {
      # Create a key for the shared sequence group
      key <- paste(file_names[i], file_names[j], sep = " & ")
      
      # Find the intersection of mature sequences between files i and j
      intersection <- intersect(mature_sequences_list[[i]], mature_sequences_list[[j]])
      
      # Store the shared sequences in the shared_sequence_groups list
      shared_sequence_groups[[key]] <- intersection
    } else {
      print("Index i or j out of bounds.")
    }
  }
}
## These are the shared sequences between two files 







# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list
csv_data <- lapply(file_names, read.csv)

# Create a data structure to track which sequences came from which files
sequence_file_mapping <- list()

# Loop through each CSV file and extract mature sequences
for (i in seq_along(file_names)) {
  # Extract mature sequences
  mature_sequences <- csv_data[[i]]$consensus.mature.sequence
  
  # Update sequence_file_mapping
  for (sequence in mature_sequences) {
    if (is.null(sequence_file_mapping[[sequence]])) {
      sequence_file_mapping[[sequence]] <- list()
    }
    sequence_file_mapping[[sequence]][[basename(file_names[i])]] <- TRUE
  }
}

# Find groups of shared sequences across all files
shared_sequences <- list()

for (sequence in names(sequence_file_mapping)) {
  # Check if the sequence is shared across files
  if (length(sequence_file_mapping[[sequence]]) > 1) {
    shared_sequences[[sequence]] <- names(sequence_file_mapping[[sequence]])
  }
}

# Print or further process shared_sequences
print(shared_sequences)
```

Yay!!! I got a large list of the shared sequences across multiple files!!!!! Good job me. Now how to utilize this info usefully...

Save shared seqs as csv -- important to note that these are not filtered by score or anything like that.
```{r}
# Create a data frame to store shared sequences and associated files
shared_sequences_novel_filt <- data.frame(
  sequence = character(),
  files = character(),
  stringsAsFactors = FALSE
)

# Convert shared_sequences list to data frame format
for (sequence in names(shared_sequences)) {
  # Extract sequence and files
  files <- paste(shared_sequences[[sequence]], collapse = ", ")
  
  # Append to the dataframe
  shared_sequences_novel_filt <- rbind(shared_sequences_novel_filt, data.frame(sequence = sequence, files = files))
}

# Save shared sequences as a CSV file
write.csv(shared_sequences_novel_filt, file = "../output/Molecular/smRNA/shared_miRNA_all.csv", row.names = FALSE)
```

Will I need to process the samples individually? Hmmmmmmmm 

Read in data so that I make one massive novel_filt of all files that includes the file name as a column
```{r}
# Define the directory path where your CSV files are located
directory_path <- "../data/Molecular/smRNA/mirdeep2/novel/"

# List all CSV file names in the directory
file_names <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Read all CSV files into a list of data frames
csv_data <- lapply(file_names, function(file) {
  # Read CSV file
  novel_filt <- read.csv(file, stringsAsFactors = FALSE)
  
  # Convert miRDeep2.score column to numeric
  novel_filt$miRDeep2.score <- as.numeric(novel_filt$miRDeep2.score)
  
  # Convert read count columns to numeric
  novel_filt$total.read.count <- as.numeric(novel_filt$total.read.count)
  novel_filt$mature.read.count <- as.numeric(novel_filt$mature.read.count)
  novel_filt$loop.read.count <- as.numeric(novel_filt$loop.read.count)
  novel_filt$star.read.count <- as.numeric(novel_filt$star.read.count)
  
  # Add a new column indicating the file name
  novel_filt$file_name <- basename(file)
  
  return(novel_filt)
})

# Combine all data frames into a single dataframe
novel_filt_all <- bind_rows(csv_data)
```

Filter so that mirdeep2 score >= 10, no rfam alerts and significant rannovel_filtold pvalues only. 
```{r}
novel_filt_filt <- novel_filt_all %>%
  filter(!is.na(miRDeep2.score)) %>%
  filter(miRDeep2.score > 10,    # Rows where values in numeric_column are above 10
         significant.rannovel_filtold.p.value == "yes",    # Rows where 'yes' appears in yes_column
        !grepl(c("rRNA/tRNA", "rfam alert"), rfam.alert, ignore.case = TRUE))  # Rows where 'rfam' does not appear in rfam_column
```

We are left with 2154 rows. Lets look at how many sequences we have and how many samples were retained
```{r}
length(unique(novel_filt_filt$consensus.mature.sequence)) # 657 unique seqs
length(unique(novel_filt_filt$file_name)) # 19 files - all samples retained after filtering 
```

Look at the shared miRNAs across samples after filtering 
```{r}
# Group by consensus.mature.sequence and aggregate file names
shared_sequences <- novel_filt_filt %>%
  group_by(consensus.mature.sequence) %>%
  summarize(files_shared = paste(unique(file_name), collapse = ", ")) %>%
  filter(lengths(strsplit(files_shared, ", ")) > 1) # Remove sequences that only have 1 sample associated with it 
```

20240206

Finally finished running mirdeep2 on all of the R1 sequences. Let's look at the novel data
```{r}
novel <- read.csv("../data/Molecular/smRNA/all_novel.csv")
length(unique(novel$provisional.id))
length(unique(novel$consensus.mature.sequence))
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant rannovel_filtold p-value
```{r}
novel_filt <- novel %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

# Save csv 
write.csv(novel_filt, "../data/Molecular/smRNA/novel_filt.csv")
```

Now look at the known data 
```{r}
known <- read.csv("../data/Molecular/smRNA/all_known.csv")
length(unique(known$tag.id))
length(unique(known$consensus.mature.sequence))
```

Because we can see some mapping to known miRNAs in cnidarians, I am only going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant rannovel_filtold p-value
```{r}
known_filt <- known %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()

# Save csv 
write.csv(known_filt, "../data/Molecular/smRNA/known_filt.csv")
```





