---
title: "mRNA-miRNA WGCNA correlation"
author: "Jill Ashey"
date: "2025-05-05"
output: html_document
---

Using modules generated from mRNA and miRNA WGCNA to predict correlations between the modules. Then look at putative interactions 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ComplexHeatmap)
library(circlize)
library(tidyverse)
library(igraph)
library(ggraph)
```

Read in mRNA ME data (sample level)
```{r}
MEs_mRNA <- read.csv("../output/Molecular/mRNA/WGCNA/WGCNA_ModuleEigengenes_mRNA.csv")
rownames(MEs_mRNA) <- MEs_mRNA$X
MEs_mRNA$X <- NULL  # Remove the sample name column
```

Read in miRNA ME data (sample level)
```{r}
MEs_miRNA <- read.csv("../output/Molecular/smRNA/shortstack/WGCNA/WGCNA_ModuleEigengenes_miRNA.csv")
rownames(MEs_miRNA) <- MEs_miRNA$X
MEs_miRNA$X <- NULL  # Remove the sample name column
```

Correlate miRNA modules with mRNA modules
```{r}
cor_matrix <- cor(MEs_miRNA, MEs_mRNA, use = "pairwise.complete.obs")

pheatmap(cor_matrix, cluster_rows = TRUE, cluster_cols = TRUE)
```

Calculate cor and pvalues and plot
```{r}
# Calculate cor and pvalues
# Initialize empty matrices for p-values
cor_matrix <- cor(MEs_miRNA, MEs_mRNA, use = "pairwise.complete.obs")
p_matrix <- matrix(NA, nrow = nrow(cor_matrix), ncol = ncol(cor_matrix))
rownames(p_matrix) <- rownames(cor_matrix)
colnames(p_matrix) <- colnames(cor_matrix)

# Calculate p-values for each pair
for (i in 1:nrow(cor_matrix)) {
  for (j in 1:ncol(cor_matrix)) {
    test <- cor.test(MEs_miRNA[, i], MEs_mRNA[, j], use = "pairwise.complete.obs")
    p_matrix[i, j] <- test$p.value
  }
}

# Format r and p into a string like "r = 0.85\np = 0.01"
text_matrix <- matrix(NA, nrow = nrow(cor_matrix), ncol = ncol(cor_matrix))
for (i in 1:nrow(cor_matrix)) {
  for (j in 1:ncol(cor_matrix)) {
    text_matrix[i, j] <- sprintf("r = %.2f\np = %.3f", cor_matrix[i, j], p_matrix[i, j])
  }
}

library(pheatmap)
p <- pheatmap(cor_matrix,
              cluster_rows = TRUE,
              cluster_cols = TRUE,
              display_numbers = text_matrix,
              number_color = "black")

# Save manually 
pdf("../output/Molecular/interactions/WGCNA/mRNA_miRNA_module_cor.pdf", width = 20, height = 20)
print(p)  # must print the object inside the pdf() context
dev.off()
png("../output/Molecular/interactions/WGCNA/mRNA_miRNA_module_cor.png", width = 2000, height = 2000)
print(p)
dev.off()
```

Plot with complexHeatmap
```{r}
# Define which cells are significant
sig_matrix <- p_matrix < 0.05

# Define a function to draw borders
border_fun <- function(j, i, x, y, width, height, fill) {
  grid.rect(x, y, width, height, 
            gp = gpar(col = ifelse(sig_matrix[i,j], "black", NA), 
                      lwd = ifelse(sig_matrix[i,j], 2, 1), 
                      fill = NA))
}

modules_heatmap <- Heatmap(
  cor_matrix,
  cell_fun = border_fun,
  rect_gp = gpar(col = NA),
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  width = unit(20, "cm"),
  height = unit(20, "cm"),
  heatmap_legend_param = list(
    title = "Correlation")); modules_heatmap
# Save manually 
pdf("../output/Molecular/interactions/WGCNA/mRNA_miRNA_module_cor_clean.pdf", width = 25, height = 20)
print(modules_heatmap)  # must print the object inside the pdf() context
dev.off()
png("../output/Molecular/interactions/WGCNA/mRNA_miRNA_module_cor_clean.png", width = 20, height = 20)
print(modules_heatmap)
dev.off()
```

Read in genes and miRNAs with module assignments
```{r}
# mRNA
module_genes <- read.csv("../output/Molecular/mRNA/WGCNA/mRNA_modules.csv")
colnames(module_genes)[colnames(module_genes) == "X"] <- "mRNA"

# miRNA
module_miRNA <- read.csv("../output/Molecular/smRNA/shortstack/WGCNA/miRNA_modules.csv")
colnames(module_miRNA)[colnames(module_miRNA) == "X"] <- "miRNA"
module_miRNA$miRNA <- sub("^(([^_]*_[^_]*))_.*", "\\1", module_miRNA$miRNA)
```

Read in miranda data 
```{r}
miranda_data <- read.delim("../output/Molecular/interactions/miranda_strict_all_1kb_apoc_shortstack_parsed.txt", header = F)
colnames(miranda_data) <- c("miRNA", "mRNA", "score", "energy", "query_start_end", "subject_start_end", "total_bp_shared", "query_similar", "subject_similar")

# Format miranda df 
miranda_data$miRNA <- sub("^>", "", miranda_data$miRNA)  # Remove leading ">"
miranda_data$miRNA <- sub("\\..*", "", miranda_data$miRNA)  # Remove everything from the first period onwards
miranda_data$mRNA <- sub(";.*", "", miranda_data$mRNA)  # Remove everything from "::" onwards
miranda_data$mRNA <- sub("ID=", "", miranda_data$mRNA)  # Remove everything from "::" onwards

dim(miranda_data)
```

Merge the mRNAs and miRNAs to see if they are present in modules. 
```{r}
miranda_data_mRNA <- miranda_data %>%
  left_join(module_genes, by = "mRNA")

miranda_data_mRNA_miRNA <- miranda_data_mRNA %>%
  left_join(module_miRNA, by = "miRNA") %>%
  na.omit()

# Remove excess cols 
miranda_data_mRNA_miRNA <- miranda_data_mRNA_miRNA %>%
  select(-starts_with("AST")) %>%
  select(-starts_with("Gene")) %>%
  rename(mrna_module = Module.x, mirna_module = Module.y)
```

Make cor and pvalue dfs 
```{r}
# Melt the correlation matrix
cor_df <- reshape2::melt(cor_matrix)
colnames(cor_df) <- c("miRNA_module", "mRNA_module", "cor_module")

# Melt the pvalue matrix
p_df <- reshape2::melt(p_matrix)
colnames(p_df) <- c("miRNA_module", "mRNA_module", "p_value_module")

# Merge dfs 
cor_p_df <- merge(cor_df, p_df, by = c("miRNA_module", "mRNA_module"))

# Remove ME from module names 
cor_p_df$miRNA_module <- gsub("ME", "", cor_p_df$miRNA_module)
cor_p_df$mRNA_module <- gsub("ME", "", cor_p_df$mRNA_module)
```

Merge miranda data and cor/pvalue df 
```{r}
merged_data <- left_join(
  miranda_data_mRNA_miRNA,
  cor_p_df,
  by = c("mirna_module" = "miRNA_module", "mrna_module" = "mRNA_module")
)
```

Read in PCC data
```{r}
pcc_df <- read.csv("../output/Molecular/interactions/miRNA_mRNA_interactions.csv")

pcc_df <- pcc_df %>%
  select(miRNA, mRNA, PCC.cor, p_value)
```

Merge with PCC data
```{r}
merged_data_pcc <- left_join(
  merged_data,
  pcc_df, 
  by = c("mRNA", "miRNA")
)
write.csv(merged_data_pcc, "../output/Molecular/interactions/miRNA_mRNA_modules_pcc_miranda.csv")
```

Plot 
```{r}
# Make edge list
edges <- merged_data_pcc %>%
  select(from = miRNA, to = mRNA, PCC.cor)

# Make node list (unique miRNA/mRNA)
nodes <- data.frame(
  id = unique(c(merged_data_pcc$miRNA, merged_data_pcc$mRNA)),
  # Assign module color: for miRNA use mirna_module, for mRNA use mrna_module
  module = c(
    merged_data_pcc$mirna_module[match(unique(merged_data_pcc$miRNA), merged_data_pcc$miRNA)],
    merged_data_pcc$mrna_module[match(unique(merged_data_pcc$mRNA), merged_data_pcc$mRNA)]
  ),
  group = c(
    rep("miRNA", length(unique(merged_data_pcc$miRNA))),
    rep("mRNA", length(unique(merged_data_pcc$mRNA)))
  )
)

g <- graph_from_data_frame(edges, vertices = nodes)

ggraph(g, layout = "igraph", algorithm = 'kk') +
  geom_edge_link(
    aes(width = abs(PCC.cor), color = PCC.cor < 0),
    alpha = 0.6) +
  scale_edge_color_manual(
    values = c("red", "blue"), # TRUE (negative) is blue, FALSE (positive) is red
    labels = c("Positive", "Negative"), 
    name = "PCC.cor sign") +
  #geom_edge_link(aes(width = abs(PCC.cor)), alpha = 0.6, color = "grey50") +
  geom_node_point(aes(color = module, size = 5, shape = group)) +
  scale_edge_width_continuous(range = c(0.5, 3), name = "|PCC.cor|") +
  scale_color_manual(values = c(
    # Add your module colors here, e.g.:
    # "turquoise" = "#1B9E77",
    # "red" = "#D95F02",
    # "yellow" = "#E6AB02",
    # "brown" = "#A6761D",
    # "lightcyan" = "#A6CEE3",
    # "orange" = "#FF7F00",
    # "grey60" = "#999999"
    "turquoise" = "turquoise",
    "red" = "red",
    "yellow" = "yellow",
    "brown" = "brown",
    "lightcyan" = "lightcyan",
    "orange" = "orange",
    "grey60" = "grey60"
    # ... add more as needed
  ), name = "Module") +
  scale_shape_manual(values = c(miRNA = 17, mRNA = 19)) +
  #theme_minimal() +
  theme(legend.position = "right") +
  guides(size = "none") +
  labs(
    title = "miRNA-mRNA Interaction Network",
    subtitle = "Node color by module, edge width by |PCC.cor|"
  )

```

Filter so that pairs must have negative cor_module, p_value_module < 0.05, PCC.cor negative and PCC p_value < 0.05
```{r}
# How many pairs have negative cor_module, p_value_module < 0.05, PCC.cor negative and p_value < 0.05
filtered_rows <- merged_data_pcc[
  merged_data_pcc$cor_module < 0 &
  merged_data_pcc$p_value_module < 0.05 &
  merged_data_pcc$PCC.cor < 0 &
  merged_data_pcc$p_value < 0.05,
]
```

```{r}
# Make edge list
edges <- filtered_rows %>%
  select(from = miRNA, to = mRNA, PCC.cor)

# Make node list (unique miRNA/mRNA)
nodes <- data.frame(
  id = unique(c(filtered_rows$miRNA, filtered_rows$mRNA)),
  # Assign module color: for miRNA use mirna_module, for mRNA use mrna_module
  module = c(
    filtered_rows$mirna_module[match(unique(filtered_rows$miRNA), filtered_rows$miRNA)],
    filtered_rows$mrna_module[match(unique(filtered_rows$mRNA), filtered_rows$mRNA)]
  ),
  group = c(
    rep("miRNA", length(unique(filtered_rows$miRNA))),
    rep("mRNA", length(unique(filtered_rows$mRNA)))
  )
)

g <- graph_from_data_frame(edges, vertices = nodes)

ggraph(g, layout = "igraph", algorithm = 'kk') +
  geom_edge_link(
    aes(width = abs(PCC.cor), color = PCC.cor < 0),
    alpha = 0.6) +
  scale_edge_color_manual(
    values = c("red", "blue"), # TRUE (negative) is blue, FALSE (positive) is red
    labels = c("Positive", "Negative"), 
    name = "PCC.cor sign") +
  #geom_edge_link(aes(width = abs(PCC.cor)), alpha = 0.6, color = "grey50") +
  geom_node_point(aes(color = module, size = 5, shape = group)) +
  scale_edge_width_continuous(range = c(0.5, 3), name = "|PCC.cor|") +
  scale_color_manual(values = c(
    # Add your module colors here, e.g.:
    # "turquoise" = "#1B9E77",
    # "red" = "#D95F02",
    # "yellow" = "#E6AB02",
    # "brown" = "#A6761D",
    # "lightcyan" = "#A6CEE3",
    # "orange" = "#FF7F00",
    # "grey60" = "#999999"
    "turquoise" = "turquoise",
    "red" = "red",
    "yellow" = "yellow",
    "brown" = "brown",
    "lightcyan" = "lightcyan",
    "orange" = "orange",
    "grey60" = "grey60"
    # ... add more as needed
  ), name = "Module") +
  scale_shape_manual(values = c(miRNA = 17, mRNA = 19)) +
  #theme_minimal() +
  theme(legend.position = "right") +
  guides(size = "none") +
  labs(
    title = "miRNA-mRNA Interaction Network",
    subtitle = "Node color by module, edge width by |PCC.cor|"
  )
```

Read in annotation info
```{r}
annot <- read.delim("/Users/jillashey/Desktop/PutnamLab/Astrangia_Genome/Apoculata_v2.0_GeneAnnotation_combined_prelim.txt", header = T)
annot$Protein_ID <- gsub("model", "TU", annot$Protein_ID)
```

Merge `filtered_rows` with annot data
```{r}
filtered_rows_annot <- filtered_rows %>%
  left_join(annot, by = c("mRNA" = "Protein_ID"))
```

Merge `merged_data_pcc` with annot data
```{r}
merged_data_pcc_annot <- merged_data_pcc %>%
  left_join(annot, by = c("mRNA" = "Protein_ID"))
```

I think I need to do PCC per condition...
