yw---
title: "Outdoor Tank Temperature"
author: "jillashey"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages 
```{r}
#Install and Load necessary libraries
if("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 

library(lubridate)
library(tidyverse)
library(emmeans)
library(gridExtra)
```

# Check working directory
```{r}
getwd()
```


## Alice A-frame logger data 
# Read in data 
```{r}
aframe <- read.csv("data/Hobo/temp_light_logger/Outdoor/A-frame_20210708.csv", header=T, na.strings = "NA")[ ,2:4] 
colnames(aframe) <- c("Date.Time", "Temperature", "Light")
aframe$Date.Time <- parse_date_time(aframe$Date.Time, "mdyHMS", tz = "EST")
```

# Plot data from A-frame
```{r}
# Plot temperature data
ggplot(aframe, aes(x=Date.Time, y=Temperature)) +
  geom_line(size = 1)+
  #ylim(5,15)+
  ylab("Temperature 째C")
ggsave("output/Hobo/temp_light_logger/A-frame_HoboTemp.pdf", plot = last_plot(), width = 8, height = 4)

# Plot light data
ggplot(aframe, aes(x=Date.Time, y=Light)) +
  geom_line(size = 1)+
  #ylim(5,15)+
  ylab("Light")
ggsave("output/Hobo/temp_light_logger/A-frame_HoboLight.pdf", plot = last_plot(), width = 8, height = 4)
```

## Outdoor tank logger data 
# Read in data 
```{r}
outdoor <- read.csv("data/Hobo/temp_light_logger/Outdoor/OutdoorTank_20210708.csv", header=T, na.strings = "NA")[ ,2:4] # data from 20201221 onward
colnames(outdoor) <- c("Date.Time", "Temperature", "Light")
outdoor$Date.Time <- parse_date_time(outdoor$Date.Time, "mdyHMS", tz = "EST")
```

# Plot data from outdoor tank
```{r}
# Plot temperature data
ggplot(outdoor, aes(x=Date.Time, y=Temperature)) +
  geom_line(size = 1)+
  #ylim(5,15)+
  ylab("Temperature 째C")
ggsave("output/Hobo/temp_light_logger/OutdoorTank_HoboTemp.pdf", plot = last_plot(), width = 8, height = 4)

# Plot light data
ggplot(outdoor, aes(x=Date.Time, y=Light)) +
  geom_line(size = 1)+
  #ylim(5,15)+
  ylab("Light")
ggsave("output/Hobo/temp_light_logger/OutdoorTank_HoboLight.pdf", plot = last_plot(), width = 8, height = 4)
```

# Plot data from June onwards from outdoor tank (time when outdoor corals for AB were collected)
```{r}
# Filter by June dates
june.outdoor <- outdoor %>% filter(Date.Time >= "2021-06-01 00:00:00")

# Plot June temperature data
ggplot(june.outdoor, aes(x=Date.Time, y=Temperature)) +
  geom_line(size = 1)+
  ylab("Temperature 째C")
ggsave("output/Hobo/temp_light_logger/OutdoorTank_HoboTemp.June.pdf", plot = last_plot(), width = 8, height = 4)

# Plot June light data 
ggplot(june.outdoor, aes(x=Date.Time, y=Light)) +
  geom_line(size = 1)+
  ylab("Light")
ggsave("output/Hobo/temp_light_logger/OutdoorTank_HoboLight.June.pdf", plot = last_plot(), width = 8, height = 4)
```

# Plot data from aframe and outdoor loggers 
```{r}
# Filter by aframe start date 
AB.outdoor <- outdoor %>% filter(Date.Time >= "2021-06-24 00:00:00")

# Add a Tank column for both datasets
AB.outdoor$Tank <- "Outdoor"
aframe$Tank <- "A-frame"

# Bind datasets together 
all <- rbind(AB.outdoor, aframe)

# Plot temperature data 
ggplot() +
  geom_point(data = all, aes(x = Date.Time, y = Temperature, color = Tank)) +
  ylab("Temperature 째C")
ggsave("output/Hobo/temp_light_logger/OutdoorTank-AFrame_HoboTemp.pdf", plot = last_plot(), width = 8, height = 4)

# Plot light data 
ggplot() +
  geom_point(data = all, aes(x = Date.Time, y = Light, color = Tank)) +
  ylab("Light")
ggsave("output/Hobo/temp_light_logger/OutdoorTank-AFrame_HoboLight.pdf", plot = last_plot(), width = 8, height = 4)


```


