---
title: "Physiology plots"
author: "Jill Ashey"
date: "2026-04-05"
output: html_document
---

Script for making physiology plots pretty for publication. All stats can be found in the metric-specific scripts. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggtext)
library(patchwork)
```

## Photosynthesis 

Read in data
```{r}
exp <- read_csv(file = "../output/Physiology/PR/PR_rates.csv")

photo_exp <- exp %>%
  filter(Light_Value > 0)
```

Set dates as timepoints
```{r}
photo_exp <- photo_exp %>%
  mutate(Timepoint = case_when(
    Date == "2021-02-04" ~ "TP0",
    Date == "2021-03-04" ~ "TP1",
    Date == "2021-04-01" ~ "TP2",
    Date == "2021-04-29" ~ "TP3",
    Date == "2021-05-27" ~ "TP4",
    Date == "2021-06-24" ~ "TP5",
    Date == "2021-07-22" ~ "TP6",
    Date == "2021-08-19" ~ "TP7"))
```

Remove outliers using IQR method
```{r}
# Look for outliers visually
ggplot(photo_exp, aes(x = Treatment, y = micromol.cm2.h)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# Calculate IQR
Q1 <- quantile(photo_exp$micromol.cm2.h, 0.25)
Q3 <- quantile(photo_exp$micromol.cm2.h, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
photo_exp_filtered <- photo_exp[photo_exp$micromol.cm2.h >= lower_bound & photo_exp$micromol.cm2.h <= upper_bound, ]

# Check dimensions before and after
dim(photo_exp)
dim(photo_exp_filtered)
```

Assign month to TPs
```{r}
photo_exp_filtered <- photo_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February",
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

photo_exp_filtered <- photo_exp_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = factor(Treatment, levels = c("Acclimation", "Ambient", "Heat")))
```

Calculate mean, sd and sem of each group for plotting 
```{r}
photo_summary <- photo_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_photo = mean(micromol.cm2.h, na.rm = TRUE),
    sd_photo = sd(micromol.cm2.h, na.rm = TRUE),
    se_photo = sd(micromol.cm2.h, na.rm = T) / sqrt(length(micromol.cm2.h)),
    .groups = "drop"
  )
```

Plot photo data 
```{r}
# Set Month as factor 
photo_summary$Month <- factor(photo_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

photo_dot <- ggplot(photo_summary, aes(x = Month, y = mean_photo, color = Treatment)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_photo - se_photo, 
                    ymax = mean_photo + se_photo),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
annotate("text", x = Inf, y = -Inf, label = "Time point p<0.001", 
         hjust = 1.1, vjust = -1.1, size = 8, fontface = "bold") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab(expression(bold('Net Photosynthesis ('*mu~ 'mol' ~cm^-2*~ h^-1*')'))) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, color = "black"),
    axis.title.y = element_text(size = 35, face = "bold", margin = margin(r = 10)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    #legend.title = element_text(size = 35, face = "bold"),
    #legend.text = element_text(size = 25),
    legend.position = "none"
    #legend.key.size = unit(1.5, "cm"),
    #legend.spacing.y = unit(0.5, "cm")
  );photo_dot
```

## Respiration

Read in data
```{r}
exp <- read_csv(file = "../output/Physiology/PR/PR_rates.csv")

resp_exp <- exp %>%
  filter(Light_Value == 0)
```

Set dates as timepoints
```{r}
resp_exp <- resp_exp %>%
  mutate(Timepoint = case_when(
    Date == "2021-02-04" ~ "TP0",
    Date == "2021-03-04" ~ "TP1",
    Date == "2021-04-01" ~ "TP2",
    Date == "2021-04-29" ~ "TP3",
    Date == "2021-05-27" ~ "TP4",
    Date == "2021-06-24" ~ "TP5",
    Date == "2021-07-22" ~ "TP6",
    Date == "2021-08-19" ~ "TP7"))
```

Remove outliers using IQR method
```{r}
# Look for outliers visually
ggplot(resp_exp, aes(x = Treatment, y = micromol.cm2.h)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# Calculate IQR
Q1 <- quantile(resp_exp$micromol.cm2.h, 0.25)
Q3 <- quantile(resp_exp$micromol.cm2.h, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
resp_exp_filtered <- resp_exp[resp_exp$micromol.cm2.h >= lower_bound & resp_exp$micromol.cm2.h <= upper_bound, ]

# Check dimensions before and after
dim(resp_exp)
dim(resp_exp_filtered)
```

Assign month to TPs
```{r}
resp_exp_filtered <- resp_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February",
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

resp_exp_filtered <- resp_exp_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = factor(Treatment, levels = c("Acclimation", "Ambient", "Heat")))
```

Calculate mean, sd and sem of each group for plotting 
```{r}
resp_summary <- resp_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_resp = mean(micromol.cm2.h, na.rm = TRUE),
    sd_resp = sd(micromol.cm2.h, na.rm = TRUE),
    se_resp = sd(micromol.cm2.h, na.rm = T) / sqrt(length(micromol.cm2.h)),
    .groups = "drop"
  )
```

Plot resp data 
```{r}
# Set Month as factor 
resp_summary$Month <- factor(resp_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

resp_dot <- ggplot(resp_summary, aes(x = Month, y = mean_resp, color = Treatment)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_resp - se_resp, 
                    ymax = mean_resp + se_resp),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  annotate("text", x = 5, y = -Inf, label = "Time point p<0.001;\nTreatment x Time point p=0.036", 
         hjust = 1.1, vjust = -0.5, size = 8, fontface = "bold") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab(expression(bold('Repsiration ('*mu~ 'mol' ~cm^-2*~ h^-1*')'))) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, color = "black"),
    axis.title.y = element_text(size = 35, face = "bold", margin = margin(r = 10)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    #legend.title = element_text(size = 35, face = "bold"),
    #legend.text = element_text(size = 25),
    legend.position = "none"
    #legend.key.size = unit(1.5, "cm"),
    #legend.spacing.y = unit(0.5, "cm")
  );resp_dot
```

## Chl-a

Read in chl data
```{r}
chl <- read.csv("../output/Physiology/Chlorophyll/Chlorophyll.csv")
```

Remove outliers using IQR method
```{r}
# Remove field data for now 
chl_exp <- chl %>%
  filter(!Treatment == "Field")

# Look for outliers visually
ggplot(chl_exp, aes(x = Treatment, y = chla.ug.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# Calculate IQR
Q1 <- quantile(chl_exp$chla.ug.cm2, 0.25)
Q3 <- quantile(chl_exp$chla.ug.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
chl_exp_filtered <- chl_exp[chl_exp$chla.ug.cm2 >= lower_bound & chl_exp$chla.ug.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(chl_exp)
dim(chl_exp_filtered)
```

Assign months to TPs
```{r}
# Assign month to TPs 
chl_exp_filtered <- chl_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

# Update to Acclimation for plotting purposes 
chl_exp_filtered <- chl_exp_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Calculate mean, sd and sem of each group for plotting 
```{r}
chla_summary <- chl_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_chla = mean(chla.ug.cm2, na.rm = TRUE),
    sd_chla = sd(chla.ug.cm2, na.rm = TRUE),
    sem_chla = sd_chla / sqrt(n),  # SEM formula
    sem_chla_offset = mean_chla + sem_chla + 0.2,
    .groups = "drop"
  )
```

Plot
```{r}
# Set Month as factor 
chla_summary$Month <- factor(chla_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

chla_dot <- ggplot(chla_summary, aes(x = Month, y = mean_chla, color = Treatment)) +
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_chla - sem_chla, 
                    ymax = mean_chla + sem_chla),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  annotate("text", x = Inf, y = -Inf, label = "Time point p=0.047", 
         hjust = 1.1, vjust = -1.1, size = 8, fontface = "bold") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  labs(y = expression(bold(Chlorophyll-a~(µg~cm^{-2})), x = "")) +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, color = "black"),
    axis.title.y = element_text(size = 35, face = "bold", margin = margin(r = 10)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    #legend.title = element_text(size = 35, face = "bold"),
    #legend.text = element_text(size = 25),
    legend.position = "none"
    #legend.key.size = unit(1.5, "cm"),
    #legend.spacing.y = unit(0.5, "cm")
  );chla_dot
```

## Protein

Load data 
```{r}
prot <- read.csv("../output/Physiology/SolubleProtein/SolubleProtein.csv")

# Remove field data for now 
prot_exp <- prot %>%
  filter(!Treatment == "Field")
```

Remove outliers 
```{r}
# Look for outliers visually
ggplot(prot_exp, aes(x = Treatment, y = prot_mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# Calculate IQR
Q1 <- quantile(prot_exp$prot_mg.cm2, 0.25)
Q3 <- quantile(prot_exp$prot_mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
prot_exp_filtered <- prot_exp[prot_exp$prot_mg.cm2 >= lower_bound & prot_exp$prot_mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(prot_exp)
dim(prot_exp_filtered)
```

Format data for plotting 
```{r}
# Assign month to TPs 
prot_exp_filtered <- prot_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

# Update to Acclimation for plotting purposes 
prot_exp_filtered <- prot_exp_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Calculate mean, sd and sem of each group for plotting 
```{r}
prot_summary <- prot_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_prot = mean(prot_mg.cm2, na.rm = TRUE),
    sd_prot = sd(prot_mg.cm2, na.rm = TRUE),
    sem_prot = sd_prot / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot
```{r}
# Set Month as factor 
prot_summary$Month <- factor(prot_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
prot_dot <- ggplot(prot_summary, aes(x = Month, y = mean_prot, color = Treatment)) +
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_prot - sem_prot, 
                    ymax = mean_prot + sem_prot),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  annotate("text", x = 4, y = -Inf, label = "Time point p<0.001", 
         hjust = 1.1, vjust = -1.1, size = 8, fontface = "bold") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab("Soluble Protein (mg/cm²)") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, color = "black"),
    axis.title.y = element_text(size = 35, face = "bold", margin = margin(r = 10)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    #legend.title = element_text(size = 35, face = "bold"),
    #legend.text = element_text(size = 25),
    legend.position = "none"
    #legend.key.size = unit(1.5, "cm"),
    #legend.spacing.y = unit(0.5, "cm")
  );prot_dot
```

## Carbs 

Read in data
```{r}
carb <- read.csv("../output/Physiology/Carbohydrates/Carbohydrates.csv")
```

Remove outliers based on IQR
```{r}
# Look for outliers visually
ggplot(carb, aes(x = Treatment, y = Carb.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# Calculate IQR
Q1 <- quantile(carb$Carb.mg.cm2, 0.25)
Q3 <- quantile(carb$Carb.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
carb_exp_filtered <- carb[carb$Carb.mg.cm2 >= lower_bound & carb$Carb.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(carb)
dim(carb_exp_filtered)
```

Format data for plotting 
```{r}
# Assign month to TPs 
carb_exp_filtered <- carb_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

# Update to Acclimation for plotting purposes 
carb_exp_filtered <- carb_exp_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Calculate mean, sd and sem of each group for plotting 
```{r}
carb_summary <- carb_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_carb = mean(Carb.mg.cm2, na.rm = TRUE),
    sd_carb = sd(Carb.mg.cm2, na.rm = TRUE),
    sem_carb = sd_carb / sqrt(n),  # SEM formula
    sem_carb_offset = mean_carb + sem_carb + 0.2,
    .groups = "drop"
  )
```

Plot
```{r}
# Set Month as factor 
carb_summary$Month <- factor(carb_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
carb_dot <- ggplot(carb_summary, aes(x = Month, y = mean_carb, color = Treatment)) +
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_carb - sem_carb, 
                    ymax = mean_carb + sem_carb),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  annotate("text", x = 4, y = -Inf, label = "Treatment p=0.043", 
         hjust = 1.1, vjust = -1.1, size = 8, fontface = "bold") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab("Carbohydrates (mg/cm²)") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, color = "black"),
    axis.title.y = element_text(size = 35, face = "bold", margin = margin(r = 10)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    #legend.title = element_text(size = 35, face = "bold"),
    #legend.text = element_text(size = 25),
    legend.position = "none"
    #legend.key.size = unit(1.5, "cm"),
    #legend.spacing.y = unit(0.5, "cm")
  ); carb_dot
```

## AFDW 

Read in data and subset host 
```{r}
biomass <- read.csv("../output/Physiology/AFDW/AFDW_biomass.csv")

biomass_host <- biomass %>%
  #dplyr::mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", as.character(Treatment))) %>%
  dplyr::mutate(Treatment = as.factor(Treatment)) %>%
  dplyr::filter(Fraction == "Host") %>%
  na.omit()
```

Remove outliers based on IQR
```{r}
# Look for outliers visually
ggplot(biomass_host, aes(x = Treatment, y = AFDW.mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# Calculate IQR
Q1 <- quantile(biomass_host$AFDW.mg.cm2, 0.25)
Q3 <- quantile(biomass_host$AFDW.mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
biomass_host_filtered <- biomass_host[biomass_host$AFDW.mg.cm2 >= lower_bound & biomass_host$AFDW.mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(biomass_host)
dim(biomass_host_filtered)
```

Format data for plotting 
```{r}
# Assign month to TPs 
biomass_host_filtered <- biomass_host_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 

# Update to Acclimation for plotting purposes 
biomass_host_filtered <- biomass_host_filtered %>%
  mutate(Treatment = ifelse(Month == "February", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

Calculate mean, sd and sem of each group for plotting 
```{r}
biomass_host_summary <- biomass_host_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_host = mean(AFDW.mg.cm2, na.rm = TRUE),
    sd_host = sd(AFDW.mg.cm2, na.rm = TRUE),
    sem_host = sd_host / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot
```{r}
# Set Month as factor 
biomass_host_summary$Month <- factor(biomass_host_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
host_dot <- ggplot(biomass_host_summary, aes(x = Month, y = mean_host, color = Treatment)) +
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_host - sem_host, 
                    ymax = mean_host + sem_host),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  annotate("text", x = Inf, y = -Inf, label = "ns", 
         hjust = 1.1, vjust = -1.1, size = 8, fontface = "bold") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab("Host Tissue Biomass (mg/cm²)") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, color = "black"),
    axis.title.y = element_text(size = 35, face = "bold", margin = margin(r = 10)),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    # legend.title = element_text(size = 35, face = "bold"),
    # legend.text = element_text(size = 25),
    legend.position = "none"
    # legend.key.size = unit(1.5, "cm"),
    # legend.spacing.y = unit(0.5, "cm")
  ); host_dot
```

## Create multi-panel phys figure and save 
```{r}
# 1. Combine the plots
phys_final <- ((chla_dot | photo_dot | resp_dot) / 
                (prot_dot | carb_dot | host_dot)) 

# 2. Apply global formatting
phys_final <- phys_final +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  # Use & to force a single legend and color mapping
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8", 
                               "Heat" = "#E41A1C"),
                     limits = c("Acclimation", "Ambient", "Heat")) &
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C"),
                    limits = c("Acclimation", "Ambient", "Heat")) &
  theme(
    legend.position = "right", 
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 30, face = "plain"), 
    legend.key.size = unit(2.5, "cm"),
    plot.tag = element_text(size = 50, face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
  )

# 3. Save with adjusted dimensions
ggsave("../output/Physiology/Physiology_panels.png", plot = phys_final, 
       width = 40, height = 20, units = "in", dpi = 300)
ggsave("../output/Physiology/Physiology_panels.pdf", plot = phys_final, 
       width = 40, height = 20, units = "in", dpi = 300)
```



