---
title: "Soluble Protein"
author: "jillashey"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")
# load packages
library(tidyverse)
library(broom)

library(dbplyr)
library(tidyverse)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(lme4)
library(lmerTest)

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("directlabels")) install.packages("directlabels")
# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(directlabels)
library(multcomp)
```


## Import data 
```{r}
## There is definitely a better way to do this

# Plate 1
p1_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_data.csv")
p1_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_platemap.csv")
p1 <- left_join(p1_data, p1_platemap)

# Plate 2
p2_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate2_data.csv")
p2_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_platemap.csv")
p2 <- left_join(p2_data, p2_platemap)

# Plate 3
p3_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate3_data.csv")
p3_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate3_platemap.csv")
p3 <- left_join(p3_data, p3_platemap)

# Plate 4
p4_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate4_data.csv")
p4_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate4_platemap.csv")
p4 <- left_join(p4_data, p4_platemap)

# Plate 5
p5_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate5_data.csv")
p5_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate5_platemap.csv")
p5 <- left_join(p5_data, p5_platemap)

```

## Subtract blank means for each run (Standard I is the blank)
```{r}
p1.blank <- p1 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p1$abs.corr <- p1$X562 - p1.blank$blk.avg

p2.blank <- p2 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p2$abs.corr <- p2$X562 - p2.blank$blk.avg

p3.blank <- p3 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p3$abs.corr <- p3$X562 - p3.blank$blk.avg

p4.blank <- p4 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p4$abs.corr <- p4$X562 - p4.blank$blk.avg

p5.blank <- p5 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p5$abs.corr <- p5$X562 - p5.blank$blk.avg
```

## Plot standard curve 
```{r}
# Select only standards and plot 
## Plate 1
p1.standards <- p1[str_detect(p1$ID, "Standard"),]
p1.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p1.standards.plot <- ggplot(data = p1.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1.standards.plot

## Plate 2
p2.standards <- p2[str_detect(p2$ID, "Standard"),]
p2.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p2.standards.plot <- ggplot(data = p2.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p2.standards.plot

## Plate 3
p3.standards <- p3[str_detect(p3$ID, "Standard"),]
p3.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p3.standards.plot <- ggplot(data = p3.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p3.standards.plot

## Plate 4
p4.standards <- p4[str_detect(p4$ID, "Standard"),]
p4.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p4.standards.plot <- ggplot(data = p4.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p4.standards.plot

## Plate 5
p5.standards <- p5[str_detect(p5$ID, "Standard"),] %>% na.omit()
p5.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p5.standards.plot <- ggplot(data = p5.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p5.standards.plot

## reminder - concentration units is (ug/mL)
```

## Create model for standards data 
```{r}
p1.standards.lm <- lm(Concentration ~ abs.corr, data = p1.standards)
summary(p1.standards.lm)

p2.standards.lm <- lm(Concentration ~ abs.corr, data = p2.standards)
summary(p2.standards.lm)

p3.standards.lm <- lm(Concentration ~ abs.corr, data = p3.standards)
summary(p3.standards.lm)

p4.standards.lm <- lm(Concentration ~ abs.corr, data = p4.standards)
summary(p4.standards.lm)

p5.standards.lm <- lm(Concentration ~ abs.corr, data = p5.standards)
summary(p5.standards.lm)
```

## Use model to find concentration of samples 
```{r}
p1.samples <- p1[!str_detect(p1$ID, "Standard"),]
p1.samples$Concentration <- predict(p1.standards.lm, newdata = p1.samples)

p2.samples <- p2[!str_detect(p2$ID, "Standard"),]
p2.samples$Concentration <- predict(p2.standards.lm, newdata = p2.samples)

p3.samples <- p3[!str_detect(p3$ID, "Standard"),]
p3.samples$Concentration <- predict(p3.standards.lm, newdata = p3.samples)

p4.samples <- p4[!str_detect(p4$ID, "Standard"),]
p4.samples$Concentration <- predict(p4.standards.lm, newdata = p4.samples)

p5.samples <- p5[!str_detect(p5$ID, "Standard"),]
p5.samples$Concentration <- predict(p5.standards.lm, newdata = p5.samples)
```

## Combine all datasets 
```{r}
p.all <- rbind(p1.samples, p2.samples, p3.samples, p4.samples, p5.samples)

p.all <- p.all %>%
  #select(ID, Concentration) %>%
  dplyr::rename(PlugID = ID) %>%
  na.omit() %>%
  dplyr::group_by(PlugID) %>%
  dplyr::summarise(mean_concentration = mean(Concentration))
```

## Read in surface area, homogenate volume and metadata 
```{r}
surface_area <- read.csv("data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("data/MasterFragmentSheet.csv", header = T, na.strings = "") %>% 
  dplyr::filter(Metric == "P/R, molec") %>% 
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

# Merge dataframes
metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
#metadata$vol_mL <- gsub("NA", "", metadata$vol_mL)
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge protein data w/ metadata and normalize data 
```{r}
prot <- left_join(metadata, p.all) %>%
  mutate(prot_ug = mean_concentration * vol_mL,
         prot_ug.cm2 = prot_ug / Area_cm,
         prot_mg.cm2 = prot_ug.cm2 / 1000) %>% 
  na.omit() # omitting NAs for now 

## The FLD TP2 corals do not have surface area (I don't think I have pictures of them. Will need to get their surface area via wax dipping)
```

## Summarize 
```{r}
mean.prot <- summarySE(prot, measurevar = "prot_mg.cm2", groupvars = c("Treatment", "Timepoint"))

mean.prot_site <- summarySE(prot, measurevar = "prot_mg.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
# I did not equally sample across site, which was a mistake on my part. I need to figure out how to present the data with site 
```

## Save data 
```{r}
write.csv(prot, "../output/Physiology/SolubleProtein/SolubleProtein.csv")
```

## Plot 
```{r}
gd <- prot %>% 
  dplyr::group_by(Timepoint, Treatment) %>%
  dplyr::summarise(avg = mean(prot_mg.cm2),
            sem = sd(prot_mg.cm2)/sqrt(length(prot_mg.cm2)))

plot <- ggplot(prot, aes(x = Timepoint, y = prot_mg.cm2, group = Treatment)) +
  geom_point(aes(color = Treatment), size = 2) +
  geom_point(data = gd, aes(x = Timepoint, y = avg-sem, group=Treatment, colour = Treatment), size=2, shape=8)+
  geom_point(data = gd, aes(x = Timepoint, y = avg+sem, group=Treatment, colour = Treatment), size=2, shape=8)+
  geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_color_manual(values = c("blue", "gray", "red")) +
  theme_classic()
plot

## Right now, I'm only summarizing and plotting by Timepoint and Treatment. At some point, I will include site as a random effect variable 

# Remove outliers 
prot_sub <- prot %>%
  filter(prot_mg.cm2 <= 10)

ggplot(prot_sub, aes(x = Timepoint, y = prot_mg.cm2, group = Treatment)) +
  geom_point(aes(color = Treatment), size = 2) +
  geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_color_manual(values = c("blue", "gray", "red")) +
  theme_classic()

ggplot(prot_sub, aes(x = Timepoint, y = prot_mg.cm2, fill = Treatment)) +
  geom_boxplot() +
  #geom_point(aes(color = Treatment), size = 2) +
  #geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_fill_manual(values = c("blue", "gray", "red")) + 
  #facet_wrap(~Timepoint) +
  #ylab("Soluble protein (mg/cm²)") +
  theme_classic()






# Remove field data and outlier
prot_exp <- prot %>%
  filter(!Treatment == "Field") %>%
  filter(prot_mg.cm2 <= 10) 

# plot by exp only 
plot_trt <- ggplot(prot_exp, aes(x = Timepoint, y = prot_mg.cm2, fill = Treatment)) +
  geom_boxplot(lwd = 1, color = "black") +
  #geom_point(aes(color = Treatment), size = 2) +
  #geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_fill_manual(values = c("blue", "red")) +
  ylab("Soluble protein (mg/cm²)") +
  xlab("") +
  theme_classic() +
  theme(legend.title = element_text(size = 28),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))) 
plot_trt

ggsave("output/Physiology/SolubleProtein/SolubleProtein_Treatment.pdf", plot_trt, width = 50, height = 25, units = "cm")
ggsave("output/Physiology/SolubleProtein/SolubleProtein_Treatment.png", plot_trt, width = 50, height = 25, units = "cm")
```

## Stats 
```{r}
## Summarize w/o field data
mean.prot_exp <- summarySE(prot_exp, measurevar = "prot_mg.cm2", groupvars = c("Treatment", "Timepoint"))
mean.prot_site_exp <- summarySE(prot_exp, measurevar = "prot_mg.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
```

## Test for differences 
For now (as of 20230420), I'm using code based on this [tutorial](http://www.sthda.com/english/wiki/two-way-anova-test-in-r). I need to do some more investigations on stats stuff 
```{r}
## Compute two-way ANOVA test 
model <- aov(log(prot_mg.cm2) ~ Treatment*Timepoint, data = prot_exp)
summary(model)

#                     Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment            1   0.15  0.1522   0.314    0.577    
# Timepoint            7  21.91  3.1294   6.464 5.36e-06 ***
# Treatment:Timepoint  7   5.81  0.8300   1.714    0.118    
# Residuals           77  37.28  0.4841                     
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

## Compute summary stats 
model.tables(model, type = "means", se = T)


## Multiple pairwise-comparisons between the means of groups 
# Tukey HSD 
TukeyHSD(model, which = "Timepoint")
## TP6 differs from TP0, TP1, TP2, TP3, TP4, TP5, and TP7

# Multiple comparisons using multcomp package

# Pairwise test 
pairwise.t.test(log(prot_exp$prot_mg.cm2), prot_exp$Timepoint, p.adjust.method = "bonf") # should I specify a p.adjust.method? 


## Check ANOVA assumptions 
# ANOVA assumes data is normally distributed and variance across groups is homogenous 
# 1. Check homogeneity of variance assumption
# a. Residuals vs fits plot
plot(model, 1)

# b. Levene's test 
leveneTest(prot_mg.cm2 ~ Timepoint*Treatment, data = prot_exp)
## p-value is  less than 0.05, so variance is not homogenous 

# 2. Check normality assumption 
# a. Normality plot of residuals 
plot(model, 2) ## Similar to the other residual plot, looks like rows 17, 73, and 107 are outliers. May be good to remove 
hist(residuals(model))

#b. Shapiro-Wilk test 
aov_residuals <- residuals(object = model)
shapiro.test(x = aov_residuals)
## p-value is not less than 0.05, so normality is all good 
```



### 20240108

Going to remove outliers, plot everything and do stats 

```{r}
prot <- read_csv("../output/Physiology/SolubleProtein/SolubleProtein.csv")
```

Assign Month to TPs
```{r}
prot <- prot %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "Febuary", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Remove field data and change Treatment for TP0 to be "Acclimation"
```{r}
prot <- prot %>%
#   filter(!Timepoint == "TP0") %>%
   filter(!Treatment == "Field") %>%
   mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", Treatment))
```

Set levels for plotting by month 
```{r}
prot$Month <- factor(prot$Month, levels = c("Febuary", "March", "Early April", "Late April", "May", "June", "July", "August"))
```

Plot 
```{r}
plot <- ggplot(prot, aes(x = Month, y = prot_mg.cm2, colour = Treatment)) +
  geom_boxplot(aes(colour=Treatment), lwd=2) +
  #geom_point(aes(fill=Treatment, group=Treatment), colour="black", pch = 21, size=4, position = position_jitterdodge(0.3), alpha=0.9) + 
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) +
  xlab("") +
  ylab("Soluble protein (mg/cm²)") +
  #hjust(x = 0) +
  #facet_wrap(~Assay, ncol = 2) +
  theme_classic() +
    theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); plot 

ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_20240521.pdf", plot, width = 15, height = 10)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_20240521.png", plot, width = 15, height = 10)
```

Looking at this plot, I see that not all of the treatments/timepoints have n=8. I need to look through my data and notebooks so that I can make sure I have all the data that I need and that I have processed everything. 

Going to do stats just for Temperate coral conference 
```{r}
model <- aov(log(prot_mg.cm2) ~ Treatment*Timepoint, data = prot)
summary(model)

#                     Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment            2   2.21   1.105   1.996    0.143    
# Timepoint            6  21.19   3.532   6.380 1.65e-05 ***
# Treatment:Timepoint  6   3.55   0.592   1.069    0.388    
# Residuals           79  43.74   0.554                     
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

## Compute summary stats 
model.tables(model, type = "means", se = T)

## Multiple pairwise-comparisons between the means of groups 
# Tukey HSD 
TukeyHSD(model, which = "Timepoint")
## TP6 differs from TP0, TP1, TP2, TP3, TP4, and TP5
```

STILL NEED TO REMOVE OUTLIERS AND CONFIRM THAT I HAVE ALL SAMPLES 














Summarize data 
```{r}
summary <- prot %>%
  dplyr::select(Timepoint, Treatment, prot_mg.cm2)%>%  # Select the necessary columns
  group_by(Timepoint, Treatment) %>%
  summarise(mean_prot_mg.cm2 = mean(prot_mg.cm2, na.rm = TRUE),
            sd_prot_mg.cm2 = sd(prot_mg.cm2, na.rm = TRUE))

summary <- prot %>% 
  group_by(Timepoint, Treatment) %>%
  dplyr::select(prot_mg.cm2) %>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

summary_tp <- prot %>% 
  group_by(Timepoint) %>%
  select(prot_mg.cm2) %>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary_tp
```

Stats!
```{r}
## Compute two-way ANOVA test 
model <- aov(prot_mg.cm2 ~ Treatment*Timepoint, data = prot)
summary(model)

## TP significant, Trt:TP different significant

## Compute summary stats 
model.tables(model, type = "means", se = T)

## Multiple pairwise-comparisons between the means of groups 
# Tukey HSD 
TukeyHSD(model, which = "Timepoint")

emm = emmeans(model, ~ Treatment*Timepoint)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

## Check ANOVA assumptions 
# ANOVA assumes data is normally distributed and variance across groups is homogenous 
# 1. Check homogeneity of variance assumption
# a. Residuals vs fits plot
plot(model, 1)

# b. Levene's test 
leveneTest(residuals(model) ~ Timepoint*Treatment, data = prot)
## p-value is  less than 0.05, so variance is not homogenous 

# 2. Check normality assumption 
# a. Normality plot of residuals 
plot(model, 2) 
hist(residuals(model))

#b. Shapiro-Wilk test 
aov_residuals <- residuals(object = model)
shapiro.test(x = aov_residuals)
# pvalue is less than 0.05, normality is violated





## Data transformation using log 
## Compute two-way ANOVA test 
model <- aov(log(prot_mg.cm2) ~ Treatment*Timepoint, data = prot)
summary(model)

## TP significant, Trt:TP different significant

## Compute summary stats 
model.tables(model, type = "means", se = T)

## Multiple pairwise-comparisons between the means of groups 
# Tukey HSD 
TukeyHSD(model, which = "Timepoint")

emm = emmeans(model, ~ Treatment*Timepoint)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

## Check ANOVA assumptions 
# ANOVA assumes data is normally distributed and variance across groups is homogenous 
# 1. Check homogeneity of variance assumption
# a. Residuals vs fits plot
plot(model, 1)

# b. Levene's test 
leveneTest(residuals(model) ~ Timepoint*Treatment, data = prot)
## p-value is  greater than 0.05, so variance is  homogenous 

# 2. Check normality assumption 
# a. Normality plot of residuals 
plot(model, 2) 
hist(residuals(model))

#b. Shapiro-Wilk test 
aov_residuals <- residuals(object = model)
shapiro.test(x = aov_residuals)
# pvalue is greater than 0.05, normality is not violated
```

Remove outliers 
```{r}
#set quantile values
q <- c(0.25, 0.75)

# calculate quantile values by Temperature and Treatment groups
Quants <- prot %>%
  group_by(Treatment, Timepoint) %>%
  summarize(quant25 = quantile(prot, probs = q[1]),
            quant75 = quantile(prot, probs = q[2]),
            IQRbyGroup=IQR(prot))

## getting errors...
```







I also need to go back and see which AST samples I still need to run for carb and protein in the lab
