---
title: "Soluble Protein"
author: "jillashey"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")
# load packages
library(tidyverse)
library(broom)

library(dbplyr)
library(tidyverse)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(lme4)
library(lmerTest)

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("directlabels")) install.packages("directlabels")
# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(directlabels)
```


## Import data 
```{r}
## There is definitely a better way to do this

# Plate 1
p1_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_data.csv")
p1_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_platemap.csv")
p1 <- left_join(p1_data, p1_platemap)

# Plate 2
p2_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate2_data.csv")
p2_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_platemap.csv")
p2 <- left_join(p2_data, p2_platemap)

# Plate 3
p3_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate3_data.csv")
p3_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate3_platemap.csv")
p3 <- left_join(p3_data, p3_platemap)

# Plate 4
p4_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate4_data.csv")
p4_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate4_platemap.csv")
p4 <- left_join(p4_data, p4_platemap)

# Plate 5
p5_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate5_data.csv")
p5_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate5_platemap.csv")
p5 <- left_join(p5_data, p5_platemap)

```

## Subtract blank means for each run (Standard I is the blank)
```{r}
p1.blank <- p1 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p1$abs.corr <- p1$X562 - p1.blank$blk.avg

p2.blank <- p2 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p2$abs.corr <- p2$X562 - p2.blank$blk.avg

p3.blank <- p3 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p3$abs.corr <- p3$X562 - p3.blank$blk.avg

p4.blank <- p4 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p4$abs.corr <- p4$X562 - p4.blank$blk.avg

p5.blank <- p5 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p5$abs.corr <- p5$X562 - p5.blank$blk.avg
```

## Plot standard curve 
```{r}
# Select only standards and plot 
## Plate 1
p1.standards <- p1[str_detect(p1$ID, "Standard"),]
p1.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p1.standards.plot <- ggplot(data = p1.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1.standards.plot

## Plate 2
p2.standards <- p2[str_detect(p2$ID, "Standard"),]
p2.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p2.standards.plot <- ggplot(data = p2.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p2.standards.plot

## Plate 3
p3.standards <- p3[str_detect(p3$ID, "Standard"),]
p3.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p3.standards.plot <- ggplot(data = p3.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p3.standards.plot

## Plate 4
p4.standards <- p4[str_detect(p4$ID, "Standard"),]
p4.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p4.standards.plot <- ggplot(data = p4.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p4.standards.plot

## Plate 5
p5.standards <- p5[str_detect(p5$ID, "Standard"),] %>% na.omit()
p5.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p5.standards.plot <- ggplot(data = p5.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p5.standards.plot

## reminder - concentration units is (ug/mL)
```

## Create model for standards data 
```{r}
p1.standards.lm <- lm(Concentration ~ abs.corr, data = p1.standards)
summary(p1.standards.lm)

p2.standards.lm <- lm(Concentration ~ abs.corr, data = p2.standards)
summary(p2.standards.lm)

p3.standards.lm <- lm(Concentration ~ abs.corr, data = p3.standards)
summary(p3.standards.lm)

p4.standards.lm <- lm(Concentration ~ abs.corr, data = p4.standards)
summary(p4.standards.lm)

p5.standards.lm <- lm(Concentration ~ abs.corr, data = p5.standards)
summary(p5.standards.lm)
```

## Use model to find concentration of samples 
```{r}
p1.samples <- p1[!str_detect(p1$ID, "Standard"),]
p1.samples$Concentration <- predict(p1.standards.lm, newdata = p1.samples)

p2.samples <- p2[!str_detect(p2$ID, "Standard"),]
p2.samples$Concentration <- predict(p2.standards.lm, newdata = p2.samples)

p3.samples <- p3[!str_detect(p3$ID, "Standard"),]
p3.samples$Concentration <- predict(p3.standards.lm, newdata = p3.samples)

p4.samples <- p4[!str_detect(p4$ID, "Standard"),]
p4.samples$Concentration <- predict(p4.standards.lm, newdata = p4.samples)

p5.samples <- p5[!str_detect(p5$ID, "Standard"),]
p5.samples$Concentration <- predict(p5.standards.lm, newdata = p5.samples)
```

## Combine all datasets 
```{r}
p.all <- rbind(p1.samples, p2.samples, p3.samples, p4.samples, p5.samples)

p.all <- p.all %>%
  #select(ID, Concentration) %>%
  dplyr::rename(PlugID = ID) %>%
  na.omit() %>%
  dplyr::group_by(PlugID) %>%
  dplyr::summarise(mean_concentration = mean(Concentration))
```

## Read in surface area, homogenate volume and metadata 
```{r}
surface_area <- read.csv("data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("data/MasterFragmentSheet.csv", header = T, na.strings = "") %>% 
  dplyr::filter(Metric == "P/R, molec") %>% 
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

# Merge dataframes
metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
#metadata$vol_mL <- gsub("NA", "", metadata$vol_mL)
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge protein data w/ metadata and normalize data 
```{r}
prot <- left_join(metadata, p.all) %>%
  mutate(prot_ug = mean_concentration * vol_mL,
         prot_ug.cm2 = prot_ug / Area_cm,
         prot_mg.cm2 = prot_ug.cm2 / 1000) %>% 
  na.omit() # omitting NAs for now 

## The FLD TP2 corals do not have surface area (I don't think I have pictures of them. Will need to get their surface area via wax dipping)
```

## Summarize 
```{r}
mean.prot <- summarySE(prot, measurevar = "prot_mg.cm2", groupvars = c("Treatment", "Timepoint"))

mean.prot_site <- summarySE(prot, measurevar = "prot_mg.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
# I did not equally sample across site, which was a mistake on my part. I need to figure out how to present the data with site 
```

## Save data 
```{r}
write.csv(prot, "output/Physiology/SolubleProtein/SolubleProtein.csv")
```

## Plot 
```{r}
gd <- prot %>% 
  dplyr::group_by(Timepoint, Treatment) %>%
  dplyr::summarise(avg = mean(prot_mg.cm2),
            sem = sd(prot_mg.cm2)/sqrt(length(prot_mg.cm2)))

plot <- ggplot(prot, aes(x = Timepoint, y = prot_mg.cm2, group = Treatment)) +
  geom_point(aes(color = Treatment), size = 2) +
  geom_point(data = gd, aes(x = Timepoint, y = avg-sem, group=Treatment, colour = Treatment), size=2, shape=8)+
  geom_point(data = gd, aes(x = Timepoint, y = avg+sem, group=Treatment, colour = Treatment), size=2, shape=8)+
  geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_color_manual(values = c("blue", "gray", "red")) +
  theme_classic()
plot

## Right now, I'm only summarizing and plotting by Timepoint and Treatment. At some point, I will include site as a random effect variable 

# Remove outliers 
prot_sub <- prot %>%
  filter(prot_mg.cm2 <= 10)

ggplot(prot_sub, aes(x = Timepoint, y = prot_mg.cm2, group = Treatment)) +
  geom_point(aes(color = Treatment), size = 2) +
  geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_color_manual(values = c("blue", "gray", "red")) +
  theme_classic()

ggplot(prot_sub, aes(x = Timepoint, y = prot_mg.cm2, fill = Treatment)) +
  geom_boxplot() +
  #geom_point(aes(color = Treatment), size = 2) +
  #geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_fill_manual(values = c("blue", "gray", "red")) + 
  #facet_wrap(~Timepoint) +
  #ylab("Soluble protein (mg/cm²)") +
  theme_classic()






# Remove field data and outlier
prot_exp <- prot %>%
  filter(!Treatment == "Field") %>%
  filter(prot_mg.cm2 <= 10) 

# plot by exp only 
plot_trt <- ggplot(prot_exp, aes(x = Timepoint, y = prot_mg.cm2, fill = Treatment)) +
  geom_boxplot(lwd = 1, color = "black") +
  #geom_point(aes(color = Treatment), size = 2) +
  #geom_line(data = gd, aes(x = Timepoint, y = avg, group=Treatment, colour = Treatment)) +
  scale_fill_manual(values = c("blue", "red")) +
  ylab("Soluble protein (mg/cm²)") +
  xlab("") +
  theme_classic() +
  theme(legend.title = element_text(size = 28),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))) 
plot_trt

ggsave("output/Physiology/SolubleProtein/SolubleProtein_Treatment.pdf", plot_trt, width = 50, height = 25, units = "cm")
ggsave("output/Physiology/SolubleProtein/SolubleProtein_Treatment.png", plot_trt, width = 50, height = 25, units = "cm")
```

## Stats 
```{r}
## Summarize w/o field data
mean.prot_exp <- summarySE(prot_exp, measurevar = "prot_mg.cm2", groupvars = c("Treatment", "Timepoint"))
mean.prot_site_exp <- summarySE(prot_exp, measurevar = "prot_mg.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
```

## Test for differences 
For now (as of 20230420), I'm using code based on this [tutorial](http://www.sthda.com/english/wiki/two-way-anova-test-in-r). I need to do some more investigations on stats stuff 
```{r}
## Compute two-way ANOVA test 
model <- aov(log(prot_mg.cm2) ~ Treatment*Timepoint, data = prot_exp)
summary(model)

#                     Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment            1   0.15  0.1522   0.314    0.577    
# Timepoint            7  21.91  3.1294   6.464 5.36e-06 ***
# Treatment:Timepoint  7   5.81  0.8300   1.714    0.118    
# Residuals           77  37.28  0.4841                     
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

## Compute summary stats 
model.tables(model, type = "means", se = T)


## Multiple pairwise-comparisons between the means of groups 
# Tukey HSD 
TukeyHSD(model, which = "Timepoint")
## TP6 differs from TP0, TP1, TP2, TP3, TP4, TP5, and TP7

# Multiple comparisons using multcomp package

# Pairwise test 
pairwise.t.test(log(prot_exp$prot_mg.cm2), prot_exp$Timepoint, p.adjust.method = "bonf") # should I specify a p.adjust.method? 


## Check ANOVA assumptions 
# ANOVA assumes data is normally distributed and variance across groups is homogenous 
# 1. Check homogeneity of variance assumption
# a. Residuals vs fits plot
plot(model, 1)
# 1, 3, 88 may be outliers 

# b. Levene's test 
leveneTest(prot_mg.cm2 ~ Timepoint*Treatment, data = prot_exp)
## p-value is  less than 0.05, so variance is not homogenous 

# 2. Check normality assumption 
# a. Normality plot of residuals 
plot(model, 2) ## Similar to the other residual plot, looks like rows 17, 73, and 107 are outliers. May be good to remove 
hist(residuals(model))

#b. Shapiro-Wilk test 
aov_residuals <- residuals(object = model)
shapiro.test(x = aov_residuals)
## p-value is not less than 0.05, so normality is all good 
```

