---
title: "Soluble Protein"
author: "jillashey"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")
# load packages
library(tidyverse)
library(broom)

library(dbplyr)
library(tidyverse)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(lme4)
library(lmerTest)

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("directlabels")) install.packages("directlabels")
# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(directlabels)
library(multcomp)
```


## Import data 
```{r}
## There is definitely a better way to do this

# Plate 1
p1_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_data.csv")
p1_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_platemap.csv")
p1 <- left_join(p1_data, p1_platemap)

# Plate 2
p2_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate2_data.csv")
p2_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate1_platemap.csv")
p2 <- left_join(p2_data, p2_platemap)

# Plate 3
p3_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate3_data.csv")
p3_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate3_platemap.csv")
p3 <- left_join(p3_data, p3_platemap)

# Plate 4
p4_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate4_data.csv")
p4_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate4_platemap.csv")
p4 <- left_join(p4_data, p4_platemap)

# Plate 5
p5_data <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate5_data.csv")
p5_platemap <- read.csv("data/Physiology/Protein/SP/20230327_SP_plate5_platemap.csv")
p5 <- left_join(p5_data, p5_platemap)

```

## Subtract blank means for each run (Standard I is the blank)
```{r}
p1.blank <- p1 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p1$abs.corr <- p1$X562 - p1.blank$blk.avg

p2.blank <- p2 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p2$abs.corr <- p2$X562 - p2.blank$blk.avg

p3.blank <- p3 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p3$abs.corr <- p3$X562 - p3.blank$blk.avg

p4.blank <- p4 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p4$abs.corr <- p4$X562 - p4.blank$blk.avg

p5.blank <- p5 %>%
  filter(ID == "StandardI") %>%
  summarise(blk.avg = mean(X562))
p5$abs.corr <- p5$X562 - p5.blank$blk.avg
```

## Plot standard curve 
```{r}
# Select only standards and plot 
## Plate 1
p1.standards <- p1[str_detect(p1$ID, "Standard"),]
p1.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p1.standards.plot <- ggplot(data = p1.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1.standards.plot

## Plate 2
p2.standards <- p2[str_detect(p2$ID, "Standard"),]
p2.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p2.standards.plot <- ggplot(data = p2.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p2.standards.plot

## Plate 3
p3.standards <- p3[str_detect(p3$ID, "Standard"),]
p3.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p3.standards.plot <- ggplot(data = p3.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p3.standards.plot

## Plate 4
p4.standards <- p4[str_detect(p4$ID, "Standard"),]
p4.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p4.standards.plot <- ggplot(data = p4.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p4.standards.plot

## Plate 5
p5.standards <- p5[str_detect(p5$ID, "Standard"),] %>% na.omit()
p5.standards$Concentration <- c(2000,2000, 1500, 1500, 1000, 1000, 750, 750, 500, 500, 250, 250, 125, 125, 25, 25, 0, 0) # making column for concentrations
p5.standards.plot <- ggplot(data = p5.standards, aes(x = Concentration, y = abs.corr)) +
  ylab("Absorbance (nm)") +
  xlab("Concentration (ug/mL)") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p5.standards.plot

## reminder - concentration units is (ug/mL)
```

## Create model for standards data 
```{r}
p1.standards.lm <- lm(Concentration ~ abs.corr, data = p1.standards)
summary(p1.standards.lm)

p2.standards.lm <- lm(Concentration ~ abs.corr, data = p2.standards)
summary(p2.standards.lm)

p3.standards.lm <- lm(Concentration ~ abs.corr, data = p3.standards)
summary(p3.standards.lm)

p4.standards.lm <- lm(Concentration ~ abs.corr, data = p4.standards)
summary(p4.standards.lm)

p5.standards.lm <- lm(Concentration ~ abs.corr, data = p5.standards)
summary(p5.standards.lm)
```

## Use model to find concentration of samples 
```{r}
p1.samples <- p1[!str_detect(p1$ID, "Standard"),]
p1.samples$Concentration <- predict(p1.standards.lm, newdata = p1.samples)

p2.samples <- p2[!str_detect(p2$ID, "Standard"),]
p2.samples$Concentration <- predict(p2.standards.lm, newdata = p2.samples)

p3.samples <- p3[!str_detect(p3$ID, "Standard"),]
p3.samples$Concentration <- predict(p3.standards.lm, newdata = p3.samples)

p4.samples <- p4[!str_detect(p4$ID, "Standard"),]
p4.samples$Concentration <- predict(p4.standards.lm, newdata = p4.samples)

p5.samples <- p5[!str_detect(p5$ID, "Standard"),]
p5.samples$Concentration <- predict(p5.standards.lm, newdata = p5.samples)
```

## Combine all datasets 
```{r}
p.all <- rbind(p1.samples, p2.samples, p3.samples, p4.samples, p5.samples)

p.all <- p.all %>%
  #select(ID, Concentration) %>%
  dplyr::rename(PlugID = ID) %>%
  na.omit() %>%
  dplyr::group_by(PlugID) %>%
  dplyr::summarise(mean_concentration = mean(Concentration))
```

## Read in surface area, homogenate volume and metadata 
```{r}
surface_area <- read.csv("data/Physiology/SurfaceArea.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, Area_cm)

homogenate_vol <- read.csv("data/Physiology/homogenate_volume.csv", header = T, na.strings = "") %>% dplyr::select(PlugID, vol_mL) 

master <- read.csv("data/MasterFragmentSheet.csv", header = T, na.strings = "") %>% 
  dplyr::filter(Metric == "P/R, molec") %>% 
  dplyr::select(PlugID, Site, Treatment, Timepoint, ExperimentalTank)

# Merge dataframes
metadata <- join_all(list(surface_area, homogenate_vol, master), by = "PlugID", type = "left")
#metadata$vol_mL <- gsub("NA", "", metadata$vol_mL)
metadata$vol_mL <- as.numeric(metadata$vol_mL)
```

## Merge protein data w/ metadata and normalize data 
```{r}
prot <- left_join(metadata, p.all) %>%
  mutate(prot_ug = mean_concentration * vol_mL,
         prot_ug.cm2 = prot_ug / Area_cm,
         prot_mg.cm2 = prot_ug.cm2 / 1000) %>% 
  na.omit() # omitting NAs for now 

## The FLD TP2 corals do not have surface area (I don't think I have pictures of them. Will need to get their surface area via wax dipping)
```

## Summarize 
```{r}
mean.prot <- summarySE(prot, measurevar = "prot_mg.cm2", groupvars = c("Treatment", "Timepoint"))

mean.prot_site <- summarySE(prot, measurevar = "prot_mg.cm2", groupvars = c("Site", "Treatment", "Timepoint"))
# I did not equally sample across site, which was a mistake on my part. I need to figure out how to present the data with site 
```

## Save data 
```{r}
write.csv(prot, "../output/Physiology/SolubleProtein/SolubleProtein.csv")
```

# Stats and plotting 

Load data 
```{r}
prot <- read.csv("../output/Physiology/SolubleProtein/SolubleProtein.csv")

prot <- prot %>%
  mutate(Treatment = ifelse(Timepoint == "TP0", "Acclimation", as.character(Treatment))) %>%
  mutate(Treatment = as.factor(Treatment))
```

## Check for outliers 
```{r}
# Remove field data for now 
prot_exp <- prot %>%
  filter(!Treatment == "Field")

# Look for outliers visually
ggplot(prot_exp, aes(x = Treatment, y = prot_mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(prot_exp$prot_mg.cm2, 0.25)
Q3 <- quantile(prot_exp$prot_mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- prot_exp[prot_exp$prot_mg.cm2 < lower_bound | prot_exp$prot_mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
prot_exp$z_score <- scale(prot_exp$prot_mg.cm2)
outliers_z <- prot_exp[abs(prot_exp$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(prot_exp$prot_mg.cm2, 0.25)
Q3 <- quantile(prot_exp$prot_mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
prot_exp_filtered <- prot_exp[prot_exp$prot_mg.cm2 >= lower_bound & prot_exp$prot_mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(prot_exp)
dim(prot_exp_filtered)
```
5 outliers removed. 

Assess normality of data 
```{r}
# Histogram
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(prot_exp_filtered$prot_mg.cm2)
qqline(prot_exp_filtered$prot_mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(prot_exp_filtered$prot_mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
prot_exp_filtered$prot_mg.cm2_sqrt <- sqrt(prot_exp_filtered$prot_mg.cm2)
shapiro.test(prot_exp_filtered$prot_mg.cm2_sqrt)
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(prot_exp_filtered$prot_mg.cm2_sqrt)
qqline(prot_exp_filtered$prot_mg.cm2_sqrt, col = "red")

# log 
prot_exp_filtered$prot_mg.cm2_log <- log(prot_exp_filtered$prot_mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(prot_exp_filtered$prot_mg.cm2_log)  # Check normality
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(prot_exp_filtered$prot_mg.cm2_log)
qqline(prot_exp_filtered$prot_mg.cm2_log, col = "red")

# Inverse 
prot_exp_filtered$prot_mg.cm2_inv <- 1 / (prot_exp_filtered$prot_mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(prot_exp_filtered$prot_mg.cm2_inv)
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(prot_exp_filtered$prot_mg.cm2_inv)
qqline(prot_exp_filtered$prot_mg.cm2_inv, col = "red")
```

Inverse transformation produces normality. 

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
prot_exp_filtered$Treatment <- as.factor(prot_exp_filtered$Treatment)
prot_exp_filtered$Timepoint <- as.factor(prot_exp_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(prot_mg.cm2_inv ~ Treatment * Timepoint, data = prot_exp_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(prot_mg.cm2_inv ~ Treatment * Timepoint, data = prot_exp_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

Timepoint is significant but Treatment and interaction is not. 

Run post-hoc test for Timepoint
```{r}
TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)

plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Format data for plotting 
```{r}
# Assign month to TPs 
prot_exp_filtered <- prot_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP0" ~ "February", 
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Plot inverse 
```{r}
anova_summary <- summary(anova_result)

# Extract p-values
p_treatment <- anova_summary[[1]]$`Pr(>F)`[1]  # Treatment p-value
p_timepoint <- anova_summary[[1]]$`Pr(>F)`[2]  # Timepoint p-value
p_interaction <- anova_summary[[1]]$`Pr(>F)`[3]  # Interaction p-value

# Format text for annotation
anova_text <- paste0("Two-way ANOVA Results:",
                     "Treatment p = ", signif(p_treatment, 3),
                     "Timepoint p = ", signif(p_timepoint, 3),
                     "Interaction p = ", signif(p_interaction, 3))

# Set Month as factor 
prot_exp_filtered$Month <- factor(prot_exp_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# Create the plot
prot_inv <- ggplot(prot_exp_filtered, aes(x = Month, y = prot_mg.cm2_inv, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7) + 
  theme_classic() +
  labs(x = "",
       y = "Soluble Protein (inverse-transformed mg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
      theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 25),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); chla
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoint_InverseTransformed.pdf", prot_inv, width = 15, height = 10)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoint_InverseTransformed.png", prot_inv, width = 15, height = 10)
```

Plot untransformed data 
```{r}
# Set Month as factor 
prot_exp_filtered$Month <- factor(prot_exp_filtered$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

prot <- ggplot(prot_exp_filtered, aes(x = Month, y = prot_mg.cm2, fill = Treatment, color = Treatment)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA, size = 1) +  
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.7, size = 4) + 
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted")+
  theme_classic() +
  xlab("") +
  ylab("Soluble Protein (mg/cm²)") +
  scale_fill_manual(values = c("grey", "blue", "red")) +
  scale_colour_manual(values = c("grey","blue", "red")) + 
  theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 15),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
  theme(axis.text.x = element_text(size = 25,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1, 
                                   face = "bold"),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.y = element_text(size = 25,
                                   color = "black"),
        axis.title.y = element_text(face = "bold", size = 30,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); prot
  #annotate("text", x = max(as.numeric(chl_exp_filtered$Timepoint)) - 0.2, 
  #         y = max(chl_exp_filtered$chla_sqrt), label = anova_text, 
  #         hjust = 1, vjust = 1, size = 5, color = "black")

ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoint.pdf", prot, width = 20, height = 15)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoint.png", prot, width = 20, height = 15)
```

Calculate mean, sd and sem of each group for plotting 
```{r}
prot_summary <- prot_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_prot = mean(prot_mg.cm2, na.rm = TRUE),
    sd_prot = sd(prot_mg.cm2, na.rm = TRUE),
    sem_prot = prot_mg.cm2 / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
prot_summary$Month <- factor(prot_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
prot_dot_sem <- ggplot(prot_summary, aes(x = Month, y = mean_prot, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_prot - sem_prot, 
                    ymax = mean_prot + sem_prot), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  xlab("") +
  ylab("Soluble Protein (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); prot_dot_sem
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sem.pdf", prot_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sem.png", prot_dot_sem, width = 20, height = 15)

# sd
prot_dot_sd <- ggplot(prot_summary, aes(x = Month, y = mean_prot, color = Treatment)) +
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_prot - sd_prot, 
                    ymax = mean_prot + sd_prot),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab("Soluble Protein (mg/cm²)") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );prot_dot_sd
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sd.pdf", prot_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sd.png", prot_dot_sd, width = 22, height = 12)
```

# Remove February and rerun stats

Read in data 
```{r}
prot <- read.csv("../output/Physiology/SolubleProtein/SolubleProtein.csv")
```

Remove field and Feb samples 
```{r}
prot_exp <- prot %>%
  filter(!Treatment == "Field") %>%
  filter(!Timepoint == "TP0")
```

## Check for outliers 
```{r}
# Look for outliers visually
ggplot(prot_exp, aes(x = Treatment, y = prot_mg.cm2)) + 
  geom_boxplot(outlier.color = "red", outlier.shape = 16) +
  theme_minimal()

# IQR method 
Q1 <- quantile(prot_exp$prot_mg.cm2, 0.25)
Q3 <- quantile(prot_exp$prot_mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val
outliers <- prot_exp[prot_exp$prot_mg.cm2 < lower_bound | prot_exp$prot_mg.cm2 > upper_bound, ]
print(outliers)

# Z-score method
prot_exp$z_score <- scale(prot_exp$prot_mg.cm2)
outliers_z <- prot_exp[abs(prot_exp$z_score) > 3, ]
print(outliers_z)
```

Remove outliers based on IQR
```{r}
# Calculate IQR
Q1 <- quantile(prot_exp$prot_mg.cm2, 0.25)
Q3 <- quantile(prot_exp$prot_mg.cm2, 0.75)
IQR_val <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Filter dataset to remove outliers
prot_exp_filtered <- prot_exp[prot_exp$prot_mg.cm2 >= lower_bound & prot_exp$prot_mg.cm2 <= upper_bound, ]

# Check dimensions before and after
dim(prot_exp)
dim(prot_exp_filtered)
```

4 outliers removed. 

Assess normality of data 
```{r}
# Histogram
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()

# Q-Q Plot
qqnorm(prot_exp_filtered$prot_mg.cm2)
qqline(prot_exp_filtered$prot_mg.cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(prot_exp_filtered$prot_mg.cm2)
```

Data is not normal, perform data transformation
```{r}
# Sqrt
prot_exp_filtered$prot_mg.cm2_sqrt <- sqrt(prot_exp_filtered$prot_mg.cm2)
shapiro.test(prot_exp_filtered$prot_mg.cm2_sqrt)
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2_sqrt)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(prot_exp_filtered$prot_mg.cm2_sqrt)
qqline(prot_exp_filtered$prot_mg.cm2_sqrt, col = "red")

# log 
prot_exp_filtered$prot_mg.cm2_log <- log(prot_exp_filtered$prot_mg.cm2 + 1)  # Add 1 to avoid log(0)
shapiro.test(prot_exp_filtered$prot_mg.cm2_log)  # Check normality
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2_log)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(prot_exp_filtered$prot_mg.cm2_log)
qqline(prot_exp_filtered$prot_mg.cm2_log, col = "red")

# Inverse 
prot_exp_filtered$prot_mg.cm2_inv <- 1 / (prot_exp_filtered$prot_mg.cm2 + 1)  # Add 1 to avoid division by zero
shapiro.test(prot_exp_filtered$prot_mg.cm2_inv)
ggplot(prot_exp_filtered, aes(x = prot_mg.cm2_inv)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  theme_minimal()
qqnorm(prot_exp_filtered$prot_mg.cm2_inv)
qqline(prot_exp_filtered$prot_mg.cm2_inv, col = "red")
```

Inverse and log transformation produces normality. Let's stick with inverse, as that is what I used prior to removing Feb. 

Run a two-way ANOVA
```{r}
# Ensure factors are properly set
prot_exp_filtered$Treatment <- as.factor(prot_exp_filtered$Treatment)
prot_exp_filtered$Timepoint <- as.factor(prot_exp_filtered$Timepoint)

# Run two-way ANOVA
anova_result <- aov(prot_mg.cm2_inv ~ Treatment * Timepoint, data = prot_exp_filtered)
summary(anova_result)

# Check for homogenity of variance 
leveneTest(prot_mg.cm2_inv ~ Treatment * Timepoint, data = prot_exp_filtered)

# Check residuals 
# Residuals vs Fitted Plot (Check homoscedasticity)
plot(anova_result, which = 1)

# Q-Q Plot of Residuals (Check normality)
plot(anova_result, which = 2)
```

Timepoint is significant but Treatment and interaction is not. 

Run post-hoc test for Timepoint
```{r}
TukeyHSD(anova_result, "Timepoint", conf.level = 0.95)
plot(TukeyHSD(anova_result, conf.level=.95), las = 2)
```

Assign month to TPs
```{r}
prot_exp_filtered <- prot_exp_filtered %>% 
  mutate(Month = case_when(
    Timepoint == "TP1" ~ "March", 
    Timepoint == "TP2" ~ "Early April", 
    Timepoint == "TP3" ~ "Late April", 
    Timepoint == "TP4" ~ "May", 
    Timepoint == "TP5" ~ "June", 
    Timepoint == "TP6" ~ "July", 
    Timepoint == "TP7" ~ "August")) 
```

Calculate mean, sd and sem of each group for plotting 
```{r}
prot_summary <- prot_exp_filtered %>%
  group_by(Treatment, Month) %>%
  dplyr::summarize(
    n = n(),  # Count observations per group
    mean_prot = mean(prot_mg.cm2, na.rm = TRUE),
    sd_prot = sd(prot_mg.cm2, na.rm = TRUE),
    sem_prot = prot_mg.cm2 / sqrt(n),  # SEM formula
    .groups = "drop"
  )
```

Plot as dot and whisker plots
```{r}
# Set Month as factor 
prot_summary$Month <- factor(prot_summary$Month, levels = c("February", "March", "Early April", "Late April", "May", "June", "July", "August"))

# sem
prot_dot_sem <- ggplot(prot_summary, aes(x = Month, y = mean_prot, color = Treatment)) +
  geom_line(aes(group = Treatment), linewidth = 1.5) +  # Connection lines
  geom_errorbar(aes(ymin = mean_prot - sem_prot, 
                    ymax = mean_prot + sem_prot), 
                width = 0.2) +
  geom_point(size = 5, aes(fill = Treatment), shape = 21, color = "black") +
  xlab("") +
  ylab("Soluble Protein (mg/cm²)") +
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",  # Blue
                               "Heat" = "#E41A1C")) +  # Red
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right"); prot_dot_sem
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sem_noFeb.pdf", prot_dot_sem, width = 20, height = 15)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sem_noFeb.png", prot_dot_sem, width = 20, height = 15)

# sd
prot_dot_sd <- ggplot(prot_summary, aes(x = Month, y = mean_prot, color = Treatment)) +
  #geom_hline(yintercept = 0, color = "black", linetype = "dotted", linewidth = 1)+
  geom_line(aes(group = Treatment), linewidth = 4) +
  geom_errorbar(aes(ymin = mean_prot - sd_prot, 
                    ymax = mean_prot + sd_prot),
                linewidth = 2,  # Thicker error bars
                width = 0.5) +
  geom_point(size = 10, aes(fill = Treatment), 
             shape = 21, color = "black", stroke = 2) +  # Thicker point borders
  scale_color_manual(values = c("Acclimation" = "grey60", 
                               "Ambient" = "#377EB8",
                               "Heat" = "#E41A1C")) +
  scale_fill_manual(values = c("Acclimation" = "grey80", 
                              "Ambient" = "#377EB8", 
                              "Heat" = "#E41A1C")) +
  xlab("") +
  ylab("Soluble Protein (mg/cm²)") +
  theme_classic() +  # Increased base font size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold", color = "black"),
    axis.text.y = element_text(size = 30, face = "bold", color = "black"),
    axis.title.y = element_text(size = 40, margin = margin(r = 15), face = "bold", color = "black"),
    axis.ticks.x = element_line(size = 1, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    legend.title = element_text(size = 35, face = "bold"),
    legend.text = element_text(size = 25),
    legend.key.size = unit(1.5, "cm"),
    legend.spacing.y = unit(0.5, "cm")
  );prot_dot_sd
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sd_noFeb.pdf", prot_dot_sd, width = 22, height = 12)
ggsave("../output/Physiology/SolubleProtein/SolubleProtein_Treatment_Timepoin_dotwhisker_sd_noFeb.png", prot_dot_sd, width = 22, height = 12)
```
