---
title: "TemperatureData"
author: "jillashey"
date: "2023-05-21"
output: html_document
---

This script will plot my experimental temperature data and the Newport 2021 NOAA Buoy data on one plot. This is the [link](https://www.ndbc.noaa.gov/station_realtime.php?station=nwpr1) to the real time Newport NOAA data. 

## Load libraries 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(tidyr)
library(ggplot2)
library(plotrix)
library(tidyverse)
library(dplyr)
library(reshape2)
library(gridExtra)
library(FSA)
library(naniar)
```

## Load my data 
```{r}
my.data <- read.csv("../output/Environmental/Hobo/temp_light_logger/All_Tank_HoboTempLight_data.csv", header = T)
colnames(my.data)[colnames(my.data) == 'Date'] <- 'Date.Time'
my.data <- my.data[,c("Date.Time", 
                      "Tank1_Temp", 
                      "Tank2_Temp", 
                      "Tank3_Temp",
                      "Tank4_Temp",
                      "Tank5_Temp", 
                      "Tank6_Temp", 
                      "Tank7_Temp",
                      "Tank8_Temp",
                      "Tank9_Temp",
                      "Tank10_Temp", 
                      "Tank11_Temp", 
                      "Tank12_Temp", 
                      "Tank15_Temp",
                      "Tank16_Temp", 
                      "Tank17_Temp", 
                      "Tank18_Temp")]
my.data <- na.omit(my.data) ### should I omit NAs????
```


I am filtering out the random temp data points (ie when the logger was removed from the water to be read) that do not fit the specific temp window that was set at a specific time. 

AMBIENT
```{r}
# Parse data and time 
my.data$Date.Time <- as.POSIXct(my.data$Date.Time, "%Y-%m-%d %H:%M:%OS", tz = "EST")

# Replace values outside specified ranges with NAs -- ambient 
test <- my.data %>%
  mutate(across(c(Tank1_Temp, Tank2_Temp, Tank5_Temp, Tank8_Temp, Tank10_Temp, Tank12_Temp, Tank16_Temp, Tank18_Temp),
                ~ case_when(
                   # Date range: 02/17 to 03/17
                  Date.Time >= as.Date.POSIXct("2021-02-07 01:00:00", tz = "EST") & Date.Time <= as.Date.POSIXct("2021-03-17 01:00:00", tz = "EST") & (.x < 4.4 | .x > 5.7) ~ NA_real_,
                  # Date range: 03/18 to 03/29
                  Date.Time >= as.Date("2021-03-17 01:10:00", tz = "EST") & Date.Time <= as.Date("2021-03-29 01:00:00", tz = "EST") & (.x > 6.3 | .x < 6) ~ NA_real_,
                  # Date range: 03/30 to 04/05
                  Date.Time >= as.Date("2021-03-30 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-05 01:00:00", tz = "EST") & (.x > 6.9 | .x < 6.5) ~ NA_real_,
                  # Date range: 04/06 to 04/12
                  Date.Time >= as.Date("2021-04-06 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-12 01:00:00", tz = "EST") & (.x > 8.1 | .x < 7) ~ NA_real_,
                  # Date range: 04/13 to 04/19
                  Date.Time >= as.Date("2021-04-13 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-19 01:00:00", tz = "EST") & (.x > 9.2 | .x < 8.2) ~ NA_real_,
                  # Date range: 04/20 to 04/23
                  Date.Time >= as.Date("2021-04-20 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-23 01:00:00", tz = "EST") & (.x > 9.8 | .x < 9.3) ~ NA_real_,
                  # Date range: 04/24 to 05/03
                  Date.Time >= as.Date("2021-04-24 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-03 01:00:00", tz = "EST") & (.x > 10.7 | .x < 10) ~ NA_real_,
                  # Date range: 05/04 to 05/12
                  Date.Time >= as.Date("2021-05-04 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-12 01:00:00", tz = "EST") & (.x > 11.4 | .x < 10.8) ~ NA_real_,
                  # Date range: 05/13 to 05/20
                  Date.Time >= as.Date("2021-05-13 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-20 01:00:00", tz = "EST") & (.x > 12.7 | .x < 11.6) ~ NA_real_,
                  # Date range: 05/21 to 05/29
                  Date.Time >= as.Date("2021-05-21 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-29 01:00:00", tz = "EST") & (.x > 14.5 | .x < 13.3) ~ NA_real_,
                  # Date range: 05/30 to 06/09
                  Date.Time >= as.Date("2021-05-30 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-06-09 01:00:00", tz = "EST") & (.x > 15.5 | .x < 14.6) ~ NA_real_,
                  # Date range: 06/10 to 06/17
                  Date.Time >= as.Date("2021-06-10 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-06-17 01:00:00", tz = "EST") & (.x > 16.4 | .x < 16) ~ NA_real_,
                  # Date range: 06/18 to 06/25
                  Date.Time >= as.Date("2021-06-18 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-06-25 01:00:00", tz = "EST") & (.x > 16.9 | .x < 16.5) ~ NA_real_,
                  # Date range: 06/26 to 07/01
                  Date.Time >= as.Date("2021-06-26 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-01 01:00:00", tz = "EST") & (.x > 17.7 | .x < 17) ~ NA_real_,
                  # Date range: 07/02 to 07/12
                  Date.Time >= as.Date("2021-07-02 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-12 01:00:00", tz = "EST") & (.x > 18.3 | .x < 17.8) ~ NA_real_,
                  # Date range: 07/13 to 07/17
                  Date.Time >= as.Date("2021-07-13 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-17 01:00:00", tz = "EST") & (.x > 18.9 | .x < 18.4) ~ NA_real_,
                  # Date range: 07/18 to 07/22
                  Date.Time >= as.Date("2021-07-18 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-22 01:00:00", tz = "EST") & (.x > 19.9 | .x < 19) ~ NA_real_,
                  # Date range: 07/23 to 08/20
                  Date.Time >= as.Date("2021-07-23 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-08-20 01:00:00", tz = "EST") & (.x > 23 | .x < 20) ~ NA_real_,
                  TRUE ~ .x
                )))
                  
# Calculate mean of each row
test$Temp_Amb <- rowMeans(test[,c("Tank1_Temp", "Tank2_Temp", "Tank5_Temp",
                                        "Tank8_Temp", "Tank10_Temp", "Tank12_Temp",
                                        "Tank16_Temp", "Tank18_Temp")], na.rm=TRUE)

# Calculate standard deviation of each row
test$Temp_Amb_StdDev <- apply(test[, c("Tank1_Temp", "Tank2_Temp", "Tank5_Temp",
                                        "Tank8_Temp", "Tank10_Temp", "Tank12_Temp",
                                        "Tank16_Temp", "Tank18_Temp")], 
                               1,   # 1 indicates row-wise operation
                               function(x) sd(x, na.rm = TRUE))

# Calculate upper and lower std deviations from the mean
test <- test %>%
  mutate(upper = Temp_Amb+Temp_Amb_StdDev,
         lower = Temp_Amb-Temp_Amb_StdDev)

# Select specific columns 
my.data_amb <- test %>% 
  dplyr::select(Date.Time, Temp_Amb, Temp_Amb_StdDev, upper, lower) %>%
  dplyr::rename("Temperature" = "Temp_Amb") %>%
  dplyr::rename("sd" = "Temp_Amb_StdDev") %>%
  #separate(col = Date.Time, into = c("Date", "Time"), sep = " ") %>%
  na.omit()
my.data_amb$Treatment <- "Ambient"

ggplot(data = my.data_amb, aes(x = Date.Time, y = Temperature, color = Treatment)) +
  geom_line()

ggplot(data = my.data_amb, aes(x=Date.Time, y=Temperature, colour=Treatment, fill=Treatment))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=Treatment), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=Treatment), alpha=1, show.legend=TRUE)
```

HEAT
```{r}
# Parse data and time 
#my.data$Date.Time <- as.POSIXct(my.data$Date.Time, "%Y-%m-%d %H:%M:%OS", tz = "EST")

# Replace values outside specified ranges with NAs -- ambient 
test2 <- my.data %>%
  mutate(across(c(Tank3_Temp, Tank4_Temp, Tank6_Temp, Tank7_Temp, Tank9_Temp, Tank11_Temp, Tank15_Temp, Tank17_Temp),
                ~ case_when(
                   # Date range: 02/17 to 03/17
                  Date.Time >= as.Date.POSIXct("2021-02-07 01:00:00", tz = "EST") & Date.Time <= as.Date.POSIXct("2021-03-17 01:00:00", tz = "EST") & (.x < 7.5 | .x > 9.2) ~ NA_real_,
                  # Date range: 03/18 to 03/29
                  Date.Time >= as.Date("2021-03-17 01:10:00", tz = "EST") & Date.Time <= as.Date("2021-03-29 01:00:00", tz = "EST") & (.x > 9.8 | .x < 9.3) ~ NA_real_,
                  # Date range: 03/30 to 04/05
                  Date.Time >= as.Date("2021-03-30 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-05 01:00:00", tz = "EST") & (.x > 10.2 | .x < 9.9) ~ NA_real_,
                  # Date range: 04/06 to 04/12
                  Date.Time >= as.Date("2021-04-06 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-12 01:00:00", tz = "EST") & (.x > 11.5 | .x < 10.3) ~ NA_real_,
                  # Date range: 04/13 to 04/19
                  Date.Time >= as.Date("2021-04-13 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-19 01:00:00", tz = "EST") & (.x > 12.2 | .x < 11.6) ~ NA_real_,
                  # Date range: 04/20 to 04/23
                  Date.Time >= as.Date("2021-04-20 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-04-23 01:00:00", tz = "EST") & (.x > 12.9 | .x < 12.3) ~ NA_real_,
                  # Date range: 04/24 to 05/03
                  Date.Time >= as.Date("2021-04-24 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-03 01:00:00", tz = "EST") & (.x > 13.7 | .x < 13) ~ NA_real_,
                  # Date range: 05/04 to 05/12
                  Date.Time >= as.Date("2021-05-04 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-12 01:00:00", tz = "EST") & (.x > 14.6 | .x < 13.8) ~ NA_real_,
                  # Date range: 05/13 to 05/20
                  Date.Time >= as.Date("2021-05-13 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-20 01:00:00", tz = "EST") & (.x > 16 | .x < 14.7) ~ NA_real_,
                  # Date range: 05/21 to 05/29
                  Date.Time >= as.Date("2021-05-21 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-05-29 01:00:00", tz = "EST") & (.x > 17.5 | .x < 16.1) ~ NA_real_,
                  # Date range: 05/30 to 06/09
                  Date.Time >= as.Date("2021-05-30 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-06-09 01:00:00", tz = "EST") & (.x > 18.5 | .x < 17.6) ~ NA_real_,
                  # Date range: 06/10 to 06/17
                  Date.Time >= as.Date("2021-06-10 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-06-17 01:00:00", tz = "EST") & (.x > 19.6 | .x < 18.6) ~ NA_real_,
                  # Date range: 06/18 to 06/25
                  Date.Time >= as.Date("2021-06-18 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-06-25 01:00:00", tz = "EST") & (.x > 20.4 | .x < 19.7) ~ NA_real_,
                  # Date range: 06/26 to 07/01
                  Date.Time >= as.Date("2021-06-26 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-01 01:00:00", tz = "EST") & (.x > 21 | .x < 20.2) ~ NA_real_,
                  # Date range: 07/02 to 07/12
                  Date.Time >= as.Date("2021-07-02 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-12 01:00:00", tz = "EST") & (.x > 22.6 | .x < 21.1) ~ NA_real_,
                  # Date range: 07/13 to 07/17
                  Date.Time >= as.Date("2021-07-13 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-17 01:00:00", tz = "EST") & (.x > 23.6 | .x < 22.7) ~ NA_real_,
                  # Date range: 07/18 to 07/22
                  Date.Time >= as.Date("2021-07-18 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-07-22 01:00:00", tz = "EST") & (.x > 24.4 | .x < 23.4) ~ NA_real_,
                  # Date range: 07/23 to 08/20
                  Date.Time >= as.Date("2021-07-23 01:00:00", tz = "EST") & Date.Time <= as.Date("2021-08-20 01:00:00", tz = "EST") & (.x > 26 | .x < 23) ~ NA_real_,
                  TRUE ~ .x
                )))
                  
# Calculate mean of each row
test2$Temp_Heat <- rowMeans(test2[,c("Tank3_Temp", "Tank4_Temp", "Tank6_Temp",
                                        "Tank7_Temp", "Tank9_Temp", "Tank11_Temp",
                                        "Tank15_Temp", "Tank17_Temp")], na.rm=TRUE)

# Calculate standard deviation of each row
test2$Temp_Heat_StdDev <- apply(test2[, c("Tank3_Temp", "Tank4_Temp", "Tank6_Temp",
                                        "Tank7_Temp", "Tank9_Temp", "Tank11_Temp",
                                        "Tank15_Temp", "Tank17_Temp")], 
                               1,   # 1 indicates row-wise operation
                               function(x) sd(x, na.rm = TRUE))

# Calculate upper and lower std deviations from the mean
test2 <- test2 %>%
  mutate(upper = Temp_Heat+Temp_Heat_StdDev,
         lower = Temp_Heat-Temp_Heat_StdDev)

# Select specific columns 
my.data_heat <- test2 %>% 
  dplyr::select(Date.Time, Temp_Heat, Temp_Heat_StdDev, upper, lower) %>%
  dplyr::rename("Temperature" = "Temp_Heat") %>%
  dplyr::rename("sd" = "Temp_Heat_StdDev") %>%
  #separate(col = Date.Time, into = c("Date", "Time"), sep = " ") %>%
  na.omit()
my.data_heat$Treatment <- "Heat"

ggplot(data = my.data_heat, aes(x = Date.Time, y = Temperature, color = Treatment)) +
  geom_line()

ggplot(data = my.data_heat, aes(x=Date.Time, y=Temperature, colour=Treatment, fill=Treatment))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=Treatment), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=Treatment), alpha=1, show.legend=TRUE)
```

Bind ambient and heat dfs and plot 
```{r}
all <- bind_rows(my.data_amb, my.data_heat)

ggplot(data = all, aes(x = Date.Time, y = Temperature, color = Treatment)) +
  geom_line()

temp_plot <- ggplot(data = all, aes(x=Date.Time, y=Temperature, colour=Treatment, fill=Treatment))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=Treatment), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=Treatment), alpha=1, show.legend=TRUE) +
  geom_vline(xintercept=as.POSIXct("2021-04-02 13:00:00", tz = "EST"), linetype="dashed", color="black")+ # TP2
  geom_vline(xintercept=as.POSIXct("2021-04-29 13:00:00", tz = "EST"), linetype="dashed", color="black")+ # TP3
  geom_vline(xintercept=as.POSIXct("2021-05-28 13:30:00", tz = "EST"), linetype="dashed", color="black")+ # TP4
  geom_vline(xintercept=as.POSIXct("2021-06-25 13:30:00", tz = "EST"), linetype="dashed", color="black")+ # TP5 
  geom_vline(xintercept=as.POSIXct("2021-07-23 13:30:00", tz = "EST"), linetype="dashed", color="black")+ # TP6
  geom_vline(xintercept=as.POSIXct("2021-08-20 13:30:00", tz = "EST"), linetype="dashed", color="black")+ # TP7
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  #xlab("Date")+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  scale_fill_manual(values=c("blue", "red"), name="Temperature")+
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b") +
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14),
    legend.position="right"
  ); temp_plot
```

I still need to make the plots pretty and get rid of the some of the ouliers but I'm happy with how this turned out. 

Read in NOAA data 
```{r}
# Get a list of all text files in the folder
file_list <- list.files("../data/Environmental/NOAA_temp/blah/", pattern = "\\.txt$", full.names = TRUE)

# Read and process each file
data_list <- lapply(file_list, function(file) {
  # Read the file
  data <- read.csv(file, sep = "\t", skip = 0, header = TRUE, na.strings = "NA")[, c(1:5, 15)]
  
  # Create Date.Time column
  #data$Date.Time <- as.POSIXct(paste(data$MM, data$DD, sep = "-"), format = "%m-%d")
  data$Date <- (paste(data$MM, data$DD, sep = "-"))

  # Replace specific values in WTMP column with NA
  data <- data %>% replace_with_na(replace = list(WTMP = c("MM", "9   9.", ".0 999", "999", "0 999.")))
  
  # Convert WTMP to numeric
  data$WTMP <- as.numeric(data$WTMP)
  
  return(data)
})

# Combine the data frames into a single data frame
combined_data <- bind_rows(data_list)
```

Add fake year of 2021 so I can plot it and the plot will not hate me. Calculate mean, sd, upper and lower thresholds
```{r}
combined_data$fake.year <- "2021"
combined_data$Date.Time <- as.POSIXct(paste(combined_data$fake.year, combined_data$Date, sep = "-"), format = "%Y-%m-%d")

noaa_data <- combined_data %>%
  dplyr::group_by(Date.Time) %>%
  dplyr::summarize(
    Temperature = mean(WTMP, na.rm = TRUE),
    sd = sd(WTMP, na.rm = TRUE)) %>%
  dplyr::mutate(upper = Temperature+sd,
         lower = Temperature-sd, 
         Treatment = "NOAA")
```

Remove unneeded columns
```{r}
noaa_data <- noaa_data %>%
  dplyr::select(Date.Time, Temperature, sd, upper, lower, Treatment)
```

Bind dfs together 
```{r}
test <- rbind(all, noaa_data)
```

Plot
```{r}
temp_plot <- ggplot(data = test, aes(x=Date.Time, y=Temperature, colour=Treatment, fill=Treatment))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=Treatment), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=Treatment), alpha=1, show.legend=TRUE, size = 3) +
  geom_vline(xintercept=as.POSIXct("2021-02-04 13:00:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP0
  geom_vline(xintercept=as.POSIXct("2021-03-04 13:00:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP1
  geom_vline(xintercept=as.POSIXct("2021-04-02 13:00:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP2
  geom_vline(xintercept=as.POSIXct("2021-04-29 13:00:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP3
  geom_vline(xintercept=as.POSIXct("2021-05-28 13:30:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP4
  geom_vline(xintercept=as.POSIXct("2021-06-25 13:30:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP5 
  geom_vline(xintercept=as.POSIXct("2021-07-23 13:30:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP6
  geom_vline(xintercept=as.POSIXct("2021-08-20 13:30:00", tz = "EST"), linetype="dashed", color="black", size = 1.5)+ # TP7
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  xlab("")+
  scale_colour_manual(values=c("blue", "red", "black"), name="Temperature")+
  scale_fill_manual(values=c("blue", "red", "black"), name="Temperature")+
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b") +
  theme_classic()+
    theme(legend.title = element_text(size = 28, face = "bold"),
          legend.text = element_text(size = 25),
          legend.position = "right") +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0))); temp_plot 

ggsave("../output/Environmental/MyData_NOAA_temperature_smooth.png", temp_plot, width=20, height=13)
ggsave("../output/Environmental/MyData_NOAA_temperature_smooth.pdf", temp_plot, width=20, height=13)
```

BEAUTIFUL!!!!!!!!! Still need to QC my temp data to get rid of the weird spikes. 


