---
title: "topGO enrichment across seasons"
author: "Jill Ashey"
date: "2025-03-19"
output: html_document
---

GO enrichment of DEGs expressed in February, June and August 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#devtools::install_github("stemangiola/tidyHeatmap")

library(tidyverse)
library(topGO)
library(tidyHeatmap)
library(purrr)
library(scales)
library(igraph)
library(tidygraph)
library(ggraph)
library(UpSetR)
library(ggupset)
library(patchwork)
```

## Feb DEGs

Read in DEGs of Feb v June and August ambient
```{r}
feb_june <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP5_amb_DEG.csv")
feb_aug <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP7_amb_DEG.csv")
```

Subset genes that are upregulated (ie those that were expressed in Feb relative to June and August)
```{r}
feb_june <- feb_june %>%
  filter(log2FoldChange > 0)
feb_aug <- feb_aug %>%
  filter(log2FoldChange > 0)

# Bind 
feb_all <- rbind(feb_june,feb_aug)[,1]
feb_all_unique <- feb_all %>%
  dplyr::rename("gene_id" = "...1") %>%
  distinct(gene_id)
```

Read in gene2go information (generated in `Network_analysis.Rmd` file)
```{r}
ast_gene2go <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(feb_all_unique$gene_id)

# All genes 
all_genes <- as.character(ast_gene2go$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 150)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_ast_BP_En_sig` with GO and gene info. Then merge with Feb genes 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_ast_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_ast_BP_En_sig_gene_feb <- GO_ast_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_ast_BP_En_sig_gene_feb$ontology <- "Biological Processes"
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_CC_En_sig_gene_feb <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene_feb$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_MF_En_sig_gene_feb <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene_feb$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_ast_BP_En_sig_gene_feb, GO_CC_En_sig_gene_feb, GO_MF_En_sig_gene_feb)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))
length(unique(GO_En_sig_gene$mRNA))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

## June DEGs

Read in DEGs of June v Feb and August ambient
```{r}
feb_june <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP5_amb_DEG.csv")
june_aug <- read_csv("../output/Molecular/mRNA/TP5_amb_v_TP7_amb_DEG.csv")
```

Subset genes that are downregulated relative to February and upregulated relative to August 
```{r}
feb_june <- feb_june %>%
  filter(log2FoldChange < 0)
june_aug <- june_aug %>%
  filter(log2FoldChange > 0)

# Bind 
june_all <- rbind(feb_june,june_aug)[,1]
june_all_unique <- june_all %>%
  dplyr::rename(gene_id = ...1) %>%
  distinct(gene_id)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(june_all_unique$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 150)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_ast_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_ast_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_ast_BP_En_sig_gene_june <- GO_ast_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_ast_BP_En_sig_gene_june$ontology <- "Biological Processes"
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_CC_En_sig_gene_june <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene_june$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in June (ie those used for the functional enrichment)
GO_MF_En_sig_gene_june <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene_june$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_ast_BP_En_sig_gene_june, GO_CC_En_sig_gene_june, GO_MF_En_sig_gene_june)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

## August DEGs

Read in DEGs of August v Feb and June ambient
```{r}
feb_aug <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP7_amb_DEG.csv")
june_aug <- read_csv("../output/Molecular/mRNA/TP5_amb_v_TP7_amb_DEG.csv")
```

Subset genes that are downregulated relative to February and downregulated relative to June 
```{r}
feb_aug <- feb_aug %>%
  filter(log2FoldChange < 0)
june_aug <- june_aug %>%
  filter(log2FoldChange < 0)

# Bind 
aug_all <- rbind(feb_aug,june_aug)[,1]
aug_all_unique <- aug_all %>%
  dplyr::rename(gene_id = ...1) %>%
  distinct(gene_id)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(aug_all_unique$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 155)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_ast_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_ast_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in August (ie those used for the functional enrichment)
GO_ast_BP_En_sig_gene_aug <- GO_ast_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_ast_BP_En_sig_gene_aug$ontology <- "Biological Processes"
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in August (ie those used for the functional enrichment)
GO_CC_En_sig_gene_aug <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene_aug$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in August (ie those used for the functional enrichment)
GO_MF_En_sig_gene_aug <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene_aug$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_ast_BP_En_sig_gene_aug, GO_CC_En_sig_gene_aug, GO_MF_En_sig_gene_aug)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

## Examining any overlaps in functional terms across season

BP ONLY

```{r}
feb_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes.csv") %>%
  filter(ontology == "Biological Processes")
length(unique(feb_genes$Term))
unique_terms_by_ontology <- tapply(feb_genes$Term, feb_genes$ontology, function(x) length(unique(x)))
blah <- feb_genes %>%
   group_by(Term,ontology) %>%
   summarise(gene_count = n_distinct(mRNA))

june_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes.csv")%>%
  filter(ontology == "Biological Processes")
length(unique(june_genes$Term))
unique_terms_by_ontology <- tapply(june_genes$Term, june_genes$ontology, function(x) length(unique(x)))
blah2 <- june_genes %>%
   group_by(Term,ontology) %>%
   summarise(gene_count = n_distinct(mRNA))

aug_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes.csv")%>%
  filter(ontology == "Biological Processes")
length(unique(aug_genes$Term))
unique_terms_by_ontology <- tapply(aug_genes$Term, aug_genes$ontology, function(x) length(unique(x)))
blah3 <- aug_genes %>%
   group_by(Term,ontology) %>%
   summarise(gene_count = n_distinct(mRNA))

feb_terms <- unique(feb_genes$Term)
june_terms <- unique(june_genes$Term)
aug_terms <- unique(aug_genes$Term)

# Overlapping terms 
overlapping_terms_feb_june <- intersect(feb_terms, june_terms) 
overlapping_terms_feb_aug <- intersect(feb_terms, aug_terms) 
overlapping_terms_june_aug <- intersect(june_terms, aug_terms)

# Terms shared between Feb, June and August 
common_terms <- Reduce(intersect, list(feb_terms, june_terms, aug_terms)) # none

# Terms unique to Feb 
feb_unique_terms <- setdiff(feb_terms, union(june_terms, aug_terms))

# Terms unique to June 
june_unique_terms <- setdiff(june_terms, union(aug_terms, feb_terms))

# Terms unique to August 
aug_unique_terms <- setdiff(aug_terms, union(june_terms, feb_terms))
```

Filter by unique terms at each timepoint 
```{r}
feb_genes_unique <- feb_genes %>%
  filter(Term %in% feb_unique_terms)

june_genes_unique <- june_genes %>%
  filter(Term %in% june_unique_terms)

aug_genes_unique <- aug_genes %>%
  filter(Term %in% aug_unique_terms)
```

Filter by shared terms for June and August 
```{r}
june_genes_shared <- june_genes %>%
  filter(Term %in% overlapping_terms_june_aug)
aug_genes_shared <- aug_genes %>%
  filter(Term %in% overlapping_terms_june_aug)
```

Make upset plot with GO terms enriched 
```{r}
# ── 1. Load & prep data ────────────────────────────────────────────────────────
feb_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes.csv") %>%
  filter(ontology == "Biological Processes")

june_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes.csv") %>%
  filter(ontology == "Biological Processes")

aug_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes.csv") %>%
  filter(ontology == "Biological Processes")

feb_terms  <- unique(feb_genes$Term)
june_terms <- unique(june_genes$Term)
aug_terms  <- unique(aug_genes$Term)

# ── 2. Define overlap groups ───────────────────────────────────────────────────
groups <- list(
  "February Only"        = setdiff(feb_terms,  union(june_terms, aug_terms)),
  "June Only"            = setdiff(june_terms, union(feb_terms,  aug_terms)),
  "August Only"          = setdiff(aug_terms,  union(feb_terms,  june_terms)),
  "February-June Shared" = setdiff(intersect(feb_terms, june_terms), aug_terms),
  "February-Aug Shared"  = setdiff(intersect(feb_terms, aug_terms),  june_terms),
  "June-August Shared"   = setdiff(intersect(june_terms, aug_terms), feb_terms),
  "All Three"            = Reduce(intersect, list(feb_terms, june_terms, aug_terms))
)

# Remove empty groups
groups <- groups[sapply(groups, length) > 0]

# Order: single timepoints first (sorted by count desc), then shared groups (sorted by count desc)
single_groups <- c("February Only", "June Only", "August Only")
shared_groups <- c("February-June Shared", "February-Aug Shared", "June-August Shared", "All Three")

bar_data_raw <- tibble(
  group = names(groups),
  count = sapply(groups, length)
)

singles <- bar_data_raw %>% filter(group %in% single_groups) %>% arrange(desc(count))
shareds <- bar_data_raw %>% filter(group %in% shared_groups) %>% arrange(desc(count))

# Bar order (bottom to top in horizontal chart): shared first (bottom), then singles (top)
bar_order <- c(rev(shareds$group), rev(singles$group))

bar_data <- bind_rows(singles, shareds) %>%
  mutate(group = factor(group, levels = rev(bar_order)))

# ── 3. UpSet dot matrix data ───────────────────────────────────────────────────
months_in_group <- function(g) {
  list(
    February = g %in% c("February Only", "February-June Shared",
                         "February-Aug Shared", "All Three"),
    June     = g %in% c("June Only", "February-June Shared",
                         "June-August Shared", "All Three"),
    August   = g %in% c("August Only", "February-Aug Shared",
                         "June-August Shared", "All Three")
  )
}

dot_data <- expand.grid(
  group = levels(bar_data$group),
  month = c("August", "June", "February"),
  stringsAsFactors = FALSE
) %>%
  as_tibble() %>%
  rowwise() %>%
  mutate(
    active = switch(month,
      February = group %in% c("February Only","February-June Shared","February-Aug Shared","All Three"),
      June     = group %in% c("June Only","February-June Shared","June-August Shared","All Three"),
      August   = group %in% c("August Only","February-Aug Shared","June-August Shared","All Three")
    )
  ) %>%
  ungroup()

dot_data$group <- factor(dot_data$group, levels = levels(bar_data$group))
dot_data$month <- factor(dot_data$month, levels = c("August", "June", "February"))

# Draw connecting lines for active months within each group
line_data <- dot_data %>%
  filter(active) %>%
  group_by(group) %>%
  filter(n() > 1) %>%
  summarise(
    ymin = min(as.numeric(month)),
    ymax = max(as.numeric(month)),
    .groups = "drop"
  )

# ── 4. Top 5 GO terms per group by p-value ────────────────────────────────────
get_top5 <- function(term_list, group_name) {
  all_data <- bind_rows(
    feb_genes  %>% mutate(source = "February"),
    june_genes %>% mutate(source = "June"),
    aug_genes  %>% mutate(source = "August")
  )
  all_data %>%
    filter(Term %in% term_list) %>%
    group_by(Term) %>%
    summarise(min_pval = min(Fisher, na.rm = TRUE), .groups = "drop") %>%
    arrange(min_pval) %>%
    slice_head(n = 5) %>%
    mutate(
      group       = group_name,
      neg_log_p   = -log10(min_pval)
    )
}

top_terms <- bind_rows(
  mapply(get_top5, groups, names(groups), SIMPLIFY = FALSE)
)

top_terms$group <- factor(top_terms$group, levels = levels(bar_data$group))

# Shared x-axis limits
x_lim <- c(0, max(top_terms$neg_log_p, na.rm = TRUE) * 1.05)

# ── 5. Colour palette ──────────────────────────────────────────────────────────
group_colors <- c(
  "February Only"        = "#4E79A7",
  "June Only"            = "#F28E2B",
  "August Only"          = "#E15759",
  "February-June Shared" = "#76B7B2",
  "February-Aug Shared"  = "#A0CBE8",
  "June-August Shared"   = "#FFBE7D",
  "All Three"            = "#B07AA1"
)

# ── 6. Bar plot ────────────────────────────────────────────────────────────────
p_bar <- ggplot(bar_data, aes(x = count, y = group, fill = group)) +
  geom_col(width = 0.6, color = "black", linewidth = 0.3) +
  geom_text(aes(label = count), hjust = -0.2, size = 7, fontface = "plain") +
  scale_fill_manual(values = group_colors) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15)), limits = c(0, 160)) +
  labs(x = "Total Terms", y = NULL) +
  theme_classic(base_size = 20) +
  theme(
    legend.position  = "none",
    axis.text.y      = element_blank(),
    axis.ticks.y     = element_blank(),
    axis.title.x     = element_text(size = 19),
    axis.text.x      = element_text(size = 18),
    plot.margin      = margin(5, 5, 0, 5)
  )

# ── 7. Dot matrix — to the LEFT of the bars ───────────────────────────────────
# Shown as a vertical panel with month labels, y-axis aligned with bars
p_dot <- ggplot() +
  geom_point(data = dot_data,
             aes(x = group, y = month, color = active, size = active)) +
  geom_segment(data = line_data,
               aes(x = group, xend = group,
                   y = ymin, yend = ymax),
               color = "black", linewidth = 1.5) +
  scale_color_manual(values = c("TRUE" = "black", "FALSE" = "grey80")) +
  scale_size_manual(values  = c("TRUE" = 4, "FALSE" = 3)) +
  scale_x_discrete(limits = levels(bar_data$group)) +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme_classic(base_size = 20) +
  theme(
    legend.position    = "none",
    axis.text.y        = element_blank(),   # bar labels come from p_bar
    axis.ticks.y       = element_blank(),
    axis.line.y        = element_blank(),
    axis.text.x        = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5),
    plot.margin        = margin(5, 0, 5, 5),
    panel.grid.major.x = element_line(color = "grey92")
  )

# ── 8. GO term dot plots — ALL on the RIGHT, ordered top-to-bottom = bar order ─
# Bar order (top → bottom): February Only, August Only, June Only, then shared by count
# levels() gives bottom→top, so rev() gives top→bottom
go_order <- rev(levels(bar_data$group))

make_go_panel <- function(grp) {
  d <- top_terms %>% filter(group == grp) %>%
    mutate(Term_short = str_trunc(Term, 50))

  if (nrow(d) == 0) return(NULL)

  ggplot(d, aes(x = neg_log_p,
                y = reorder(Term_short, neg_log_p),
                color = group)) +
    geom_segment(aes(xend = 0, yend = reorder(Term_short, neg_log_p)),
                 color = "grey70", linewidth = 0.6) +
    geom_point(size = 6) +
    scale_color_manual(values = group_colors) +
    scale_x_continuous(limits = x_lim) +
    labs(x = NULL, y = NULL, title = grp) +
    theme_classic(base_size = 18) +
    theme(
      legend.position  = "none",
      plot.title       = element_text(size = 18, face = "bold", color = group_colors[grp]),
      axis.text.y      = element_text(size = 16),
      axis.text.x      = element_text(size = 16),
      plot.margin      = margin(3, 8, 3, 8)
    )
}

go_panels <- lapply(go_order, make_go_panel)
go_panels <- go_panels[!sapply(go_panels, is.null)]

# Add x-axis label only to the bottom (last) GO panel
n_panels <- length(go_panels)
go_panels[[n_panels]] <- go_panels[[n_panels]] +
  labs(x = expression(-log[10](Fisher))) +
  theme(axis.title.x = element_text(size = 18, face = "plain"))

go_stack <- wrap_plots(go_panels, ncol = 1)

# ── 9. Combine dot matrix (left) + bar plot (right) ───────────────────────────
left_panel <- wrap_plots(p_dot, p_bar, ncol = 2, widths = c(1, 2.5))

# ── 10. Final layout: [dot matrix | bars] | [all GO panels] ───────────────────
final_plot <- wrap_plots(left_panel, go_stack, ncol = 2, widths = c(1.8, 2.2)) +
  plot_annotation(
    theme    = theme(
      plot.title    = element_text(size = 22, face = "bold"),
      plot.subtitle = element_text(size = 18, color = "grey40")
    )
  )

# ── 11. Save ──────────────────────────────────────────────────────────────────
ggsave("../output/Molecular/mRNA/enrichment/seasonal_upset_go_combined.png",
       final_plot,
       width  = 20,
       height = 14,
       dpi    = 300,
       bg     = "white")
ggsave("../output/Molecular/mRNA/enrichment/seasonal_upset_go_combined.pdf",
       final_plot,
       width  = 20,
       height = 14,
       dpi    = 300,
       bg     = "white")
```

