---
title: "GO enrichment of mRNAs targeted by miRNAs"
author: "Jill Ashey"
date: "2025-05-14"
output: html_document
---

This script will perform GO enrichment for mRNAs that are targeted by miRNAs by condition (as generated from [this script](https://github.com/JillAshey/Astrangia_repo/blob/main/scripts/mRNA-miRNA_interactions_by_Condition.Rmd)).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(topGO)
library(tidyHeatmap)
library(purrr)
library(scales)
library(igraph)
library(tidygraph)
library(ggraph)
```

# TP0 v TP5 amb

Read in data - this represents the mRNA-miRNA interacting pairs with the mRNAs being the DEGs from the TP0 v TP5 amb comparison. 
```{r}
df <- read.csv("../output/Molecular/interactions/miRNA_mRNA_TP0_v_TP5_amb_pcc_lfc_annot.csv")
```

## Subset by `miRNA repression of mRNA at TP0`
```{r}
df_subset <- df %>%
  filter(miRNA_repression_inferred == "miRNA repression of mRNA at TP0")
unique_mrna <- unique(df_subset$mRNA)
```

Read in gene2go information (generated in `topGO.Rmd` file)
```{r}
go_ast <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(unique_mrna)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. Then merge with Feb genes 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_BP_En_sig_gene <- GO_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_BP_En_sig_gene$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. Then merge with Feb genes 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_CC_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_CC_En_sig_gene <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. Then merge with Feb genes 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_MF_En_sig_gene <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene, GO_CC_En_sig_gene, GO_MF_En_sig_gene_feb)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))
length(unique(GO_En_sig_gene$mRNA))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/interactions/enrichment/GO_en_sig_mRNA_repression_TP0_v_TP5.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/interactions/enrichment/GO_en_sig_mRNA_repression_TP0_v_TP5_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/interactions/enrichment/GO_en_sig_mRNA_repression_TP0_v_TP5_top10.png", last_plot(), width = 30, height = 40)
```
