---
title: "topGO enrichment of June and August ambient v heat"
author: "Jill Ashey"
date: "2025-05-05"
output: html_document
---

GO enrichment of DEGs expressed in June ambient v heat and August ambient v heat 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#devtools::install_github("stemangiola/tidyHeatmap")

library(tidyverse)
library(topGO)
library(tidyHeatmap)
library(purrr)
library(scales)
library(igraph)
library(tidygraph)
library(ggraph)
```

## June ambient v heat DEGs

Read in data. Positive LFC = upregulated in TP5 amb treatment. Negative LFC = upregulated in TP5 heat treatment
```{r}
june_deg<- read_csv("../output/Molecular/mRNA/TP5_amb_v_TP5_heat_DEG.csv")
```

Read in gene2go information (generated in `topGO.Rmd` file)
```{r}
go_ast <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

### June ambient DEGs 

Subset genes that are upregulated in TP5 ambient treatment (ie positive LFC)
```{r}
june_deg_amb <- june_deg %>%
  filter(log2FoldChange > 0) %>%
  dplyr::rename(gene_id = ...1)

length(unique(june_deg_amb$gene_id))
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(june_deg_amb$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

#### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]

# Add ontology col
GO_BP_En_sig$ontology <- "Biological Processes"
```

#### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]

# Add ontology col
GO_CC_En_sig$ontology <- "Cellular Components"
```

#### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]

# Add ontology column 
GO_MF_En_sig$ontology <- "Molecular Functions"
```

#### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig, GO_CC_En_sig, GO_MF_En_sig)

# Separate GO terms into one per row
go_long <- go_ast %>%
  separate_rows(GO, sep = ";\\s*")  # splits on semicolon and optional space

# Join go_long with GO_En_sig_gene
GO_En_sig_gene <- GO_En_sig_gene %>%
  inner_join(go_long, by = c("GO.ID" = "GO"))

# Join the datasets based on genes upregulated in June (ie those used for the functional enrichment)
GO_En_sig_gene_june_amb <- GO_En_sig_gene %>%
  filter(mRNA %in% clust_genes)
unique(GO_En_sig_gene_june_amb$Term)
length(unique(GO_En_sig_gene_june_amb$Term))

# Save as csv 
write.csv(GO_En_sig_gene_june_amb, "../output/Molecular/mRNA/enrichment/GO_en_sig_June_amb_expressed_genes.csv")
```

Plot
```{r}
plot_data <- GO_En_sig_gene_june_amb %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_ambient_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_ambient_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

### June heat DEGs 

Subset genes that are upregulated in TP5 heat treatment (ie negative LFC)
```{r}
june_deg_heat <- june_deg %>%
  filter(log2FoldChange < 0) %>%
  dplyr::rename("gene_id" = "...1")

length(unique(june_deg_heat$gene_id))
```

Read in gene2go information (generated in `topGO.Rmd` file)
```{r}
ast_gene2go <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(june_deg_heat$gene_id)

# All genes 
all_genes <- as.character(ast_gene2go$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

#### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]

# Add ontology col
GO_BP_En_sig$ontology <- "Biological Processes"
```

#### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]

# Add ontology col
GO_CC_En_sig$ontology <- "Cellular Components"
```

#### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]

# Add ontology column 
GO_MF_En_sig$ontology <- "Molecular Functions"
```

#### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig, GO_CC_En_sig, GO_MF_En_sig)

# Separate GO terms into one per row
#go_long <- go_ast %>%
#  separate_rows(GO, sep = ";\\s*")  # splits on semicolon and optional space

# Join go_long with GO_En_sig_gene
GO_En_sig_gene <- GO_En_sig_gene %>%
  inner_join(go_long, by = c("GO.ID" = "GO"))

# Join the datasets based on genes upregulated in June (ie those used for the functional enrichment)
GO_En_sig_gene_june_heat <- GO_En_sig_gene %>%
  filter(mRNA %in% clust_genes)
unique(GO_En_sig_gene_june_heat$Term)
length(unique(GO_En_sig_gene_june_heat$Term))

# Save as csv 
write.csv(GO_En_sig_gene_june_heat, "../output/Molecular/mRNA/enrichment/GO_en_sig_June_heat_expressed_genes.csv")
```

Plot
```{r}
plot_data <- GO_En_sig_gene_june_heat %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_heat_expressed_genes_vs_heat_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_heat_expressed_genes_vs_heat_top10.png", last_plot(), width = 30, height = 40)
```

### Compare enriched GO terms from June ambient and heat DEGs 

Look at any overlaps between June ambient and heat 
```{r}
june_amb_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_June_amb_expressed_genes.csv")
length(unique(june_amb_genes$Term))

june_heat_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_June_heat_expressed_genes.csv")
length(unique(june_heat_genes$Term))

june_amb_terms <- unique(june_amb_genes$Term)
june_heat_terms <- unique(june_heat_genes$Term)

overlapping_terms_june <- intersect(june_amb_terms, june_heat_terms) # 1 overlapping term
```

## August ambient v heat DEGs

Read in data. Positive LFC = upregulated in TP7 amb treatment. Negative LFC = upregulated in TP7 heat treatment
```{r}
aug_deg<- read_csv("../output/Molecular/mRNA/TP7_amb_v_TP7_heat_DEG.csv")
```

Read in gene2go information (generated in `topGO.Rmd` file)
```{r}
go_ast <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

### August ambient DEGs 

Subset genes that are upregulated in TP7 ambient treatment (ie positive LFC)
```{r}
aug_deg_amb <- aug_deg %>%
  filter(log2FoldChange > 0) %>%
  dplyr::rename(gene_id = ...1)

length(unique(aug_deg_amb$gene_id))
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(aug_deg_amb$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

#### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]

# Add ontology col
GO_BP_En_sig$ontology <- "Biological Processes"
```

#### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]

# Add ontology col
GO_CC_En_sig$ontology <- "Cellular Components"
```

#### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]

# Add ontology column 
GO_MF_En_sig$ontology <- "Molecular Functions"
```

#### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig, GO_CC_En_sig, GO_MF_En_sig)

# Separate GO terms into one per row
go_long <- go_ast %>%
  separate_rows(GO, sep = ";\\s*")  # splits on semicolon and optional space

# Join go_long with GO_En_sig_gene
GO_En_sig_gene <- GO_En_sig_gene %>%
  inner_join(go_long, by = c("GO.ID" = "GO"))

# Join the datasets based on genes upregulated in June (ie those used for the functional enrichment)
GO_En_sig_gene_aug_amb <- GO_En_sig_gene %>%
  filter(mRNA %in% clust_genes)
unique(GO_En_sig_gene_aug_amb$Term)
length(unique(GO_En_sig_gene_aug_amb$Term))

# Save as csv 
write.csv(GO_En_sig_gene_aug_amb, "../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_amb_expressed_genes.csv")
```

Plot
```{r}
plot_data <- GO_En_sig_gene_aug_amb %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_ambient_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_ambient_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

### August heat DEGs 

Subset genes that are upregulated in TP7 heat treatment (ie negative LFC)
```{r}
aug_deg_heat <- aug_deg %>%
  filter(log2FoldChange < 0) %>%
  dplyr::rename("gene_id" = "...1")

length(unique(june_deg_heat$gene_id))
```

Read in gene2go information (generated in `topGO.Rmd` file)
```{r}
ast_gene2go <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(aug_deg_heat$gene_id)

# All genes 
all_genes <- as.character(ast_gene2go$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

#### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]

# Add ontology col
GO_BP_En_sig$ontology <- "Biological Processes"
```

#### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]

# Add ontology col
GO_CC_En_sig$ontology <- "Cellular Components"
```

#### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]

# Add ontology column 
GO_MF_En_sig$ontology <- "Molecular Functions"
```

#### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig, GO_CC_En_sig, GO_MF_En_sig)

# Separate GO terms into one per row
#go_long <- go_ast %>%
#  separate_rows(GO, sep = ";\\s*")  # splits on semicolon and optional space

# Join go_long with GO_En_sig_gene
GO_En_sig_gene <- GO_En_sig_gene %>%
  inner_join(go_long, by = c("GO.ID" = "GO"))

# Join the datasets based on genes upregulated in June (ie those used for the functional enrichment)
GO_En_sig_gene_aug_heat <- GO_En_sig_gene %>%
  filter(mRNA %in% clust_genes)
unique(GO_En_sig_gene_aug_heat$Term)
length(unique(GO_En_sig_gene_aug_heat$Term))

# Save as csv 
write.csv(GO_En_sig_gene_aug_heat, "../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_heat_expressed_genes.csv")
```

Plot
```{r}
plot_data <- GO_En_sig_gene_aug_heat %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_heat_expressed_genes_vs_heat_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_heat_expressed_genes_vs_heat_top10.png", last_plot(), width = 30, height = 40)
```

### Compare enriched GO terms from August ambient and heat DEGs 

Look at any overlaps between August ambient and heat 
```{r}
aug_amb_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_amb_expressed_genes.csv")
length(unique(aug_amb_genes$Term))

aug_heat_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_Aug_heat_expressed_genes.csv")
length(unique(aug_heat_genes$Term))

aug_amb_terms <- unique(aug_amb_genes$Term)
aug_heat_terms <- unique(aug_heat_genes$Term)

overlapping_terms_aug <- intersect(aug_amb_terms, aug_heat_terms) # 1 overlapping term
```
