---
title: "topGO enrichment across seasons"
author: "Jill Ashey"
date: "2025-03-19"
output: html_document
---

GO enrichment of DEGs expressed in February, June and August 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#devtools::install_github("stemangiola/tidyHeatmap")

library(tidyverse)
library(topGO)
library(tidyHeatmap)
library(purrr)
library(scales)
library(igraph)
library(tidygraph)
library(ggraph)
library(UpSetR)
```

## Feb DEGs

Read in DEGs of Feb v June and August ambient
```{r}
feb_june <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP5_amb_DEG.csv")
feb_aug <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP7_amb_DEG.csv")
```

Subset genes that are upregulated (ie those that were expressed in Feb relative to June and August)
```{r}
feb_june <- feb_june %>%
  filter(log2FoldChange > 0)
feb_aug <- feb_aug %>%
  filter(log2FoldChange > 0)

# Bind 
feb_all <- rbind(feb_june,feb_aug)[,1]
feb_all_unique <- feb_all %>%
  dplyr::rename("gene_id" = "...1") %>%
  distinct(gene_id)
```

Read in gene2go information (generated in `topGO.Rmd` file)
```{r}
go_ast <- read.delim("../data/Molecular/Ast_gene2go.tab", sep = "\t")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(feb_all_unique$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes expressed in Feb. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Ast_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_ast_BP_En_sig` with GO and gene info. Then merge with Feb genes 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_ast_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_ast_BP_En_sig_gene_feb <- GO_ast_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_ast_BP_En_sig_gene_feb$ontology <- "Biological Processes"
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_CC_En_sig_gene_feb <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene_feb$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_MF_En_sig_gene_feb <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene_feb$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_ast_BP_En_sig_gene_feb, GO_CC_En_sig_gene_feb, GO_MF_En_sig_gene_feb)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))
length(unique(GO_En_sig_gene$mRNA))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

## June DEGs

Read in DEGs of June v Feb and August ambient
```{r}
feb_june <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP5_amb_DEG.csv")
june_aug <- read_csv("../output/Molecular/mRNA/TP5_amb_v_TP7_amb_DEG.csv")
```

Subset genes that are downregulated relative to February and upregulated relative to August 
```{r}
feb_june <- feb_june %>%
  filter(log2FoldChange < 0)
june_aug <- june_aug %>%
  filter(log2FoldChange > 0)

# Bind 
june_all <- rbind(feb_june,june_aug)[,1]
june_all_unique <- june_all %>%
  dplyr::rename(gene_id = ...1) %>%
  distinct(gene_id)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(june_all_unique$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_ast_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_ast_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_ast_BP_En_sig_gene_june <- GO_ast_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_ast_BP_En_sig_gene_june$ontology <- "Biological Processes"
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in Feb (ie those used for the functional enrichment)
GO_CC_En_sig_gene_june <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene_june$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in June (ie those used for the functional enrichment)
GO_MF_En_sig_gene_june <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene_june$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_ast_BP_En_sig_gene_june, GO_CC_En_sig_gene_june, GO_MF_En_sig_gene_june)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

## August DEGs

Read in DEGs of August v Feb and June ambient
```{r}
feb_aug <- read_csv("../output/Molecular/mRNA/TP0_amb_v_TP7_amb_DEG.csv")
june_aug <- read_csv("../output/Molecular/mRNA/TP5_amb_v_TP7_amb_DEG.csv")
```

Subset genes that are downregulated relative to February and downregulated relative to June 
```{r}
feb_aug <- feb_aug %>%
  filter(log2FoldChange < 0)
june_aug <- june_aug %>%
  filter(log2FoldChange < 0)

# Bind 
aug_all <- rbind(feb_aug,june_aug)[,1]
aug_all_unique <- aug_all %>%
  dplyr::rename(gene_id = ...1) %>%
  distinct(gene_id)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(aug_all_unique$gene_id)

# All genes 
all_genes <- as.character(go_ast$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_ast_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
ast_gene2go <- ast_gene2go %>%
  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_ast_BP_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in August (ie those used for the functional enrichment)
GO_ast_BP_En_sig_gene_aug <- GO_ast_BP_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_ast_BP_En_sig_gene_aug$ontology <- "Biological Processes"
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- ast_gene2go %>%
  left_join(GO_BP_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in August (ie those used for the functional enrichment)
GO_CC_En_sig_gene_aug <- GO_CC_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_CC_En_sig_gene_aug$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#ast_gene2go <- ast_gene2go %>%
#  separate_rows(GO, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#ast_gene2go$GO <- trimws(ast_gene2go$GO)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- ast_gene2go %>%
  left_join(GO_MF_En_sig, by = c("GO" = "GO.ID")) %>%
  na.omit()

# Join the datasets based on genes upregulated in August (ie those used for the functional enrichment)
GO_MF_En_sig_gene_aug <- GO_MF_En_sig_gene %>%
  filter(mRNA %in% clust_genes)

# Add ontology column 
GO_MF_En_sig_gene_aug$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_ast_BP_En_sig_gene_aug, GO_CC_En_sig_gene_aug, GO_MF_En_sig_gene_aug)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()
length(unique(GO_En_sig_gene$Term))

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes.csv")
```

Plot by number of genes that have enriched terms
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes_top10.pdf", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes_top10.png", last_plot(), width = 30, height = 40)
```

## Examining any overlaps in functional terms across season

```{r}
feb_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_February_expressed_genes.csv") %>%
  filter(!ontology == "Cellular Components")
length(unique(feb_genes$Term))
unique_terms_by_ontology <- tapply(feb_genes$Term, feb_genes$ontology, function(x) length(unique(x)))
blah <- feb_genes %>%
   group_by(Term,ontology) %>%
   summarise(gene_count = n_distinct(mRNA))

june_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_June_expressed_genes.csv")%>%
  filter(!ontology == "Cellular Components")
length(unique(june_genes$Term))
unique_terms_by_ontology <- tapply(june_genes$Term, june_genes$ontology, function(x) length(unique(x)))
blah2 <- june_genes %>%
   group_by(Term,ontology) %>%
   summarise(gene_count = n_distinct(mRNA))

aug_genes <- read.csv("../output/Molecular/mRNA/enrichment/GO_en_sig_August_expressed_genes.csv")%>%
  filter(!ontology == "Cellular Components")
length(unique(aug_genes$Term))
unique_terms_by_ontology <- tapply(aug_genes$Term, aug_genes$ontology, function(x) length(unique(x)))
blah3 <- aug_genes %>%
   group_by(Term,ontology) %>%
   summarise(gene_count = n_distinct(mRNA))

feb_terms <- unique(feb_genes$Term)
june_terms <- unique(june_genes$Term)
aug_terms <- unique(aug_genes$Term)

# Overlapping terms 
overlapping_terms_feb_june <- intersect(feb_terms, june_terms) # 13 overlapping terms
overlapping_terms_feb_aug <- intersect(feb_terms, aug_terms) # 1 overlapping term
overlapping_terms_june_aug <- intersect(june_terms, aug_terms) # 70 overlapping terms

# Terms shared between Feb, June and August 
common_terms <- Reduce(intersect, list(feb_terms, june_terms, aug_terms)) # 1 common term

# Terms unique to June 
june_unique_terms <- setdiff(june_terms, union(aug_terms, feb_terms))

# Terms unique to August 
aug_unique_terms <- setdiff(aug_terms, union(june_terms, feb_terms))
```

Make an upset plot
```{r}
sets_list <- list(
  February = feb_terms,
  June     = june_terms,
  August   = aug_terms
)

pset_data <- fromList(sets_list)

png("../output/Molecular/mRNA/enrichment/seasonal_upset_plot.png", width = 8, height = 6, units = "in", res = 600)
upset(upset_data,
      sets = c("February", "June", "August"),
      order.by = "freq",
      sets.bar.color = c("#8a36b6", "#f19e21", "#5eca4a"),  # <-- custom colors
      main.bar.color = "grey40",
      matrix.color = "black",
      mb.ratio = c(0.7, 0.3),      # ratio between main barplot and set size barplot
      show.numbers = "yes",        # show counts on bars
      text.scale = 1.6,            # scale up all text
      point.size = 4,              # size of matrix points
      line.size = 1)               # thickness of connecting lines
dev.off()
```



