---
title: "Weighted Jaccard Index for mRNA-miRNA networks"
author: "Jill Ashey"
date: "2025-07-16"
output: html_document
---

Steps 
- Subset samples for one treatment 
- Run PCC for those samples 
- Build networks using the PCC and miranda data 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(DESeq2)
library(pbapply)
library(igraph)
library(ggraph)
library(tidygraph)
```

Read in mRNA count data
```{r}
mrna_counts <- read.csv("../output/Molecular/mRNA/filtered_gene_counts.csv")

# Set row names
rownames(mrna_counts) <- mrna_counts[,1] #set first column that contains gene names as rownames
mrna_counts <- mrna_counts[,-1] # remove column w/ gene names 
```

Read in miRNA count data
```{r}
mirna_counts <- read.table("../output/Molecular/smRNA/shortstack/Counts.csv", header = T) 
```

Select only miRNAs with a Y in MIRNA column. Combine Name and Coords columns for miRNA name. Remove MIRNA column 
```{r}
mirna_counts <- mirna_counts %>%
  dplyr::filter(MIRNA == "Y") %>%
  mutate(miRNA_name = paste(Name, Coords, sep = "_")) %>%
  dplyr::select(-c(Name, Coords, MIRNA))

# Set row names 
rownames(mirna_counts) <- mirna_counts[,20] #set column that contains names as rownames
mirna_counts <- mirna_counts[,-20] # remove column w/ gene names 
```

Remove extra info from column names
```{r}
colnames(mirna_counts) <- gsub("_R1_001.fastq.gz_1", "", colnames(mirna_counts))
colnames(mirna_counts) <- gsub("AST.", "AST-", colnames(mirna_counts))
colnames(mrna_counts) <- gsub("AST.", "AST-", colnames(mrna_counts))
```

Remove sample that did not sequence well
```{r}
mrna_counts <- mrna_counts %>%
    dplyr::select(-"AST-1105")
```

Read in metadata 
```{r}
meta <- read.csv("../data/Molecular/RNA_metadata.csv")
meta <- dplyr::arrange(meta, ID) # rearrange metadata so IDs are sorted in descending order 
#meta$ID <- gsub("AST-", "AST.", meta$ID)

# Set variables as factors 
meta$Timepoint <- factor(meta$Timepoint, levels = c("TP0", "TP5", "TP7"))
meta$Treatment <- factor(meta$Treatment, levels = c("Acclimation", "Ambient", "Heat"))

# Remove sample that did not sequence well 
meta <- meta %>%
    dplyr::filter(!ID=="AST-1105")
```

## February network 

Subset February samples from counts dfs
```{r}
# Get TP0 sample IDs
tp0_ids <- meta$ID[meta$Timepoint == "TP0"]

# Subset columns in counts
mirna_tp0 <- mirna_counts[, tp0_ids]
mrna_tp0  <- mrna_counts[, tp0_ids]
rownames(mirna_tp0) <- rownames(mirna_counts)
rownames(mrna_tp0)  <- rownames(mrna_counts)
```

Perform rlog on counts 
```{r}
# Filter meta for TP0
tp0_meta <- meta[meta$Timepoint == "TP0", ]
tp0_ids <- tp0_meta$ID

# Subset counts to TP0 samples
mirna_tp0 <- mirna_counts[, tp0_ids]
mrna_tp0  <- mrna_counts[, tp0_ids]

# Add dummy colData if you don’t have any covariates
dds_mirna <- DESeqDataSetFromMatrix(countData = mirna_tp0,
                                    colData = tp0_meta,
                                    design = ~1)

dds_mrna  <- DESeqDataSetFromMatrix(countData = mrna_tp0,
                                    colData = tp0_meta,
                                    design = ~1)

# Run rlog (blind = TRUE for no design)
rlog_mirna <- assay(rlog(dds_mirna, blind = TRUE))
rlog_mrna  <- assay(rlog(dds_mrna, blind = TRUE))

# Check shapes
dim(rlog_mirna)
dim(rlog_mrna)

```

Run PCC on February data 
```{r}
miRNA_ids <- rownames(rlog_mirna)
mRNA_ids  <- rownames(rlog_mrna)

# Make grid of all combos
grid <- expand.grid(miRNA = miRNA_ids, mRNA = mRNA_ids, stringsAsFactors = FALSE)

# Compute PCC and p-value
results <- pbapply::pblapply(1:nrow(grid), function(i) {
  mir <- rlog_mirna[grid$miRNA[i], ]
  mr  <- rlog_mrna[grid$mRNA[i], ]
  ct  <- cor.test(mir, mr, method = "pearson")
  data.frame(miRNA = grid$miRNA[i],
             mRNA = grid$mRNA[i],
             PCC = ct$estimate,
             Pvalue = ct$p.value)
})

results_df <- do.call(rbind, results)
str(results_df)

# format miRNA names 
results_df <- results_df %>%
    mutate(miRNA = str_replace(miRNA, "^([^_]*_[^_]*)_.*$", "\\1"))
```

### Full binding 

Read in full binding miranda data 
```{r}
miranda_full <- read.delim("~/Desktop/PutnamLab/Astrangia/Molecular/miranda_strict_all_mrna_apoc_shortstack_parsed.txt", header = F)
colnames(miranda_full) <- c("miRNA", "mRNA", "score", "energy", "query_start_end", "subject_start_end", "total_bp_shared", "query_similar", "subject_similar")

# Format miranda df 
miranda_full$miRNA <- sub("^>", "", miranda_full$miRNA)  # Remove leading ">"
miranda_full$miRNA <- sub("\\..*", "", miranda_full$miRNA)  # Remove everything from the first period onwards
miranda_full$mRNA <- sub(";.*", "", miranda_full$mRNA)  # Remove everything from "::" onwards
miranda_full$mRNA <- sub("ID=", "", miranda_full$mRNA)  # Remove everything from "::" onwards
miranda_full$mRNA <- sub("model", "TU", miranda_full$mRNA)  # Remove everything from "::" onwards
```

Merge miranda results with pcc data
```{r}
full_pcc <- results_df %>%
  inner_join(miranda_full, by = c("miRNA", "mRNA"))

length(unique(full_pcc$miRNA))
length(unique(full_pcc$mRNA))
```

Subset by significant PCCs
```{r}
full_pcc_05 <- full_pcc %>%
  filter(Pvalue < 0.05)
```

Create network plot of the mRNA-miRNA interactions. 
```{r}
# Prepare edges
edges <- full_pcc_05[, c("miRNA", "mRNA", "PCC", "Pvalue", "score", "energy")]
colnames(edges) <- c("from", "to", "PCC", "Pvalue", "score", "energy")

miRNAs <- unique(edges$from)
mRNAs  <- unique(edges$to)

nodes <- data.frame(
  name = c(miRNAs, mRNAs),
  type = c(rep("miRNA", length(miRNAs)), rep("mRNA", length(mRNAs)))
)

# Create graph
graph <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE);graph

# Plot 
ggraph(graph, layout = "fr") +  # fr = Fruchterman-Reingold layout
  geom_edge_link(aes(color = PCC), arrow = arrow(length = unit(4, 'mm')),
                 end_cap = circle(3, 'mm'), width = 0.8) +
  scale_edge_color_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_node_point(aes(shape = type, color = type), size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  theme_graph() +
  labs(edge_color = "PCC") +
  ggtitle("miRNA–mRNA PCC network")
```

Another version
```{r}
# Example edges — use your filtered data
edges <- full_pcc_05[, c("miRNA", "mRNA", "PCC", "Pvalue")]
colnames(edges) <- c("from", "to", "PCC", "Pvalue")

# Create igraph object
g <- graph_from_data_frame(edges, directed = FALSE)

# Add edge weight and color
E(g)$weight <- abs(E(g)$PCC)  # Absolute PCC for line thickness
E(g)$color <- ifelse(E(g)$PCC > 0, "red", "blue")  # Red = positive, Blue = negative

# Add node type attribute
V(g)$type <- ifelse(V(g)$name %in% edges$from, "miRNA", "mRNA")

# Convert to tidygraph for ggraph
g_tbl <- as_tbl_graph(g)

# Plot
p <- ggraph(g_tbl, layout = "fr") +  # Fruchterman-Reingold layout
  geom_edge_link(aes(edge_width = weight, color = color), alpha = 0.6) +
  geom_node_point(aes(color = type), size = 5) +
  # geom_node_text(aes(label = name), repel = TRUE, size = 3) +  # Uncomment to show labels
  scale_edge_width(range = c(0.5, 3)) +
  scale_edge_color_identity() +  # Use exact edge colors we set
  scale_color_manual(values = c("miRNA" = "purple", "mRNA" = "orange")) +
  theme_graph() +
  labs(
    title = "miRNA–mRNA Interaction Network",
    subtitle = "Edge width = |PCC|, Red = positive, Blue = negative"
  )

# Show plot
p
```

### 3'UTR

Read in 3'UTR binding miranda data 
```{r}
miranda_3utr <- read.delim("../output/Molecular/interactions/miranda_strict_all_1kb_apoc_shortstack_parsed.txt", header = F)
colnames(miranda_3utr) <- c("miRNA", "mRNA", "score", "energy", "query_start_end", "subject_start_end", "total_bp_shared", "query_similar", "subject_similar")

# Format miranda df 
miranda_3utr$miRNA <- sub("^>", "", miranda_3utr$miRNA)  # Remove leading ">"
miranda_3utr$miRNA <- sub("\\..*", "", miranda_3utr$miRNA)  # Remove everything from the first period onwards
miranda_3utr$mRNA <- sub(";.*", "", miranda_3utr$mRNA)  # Remove everything from "::" onwards
miranda_3utr$mRNA <- sub("ID=", "", miranda_3utr$mRNA)  # Remove everything from "::" onwards
miranda_3utr$mRNA <- sub("model", "TU", miranda_3utr$mRNA)  # Remove everything from "::" onwards
```

Merge miranda results with pcc data
```{r}
utr_pcc <- results_df %>%
  inner_join(miranda_3utr, by = c("miRNA", "mRNA"))

length(unique(utr_pcc$miRNA))
length(unique(utr_pcc$mRNA))
```

Subset by significant PCCs
```{r}
utr_pcc_05 <- utr_pcc %>%
  filter(Pvalue < 0.05)
```

Create network plot of the mRNA-miRNA interactions. 
```{r}
# Prepare edges
edges <- utr_pcc_05[, c("miRNA", "mRNA", "PCC", "Pvalue", "score", "energy")]
colnames(edges) <- c("from", "to", "PCC", "Pvalue", "score", "energy")

miRNAs <- unique(edges$from)
mRNAs  <- unique(edges$to)

nodes <- data.frame(
  name = c(miRNAs, mRNAs),
  type = c(rep("miRNA", length(miRNAs)), rep("mRNA", length(mRNAs)))
)

# Create graph
graph <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE);graph

# Plot 
ggraph(graph, layout = "fr") +  # fr = Fruchterman-Reingold layout
  geom_edge_link(aes(color = PCC), arrow = arrow(length = unit(4, 'mm')),
                 end_cap = circle(3, 'mm'), width = 0.8) +
  scale_edge_color_gradient2(low = "red", mid = "grey", high = "blue", midpoint = 0) +
  geom_node_point(aes(shape = type, color = type), size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  theme_graph() +
  labs(edge_color = "PCC") +
  ggtitle("miRNA–mRNA PCC network")
```

Another version
```{r}
# Example edges — use your filtered data
edges <- utr_pcc_05[, c("miRNA", "mRNA", "PCC", "Pvalue")]
colnames(edges) <- c("from", "to", "PCC", "Pvalue")

# Create igraph object
g <- graph_from_data_frame(edges, directed = FALSE)

# Add edge weight and color
E(g)$weight <- abs(E(g)$PCC)  # Absolute PCC for line thickness
E(g)$color <- ifelse(E(g)$PCC > 0, "red", "blue")  # Red = positive, Blue = negative

# Add node type attribute
V(g)$type <- ifelse(V(g)$name %in% edges$from, "miRNA", "mRNA")

# Convert to tidygraph for ggraph
g_tbl <- as_tbl_graph(g)

# Plot
p <- ggraph(g_tbl, layout = "fr") +  # Fruchterman-Reingold layout
  geom_edge_link(aes(edge_width = weight, color = color), alpha = 0.6) +
  geom_node_point(aes(color = type), size = 5) +
  # geom_node_text(aes(label = name), repel = TRUE, size = 3) +  # Uncomment to show labels
  scale_edge_width(range = c(0.5, 3)) +
  scale_edge_color_identity() +  # Use exact edge colors we set
  scale_color_manual(values = c("miRNA" = "purple", "mRNA" = "orange")) +
  theme_graph() +
  labs(
    title = "miRNA–mRNA Interaction Network",
    subtitle = "Edge width = |PCC|, Red = positive, Blue = negative"
  )

# Show plot
p
```
